---
title: "TEMPEST: Porewater ICP"
author: "2023-2025 Samples"
subtitle: "Run 1 - 2026/04/15"
date: "`r Sys.Date()`"
output: html_document
---


#SET UP

```{r setup, include=FALSE}
#Set working directory
knitr::opts_knit$set(root.dir = "C:/Users/amontg35/OneDrive - Arizona State University/Data/TEMPEST/Porewater/01_RawMicroplateData/ICP")


library(tidyverse)
library(openxlsx)
library(lubridate)
library(patchwork)
library(kableExtra)
library(tinytex)
```


#USEFUL FUNCTIONS

```{r,include=FALSE}
cal_curve<-function(Conc_uM,ABS){
    model<-lm(ABS~Conc_uM)
    slope<-summary(model)$coefficients[2, 1]
    intercept<-summary(model)$coefficients[1, 1]
     R2<-summary(model)$r.squared
  return(data.frame(slope=slope,intercept=intercept,R2=R2))
}


#function to calculate percent difference
calc_percent_diff<-function(dup,original){
  avg<-(dup+original)/2
  percent_diff<-((abs(original-dup))/avg)*100
  return(percent_diff=percent_diff)
}


#Percent error
calc_percent_error<-function(chk,expected){
  percent_error<-(abs((chk-expected))/expected)*100
  return(percent_error=percent_error)
}
```


# April 15, 2026 - Run 1

## Metal Standard Curve Checks

```{r Metal Standard Curve Checks, echo = FALSE, message = FALSE, warning = FALSE,results='asis'}
run_date<-"04/15/2026"


##Read in blank standard
std_blank <- read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet = "Corrected Intensities", rows = c(1:2))

##Read in metals standards - there has to be a better way to do this
M_0.03 <- read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet = "Corrected Intensities", rows = c(3:4))
M_0.1 <- read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet = "Corrected Intensities", rows = c(5:6))
M_0.3_10 <- read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet = "Corrected Intensities", rows = c(7:11))
M_30 <- read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet = "Corrected Intensities", rows = c(12:13))
M_100 <- read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet = "Corrected Intensities", rows = c(14:15))
 
#COMBINE STANDARDS
M_stds <- bind_rows(M_0.03,M_0.1,M_0.3_10,M_30,M_100)


M_stds <- M_stds %>%
  #COMBINE WITH STANDARD BLANK
  #SELECT ONLY COLUMNS MEASURED IN M_STDS
  bind_rows(std_blank %>% select(any_of(colnames(M_stds))))%>% 
  #ADD EXPECTED CONCENTRATIONS OF STANDARDS
    separate_wider_delim(Sample.Id,names = c("Curve","Expected_Conc_ppm"),delim = " ",
                       too_many = "merge")%>%
  mutate(Expected_Conc_ppm = if_else(str_detect(Expected_Conc_ppm,"Blank"),0,as.numeric(Expected_Conc_ppm)))%>%
  #REMOVE INTERNAL STANDARDS
  select(-contains("(IS)"))%>%
  select(Curve, Expected_Conc_ppm,contains("(cps)"))%>%
  pivot_longer(cols = contains("(cps)"),
    names_to = "Element",
    values_to = "cps",
    values_drop_na = TRUE)%>%
  mutate(Element = str_remove(Element,fixed(".(cps)")))

#CALCULATE R2
M_stds_curve<-M_stds%>%
  group_by(Element)%>%
  summarise(cal_curve(Expected_Conc_ppm,cps))%>%
  mutate(across(where(is.numeric),round, digits=5))%>%
  mutate(Run_Date = run_date)


#labels for plots 
data_text<-data.frame(
  Element = M_stds_curve$Element,
  label = round(M_stds_curve$R2,4)
)


p<-ggplot(M_stds,aes(x=Expected_Conc_ppm,y=cps))+
  geom_point()+
  geom_smooth(method="lm",se=FALSE)+
  facet_wrap(~Element,scales="free_y")+
  geom_text(
    data    = data_text,
    mapping = aes(x = Inf, y = Inf, label = label),
    hjust   = 1.1,
    vjust   = 5)+
  ggtitle(run_date)
  

print(p)

cat("\n\n **TABLE OF CALIBRATION CURVE VALUES** \n")


M_stds_curve%>%
  mutate(R2=cell_spec(R2,color = ifelse(R2<0.999,"red","black")))%>%
  kbl(caption = "Red values are <0.999",
    escape = FALSE)%>%
  kable_styling()
 
```

## P, S, and Si Standard Curve Checks
```{r P Standard Curve Checks, echo = FALSE, message = FALSE, warning = FALSE,results='asis'}

#READ IN P STANDARDS
P_stds <- read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet = "Corrected Intensities", rows = c(16:23))


#COMBINE WITH BLANK
P_stds <- P_stds %>%
  bind_rows(std_blank)%>%
  select(Sample.Id:`Ga.radial.(IS).(cps)`)%>%
  #REMOVE INTERNATL STANDARDS
  select(-contains("(IS)"))%>%
  #ADD EXPECTED CONCENTRATIONS OF STANDARDS
    separate_wider_delim(Sample.Id,names = c("Curve","Expected_Conc_ppm"),delim = " ",
                       too_many = "merge")%>%
  mutate(Expected_Conc_ppm = if_else(str_detect(Expected_Conc_ppm,"Blank"),0,as.numeric(Expected_Conc_ppm)))%>%
  select(Curve, Expected_Conc_ppm,contains("(cps)"))%>%
  pivot_longer(cols = contains("(cps)"),
    names_to = "Element",
    values_to = "cps",
    values_drop_na = TRUE)%>%
  mutate(Element = str_remove(Element,fixed(".(cps)")))


#CALCULATE R2
P_stds_curve<-P_stds%>%
  group_by(Element)%>%
  summarise(cal_curve(Expected_Conc_ppm,cps))%>%
  mutate(across(where(is.numeric),round, digits=5))%>%
  mutate(Run_Date = run_date)



#LABELS FOR PLOTS
data_text<-data.frame(
  Element = P_stds_curve$Element,
  label = round(P_stds_curve$R2,4)
)

#PLOT
p<-ggplot(P_stds,aes(x=Expected_Conc_ppm,y=cps))+
  geom_point()+
  geom_smooth(method="lm",se=FALSE)+
  facet_wrap(~Element,scales="free_y")+
  geom_text(
    data    = data_text,
    mapping = aes(x = Inf, y = Inf, label = label),
    hjust   = 1.1,
    vjust   = 5)+
  ggtitle(run_date)
  

print(p)

cat("\n\n **TABLE OF CALIBRATION CURVE VALUES** \n")

P_stds_curve%>%
  mutate(R2=cell_spec(R2,color = ifelse(R2<0.999,"red","black")))%>%
  kbl(caption = "Red values are <0.999",
      escape = FALSE)%>%
  kable_styling()


```



```{r,include=FALSE}
#Combine standard curves into one data frame
stds_curve<-rbind(M_stds_curve,P_stds_curve)



```





#Import sample data

## Counts

```{r, echo=FALSE, warning = FALSE,results='asis',fig.height = 8,width = 8}
dat.cps<-read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet =  "Corrected Intensities", startRow = 24)


dat.cps <- dat.cps%>%
  select(-1)



#range of counts for standards for each element

#M Standards
M_std_range <- M_stds%>%
  group_by(Element)%>%
  summarise(min.cps = min(cps),
            min_conc = Expected_Conc_ppm[which.min(cps)],
            max.cps = max(cps),
            max_conc = Expected_Conc_ppm[which.max(cps)])


# P Standards 
P_std_range <-P_stds%>%
  group_by(Element)%>%
  summarise(min.cps = min(cps),
            min_conc = Expected_Conc_ppm[which.min(cps)],
            max.cps = max(cps),
            max_conc = Expected_Conc_ppm[which.max(cps)])



#check that samples are all within range of standards

## M 

M.dat.cps <- dat.cps %>%
  select(-contains("(IS)"), -contains("S."), -contains("P."), -contains("Si."))%>%
  filter(str_detect(Sample.Id,"TMP_"))%>%
  filter(!str_detect(Sample.Id, "_spk"))%>%
    mutate(rep = row_number())%>%
  pivot_longer(cols = contains("(cps)"),
    names_to = "Element",
    values_to = "cps",
    values_drop_na = TRUE)%>%
  mutate(Element = str_remove(Element,fixed(".(cps)")))%>%
  left_join(M_std_range, by = "Element")%>%
  mutate(cps_flag = case_when(
    cps < min.cps ~ "BELOW RANGE",
    cps > max.cps ~ "ABOVE RANGE",
    cps > min.cps & cps < max.cps ~ "IN RANGE"))%>%
  select(-R,-Acquisition.Time,-QC.Status,-Dataset.File)

p<-ggplot(M.dat.cps,aes(x=rep,y=cps))+
  facet_wrap(~Element,scales="free")+
  geom_bar(stat = "identity", position = "dodge",aes(fill = cps_flag))+
  scale_fill_manual(values = c("BELOW RANGE" = "darkred",
                               "ABOVE RANGE" = "firebrick2",
                               "IN RANGE" = "darkgreen"))+
  geom_hline(data=M_std_range, aes(yintercept = min.cps), linetype="dashed", color = "black", linewidth=1)+
    geom_hline(data=M_std_range %>% 
                 filter(str_detect(Element, "Ca|K|Mg|Na" )), 
               aes(yintercept = max.cps), 
               linetype="dashed", 
               color = "black", 
               linewidth=1)+
  labs(caption = "Dashed lines = Minimum and maximum counts of standard curve \n\nIf no upper limit, values were much lower than highest standard",
       title = "Metals")


print(p)
  

#P

P.dat.cps <- dat.cps %>%
  select(-contains("(IS)"))%>%
  select(Sample.Id, Method.File,     `P.178.221.(cps)`,`P.213.617.(cps)`,`S.axial.(cps)`,`S.radial.(cps)`,`Si.axial.(cps)`)%>%
  #only include samples
  filter(str_detect(Sample.Id,"TMP_"))%>%
  #remove spike
  filter(!str_detect(Sample.Id, "_spk"))%>%
    mutate(rep = row_number())%>%
  pivot_longer(cols = contains("(cps)"),
    names_to = "Element",
    values_to = "cps",
    values_drop_na = TRUE)%>%
  mutate(Element = str_remove(Element,fixed(".(cps)")))%>%
  left_join(P_std_range, by = "Element")%>%
  mutate(cps_flag = case_when(
    cps < min.cps ~ "BELOW RANGE",
    cps > max.cps ~ "ABOVE RANGE",
    cps > min.cps & cps < max.cps ~ "IN RANGE"))

p<-ggplot(P.dat.cps,aes(x=rep,y=cps))+
  facet_wrap(~Element,scales="free")+
  geom_bar(stat = "identity", position = "dodge",aes(fill = cps_flag))+
  scale_fill_manual(values = c("BELOW RANGE" = "darkred",
                               "ABOVE RANGE" = "firebrick2",
                               "IN RANGE" = "darkgreen"))+
  geom_hline(data=P_std_range, aes(yintercept = min.cps), linetype="dashed", color = "black", linewidth=1)+
    geom_hline(data=P_std_range %>%
                 filter(!Element == "P.213.617",!Element =="Si.axial"), 
               aes(yintercept = max.cps), 
               linetype="dashed", 
               color = "black", 
               linewidth=1)+
  labs(caption = "Dashed lines = Minimum and maximum counts of standard curve \n\nIf no upper limit, values were much lower than highest standard",
       title = "P, S, and Si")


print(p)











cps_percent <- rbind(M.dat.cps,P.dat.cps)%>%
  group_by(Element)%>%
  summarise(Pass_percent = round(mean(cps_flag == "IN RANGE")*100,1))%>%
  mutate(range_flag = ifelse(Pass_percent>80,"PASS","FAIL"))%>%
  select(Element,Pass_percent,range_flag)




cps_percent%>%
  mutate(Element=cell_spec(Element,color =ifelse(Pass_percent< 80,"red","black")))%>%
  kbl(caption = "Red elements had <80% of samples within range of standard curve",
      escape= FALSE)%>%
  kable_styling()




```

```{r, include = FALSE}
#just Fe 

Fe.dat.cps <- M.dat.cps %>%
  filter(Element == "Fe.axial")%>%
  filter(str_detect(Sample.Id,"TMP_"))%>%
  filter(!str_detect(Sample.Id,"Spike|_spk"))%>%
  mutate(rep = row_number())%>%
  mutate(Element = str_remove(Element,fixed(".(cps)")))%>%
  mutate(Fe_flag = case_when(
    cps < min.cps ~ "BELOW RANGE",
    cps > max.cps ~ "ABOVE RANGE",
    ## cps of 0.01 ppm Fe standard
    cps > min.cps & cps < 361.1 ~ "QUESTIONABLE",
    cps > 361.1 & cps < max.cps ~ "IN RANGE"))


p <- Fe.dat.cps%>%
  ggplot(aes(x=rep,y=cps))+
  facet_wrap(~Element,scales="free")+
  geom_bar(stat = "identity", position = "dodge",aes(fill = Fe_flag))+
  scale_fill_manual(values = c("BELOW RANGE" = "darkred",
                               "ABOVE RANGE" = "firebrick2",
                               "IN RANGE" = "darkgreen",
                               "QUESTIONABLE" = "goldenrod2"))+
  geom_hline(yintercept=361.1, linetype = "dashed",color ='black',linewidth = 1)

print(p)
  
```



## Concentrations
```{r,include=FALSE}
dat<-read.xlsx("20260415_AM_TMP_PW_Results.xlsx", sheet =  "Conc. in Calib Units", startRow = 24)

dat <- dat %>% 
  select(-1)
```


## Asses Metal Check Standards 
```{r,echo=FALSE, warning = FALSE,results='asis'}

# Pull out M check standards
M_chks_raw <- dat %>%
  filter(str_detect(Sample.Id,"M "))

# Remove percentage values and any S/P/Si stds
M_chks_raw <- M_chks_raw %>% 
  select(-contains("(IS)"), -contains("S."), -contains("P."), -contains("Si."))%>%
  
  # ADD EXPECTED CONCENTRAITON
  separate_wider_delim(Sample.Id,names = c("Curve","Expected_Conc_ppm"),delim = " ",
                       too_many = "merge")%>%
  mutate(Expected_Conc_ppm = if_else(str_detect(Expected_Conc_ppm,"Blank"),0,as.numeric(Expected_Conc_ppm)))%>%
  mutate(rep = row_number())


# CALCULATE PERCENT ERROR 
 
M_chks_pe<- M_chks_raw %>%
  select(rep, Curve, Expected_Conc_ppm,contains("mg/L)"))%>%
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm",
    values_drop_na = TRUE)%>%
  #CONCENTRATION OF SR IS LOWER THAN OTHER ELEMENTS
  mutate(Expected_Conc_ppm = if_else(str_detect(Element,"Sr"),Expected_Conc_ppm/10,Expected_Conc_ppm))%>%
  mutate(percent_error = calc_percent_error(ppm,Expected_Conc_ppm))%>%
  mutate(M_chks_flag = case_when(
    abs(percent_error) <=10 ~ "PASS",
    abs(percent_error)>10 & abs(percent_error)<=15 ~ "OKAY",
    abs(percent_error)>15 ~ "FAIL"
  ))

#data frame of expected concentrations for hline on graph
#necessary because Sr has a different expected concentrations from other elements
hline_dat<-data.frame(
  Element = M_chks_pe$Element,
  intercept_value = M_chks_pe$Expected_Conc_ppm
)

  
p<-ggplot(M_chks_pe, aes(x=rep,y=ppm))+
  facet_wrap(~Element,scales = "free_y")+
  #dashed line for expected concentration
  geom_hline(data=hline_dat, aes(yintercept = intercept_value), linetype="dashed", color = "black", linewidth=1) + 
  geom_bar(stat = "identity", position = position_dodge(),aes(fill=M_chks_flag))+
  scale_fill_manual(values = c("FAIL" = "darkred", 
                               "OKAY" = "goldenrod2",
                               "PASS" = "chartreuse4"),
                    labels = c(">15 %",
                               "<15 %",
                               "<10 %"),
                    name = "% error")+
  labs(x="",y="Concentration (mg/L)")+
  ggtitle("Metal Check Standards")

print(p)


chks_percent <- M_chks_pe%>%
  mutate(Element = str_remove(Element,fixed(".(mg/L)")))%>%
  group_by(Element)%>%
  summarise(Fail_percent =mean(abs(percent_error)>15)*100,
            Pass_percent = 100 - Fail_percent,
            RSD = (abs(sd(ppm))/mean(ppm))*100)%>%
  mutate(PE_flag = ifelse(Pass_percent >= 80,"PASS","FAIL"),
         RSD_flag = ifelse(RSD<=15,"PASS","FAIL"))
 




cat("\n\n**LIST OF ELEMENTS WITH <80% OF CHECK STANDARDS WITHIN 15% OF EXPECTED CONCENTRATION - REASSESS** \n\n")


chks_percent %>%
  filter(PE_flag == "FAIL")%>%
  select(Element)%>%
  kbl()%>%
  kable_styling()


```

## Assess P, S, and Si Check Standards
```{r,echo=FALSE, warning = FALSE,results='asis'}
# Pull out M check standards
P_chks_raw <- dat %>%
  filter(str_detect(Sample.Id,"P "))

# Remove percentage values and any M stds
P_chks_raw <- P_chks_raw %>% 
  select(1:6, `P.178.221.(mg/L)`,`P.213.617.(mg/L)`,`S.radial.(mg/L)`,`S.axial.(mg/L)`,`Si.axial.(mg/L)` )%>%
  
  # ADD EXPECTED CONCENTRAITON
  separate_wider_delim(Sample.Id,names = c("Curve","Expected_Conc_ppm"),delim = " ",
                       too_many = "merge")%>%
  mutate(Expected_Conc_ppm = if_else(str_detect(Expected_Conc_ppm,"Blank"),0,as.numeric(Expected_Conc_ppm)))%>%
  mutate(rep = row_number())


# CALCULATE PERCENT ERROR 
 
P_chks_pe<- P_chks_raw %>%
  select(rep, Curve, Expected_Conc_ppm,contains("mg/L)"))%>%
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm",
    values_drop_na = TRUE)%>%
  #PERCENT ERROR CALCULATIONS
  mutate(percent_error = calc_percent_error(ppm,Expected_Conc_ppm))%>%
  mutate(P_chks_flag = case_when(
    abs(percent_error) <=10 ~ "PASS",
    abs(percent_error)>10 & abs(percent_error)<=15 ~ "OKAY",
    abs(percent_error)>15 ~ "FAIL"
  ))

#data frame of expected concentrations for hline on graph
#necessary because Sr has a different expected concentrations from other elements
hline_dat<-data.frame(
  Element = P_chks_pe$Element,
  intercept_value = P_chks_pe$Expected_Conc_ppm
)

  
p<-ggplot(P_chks_pe, aes(x=rep,y=ppm))+
  facet_wrap(~Element,scales = "free_y")+
  #dashed line for expected concentration
  geom_hline(data=hline_dat, aes(yintercept = intercept_value), linetype="dashed", color = "black", linewidth=1) + 
  geom_bar(stat = "identity", position = position_dodge(),aes(fill=P_chks_flag))+
  scale_fill_manual(values = c("FAIL" = "darkred", 
                               "OKAY" = "goldenrod2",
                               "PASS" = "chartreuse4"),
                    labels = c(">15 %",
                               "<15 %",
                               "<10 %"),
                    name = "% error")+
  labs(x="",y="Concentration (mg/L)")+
  ggtitle("P, S, Si Check Standards")

print(p)


P_chks_percent <- P_chks_pe%>%
  mutate(Element = str_remove(Element,fixed(".(mg/L)")))%>%
  group_by(Element)%>%
  summarise(Fail_percent =mean(abs(percent_error)>15)*100,
            Pass_percent = 100 - Fail_percent,
            RSD = (abs(sd(ppm))/mean(ppm))*100)%>%
  mutate(PE_flag = ifelse(Pass_percent >= 80,"PASS","FAIL"),
         RSD_flag = ifelse(RSD<=15,"PASS","FAIL"))
 


cat("\n\n**LIST OF ELEMENTS WITH <80% OF CHECK STANDARDS WITHIN 15% OF EXPECTED CONCENTRATION - REASSESS** \n\n")

P_chks_percent %>%
  filter(PE_flag == "FAIL")%>%
  select(Element)%>%
  kbl()%>%
  kable_styling()



```

## Check Internal Standards
```{r,echo=FALSE, warning = FALSE,results='asis'}
IS_stds<- dat %>%
  select(1:6, contains("(IS"))%>%
   mutate(rep = row_number())%>%
  pivot_longer(cols = contains("(IS)"),
    names_to = "Element",
    values_to = "Percent",
    values_drop_na = TRUE)%>%
  mutate(IS_flag = case_when(
    Percent > 110 | Percent < 90 ~ "FAIL",
    Percent < 110 & Percent > 90 ~ "PASS"
  ))


p_IS<-ggplot(IS_stds, aes(x=rep,y=Percent))+
  geom_hline(yintercept = 100, color = 'black',linetype='dashed')+
  geom_line()+
  geom_point(aes(color=IS_flag))+
  facet_wrap(~Element)+
  scale_color_manual(values = c("PASS" = "chartreuse4", "FAIL" = "darkred"),
                    name = "% error < 10%")+
  coord_cartesian(ylim=c(80,120))

print(p_IS)





cat("\n\n",paste(IS_stds%>%
  summarise(Pass_percent = mean(IS_flag=="PASS")*100), "% of Internal Standards measurements are within 10% of 100"))



```



## Additional Checks 

### NIST 1643f and Multi-elemental standard (QCS) are used as additional checks with a 10x dilution 

```{r,echo=FALSE, warning = FALSE,fig.width=10,results='asis'}
DF_chks<-10 # dilution factor 

#Make data frame of expected values for each additional check

##QCS - Multi-elemental standard
QCS<- data.frame(
  Element = c("Al", "B","Ca","Fe","K","Mg","Mn","Na","P","S","Si","Sr","Zn"),
  Expected_Conc_ppm = c(99.8, 49, 98.9,99.9,49.78,99.0,10.03,101.1,49.50,99.1,99.1,9.92,50.3)
)

QCS<-QCS%>%
  #Account for dilution factor (10)
  mutate(Expected_Conc_ppm = Expected_Conc_ppm/DF_chks )


#Pull out check samples 
QCS_dat<-dat%>%
  filter(str_detect(Sample.Id,"QCS"))%>%
  #remove internal standards
  select(-contains("(IS)"))%>%
  mutate(rep=row_number())%>%
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm",
    values_drop_na = TRUE)%>%
  #MERGE WITH DATA FRAME OF EXPECTED VALUES
  separate_wider_delim(Element,delim =".",names = c("Element","Measurement.type","Units"),too_many = "merge")%>%
  left_join(QCS, by= "Element")%>%
  unite("Element",Element,Measurement.type,Units,sep=".")

QCS_pe<-QCS_dat%>%
  mutate(QCS_pe = calc_percent_error(ppm,Expected_Conc_ppm))%>%
  mutate(QCS_flag = case_when(
    abs(QCS_pe)>15 ~ "FAIL",
    abs(QCS_pe)<=10 ~ "PASS",
    abs(QCS_pe)>10 & abs(QCS_pe)<=15 ~ "OKAY"
  ))


hline_dat<-data.frame(
  Element = QCS_pe$Element,
  intercept_value = QCS_pe$Expected_Conc_ppm
)

  
p<-ggplot(QCS_pe, aes(x=rep,y=ppm))+
  facet_wrap(~Element,scales = "free_y")+
  #dashed line for expected concentration
  geom_hline(data=hline_dat, aes(yintercept = intercept_value), linetype="dashed", color = "black", linewidth=1) + 
  geom_bar(stat = "identity", position = position_dodge(),aes(fill=QCS_flag))+
  scale_fill_manual(values = c("FAIL" = "darkred", 
                               "OKAY" = "goldenrod2",
                               "PASS" = "chartreuse4"),
                    labels = c(">15 %",
                               "<15 %",
                               "<10 %"),
                    name = "% error")+
  labs(x="",y="Concentration (mg/L)")+
  ggtitle("QCS Multi-elemental Standard (1:10) check")

print(p)


QCS_percent <- QCS_pe%>%
  mutate(Element = str_remove(Element,fixed(".(mg/L)")))%>%
  group_by(Element)%>%
  summarise(percenterror_mean = mean(abs(QCS_pe)),
            Fail_percent = mean(abs(QCS_pe)>15)*100,
            Pass_percent = 100 - Fail_percent,
            RSD = (abs(sd(ppm))/mean(ppm))*100)%>%
  mutate(QCS_flag = ifelse(Pass_percent >= 80,"PASS","FAIL"),
         RSD_flag = ifelse(RSD<=15,"PASS","FAIL"))
 



cat("\n\n**LIST OF ELEMENTS WITH <80% OF CHECK STANDARDS WITHIN 15% OF EXPECTED CONCENTRATION - REASSESS**")


QCS_percent %>%
  filter(QCS_flag == "FAIL")%>%
  select(Element,percenterror_mean)%>%
  mutate(percenterror_mean =round(percenterror_mean,digits = 2))%>%
  rename("Avg % Error"=percenterror_mean)%>%
  kbl()%>%
  kable_styling()
  

```




```{r,echo=FALSE, warning = FALSE,fig.width=10,results='asis'}
##NIST 1643f 
NIST<-data.frame(
  Element = c("Al", "B","Ca","Fe","K","Mg","Mn","Na","Sr","Zn"),
  Expected_Conc_ug.L = c(133.8,152.3,29430,93.44,1932.6,7454,37.14,18830,314,74.4)
)

NIST<-NIST %>%
  #convert to ppm (mg/L)
  group_by(Element)%>%
  summarise(Expected_Conc_ppm = Expected_Conc_ug.L/1000)%>%
  #account for dilution factor (10)
  mutate(Expected_Conc_ppm = Expected_Conc_ppm/DF_chks)
  

NIST_dat<-dat%>%
  filter(str_detect(Sample.Id,"NIST"))%>%
  #remove internal standards
  select(-contains("(IS)"))%>%
  mutate(rep=row_number())%>%
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm",
    values_drop_na = TRUE)%>%
  #MERGE WITH DATA FRAME OF EXPECTED VALUES
  separate_wider_delim(Element,delim =".",names = c("Element","Measurement.type","Units"),too_many = "merge")%>%
  left_join(NIST, by= "Element")%>%
  unite("Element",Element,Measurement.type,Units,sep=".")

NIST_pe<-NIST_dat%>%
  filter(!is.na(Expected_Conc_ppm))%>%
  mutate(NIST_pe = calc_percent_error(ppm,Expected_Conc_ppm))%>%
  mutate(NIST_flag = case_when(
    abs(NIST_pe)>15 ~ "FAIL",
    abs(NIST_pe)<=10 ~ "PASS",
    abs(NIST_pe)>10 & abs(NIST_pe)<=15 ~ "OKAY"
  ))


hline_dat<-data.frame(
  Element = NIST_pe$Element,
  intercept_value = NIST_pe$Expected_Conc_ppm
)

  
p<-ggplot(NIST_pe,aes(x=rep,y=ppm))+
  facet_wrap(~Element,scales = "free_y",drop=TRUE)+
  #dashed line for expected concentration
  geom_hline(data=hline_dat,aes(yintercept = intercept_value), linetype="dashed", color = "black", linewidth=1) + 
  geom_bar(stat = "identity",aes(fill=NIST_flag))+
  scale_fill_manual(values = c("FAIL" = "darkred", 
                               "OKAY" = "goldenrod2",
                               "PASS" = "chartreuse4"),
                    labels = c(">15 %",
                               "<15 %",
                               "<10 %"),
                    name = "% error")+
  labs(x="",y="Concentration (mg/L)",title = "NIST 1643f Checks")


print(p)


NIST_percent <- NIST_pe%>%
  mutate(Element = str_remove(Element,fixed(".(mg/L)")))%>%
  group_by(Element)%>%
  summarise(percenterror_mean = mean(abs(NIST_pe)),
    Fail_percent = mean(abs(NIST_pe)>15)*100,
    Pass_percent = 100 - Fail_percent,
            RSD = (abs(sd(ppm))/mean(ppm))*100)%>%
  mutate(NIST_flag = ifelse(Pass_percent >= 80,"PASS","FAIL"),
         RSD_flag = ifelse(RSD<=15,"PASS","FAIL"))
 


cat("\n\n**LIST OF ELEMENTS WITH <80% OF NIST CHECKS WITHIN 15% OF EXPECTED CONCENTRATION - REASSESS**")


NIST_percent %>%
  filter(NIST_flag == "FAIL")%>%
  select(Element,percenterror_mean)%>%
  mutate(percenterror_mean =round(percenterror_mean,digits = 2))%>%
  rename("Avg % Error"=percenterror_mean)%>%
  kbl()%>%
  kable_styling()
  

```

### Duplicate samples 


```{r,echo=FALSE, warning = FALSE,fig.width=10,results='asis'}
setwd("C:/Users/amontg35/OneDrive - Arizona State University/Data/TEMPEST/Porewater/00_Metadata")


#import DF of samples and other info 
sample_list<-read.xlsx("TEMPEST_ICP_List2.xlsx",sheet="20260415")%>%
  rename(Sample.Id = Sample.ID)
  

#combine with sample concentrations 
dat2<-left_join(dat,sample_list,by="Sample.Id")%>%
  #FIX SAMPLE IDs ON DUPES AND MISSING SAMPLES 
  #CORRECT IDs are in Notes column
  mutate(Sample.Id = if_else(is.na(Notes),Sample.Id,Notes))


#pull out dupes
dups<-dat2%>%
  filter(str_detect(Sample.Id,"_dup"))%>%
  select(!contains("(IS)"))%>%
  #REMOVE _DUP FROM SAMPLE IDS
  mutate(Sample.Id = gsub("_dup","",Sample.Id),
         rep = row_number())%>%
  #CALCULATE CONCENTRATIONS WITH DILUTION FACTOR
  mutate(across(contains("(mg/L"), ~ .x * DF))%>%
  #pivot longer to make calculations and plotting easier
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm_dup",
    values_drop_na = TRUE)


#remove dupes from dat2
dups_conc<-dat2%>%
  filter(!str_detect(Sample.Id,"_dup|_spk"))%>%
  select(!contains("(IS)"))%>%
  #ACCOUNT FOR DILUTION FACTOR IN CONCENTRATION
  mutate(across(contains("(mg/L)"),~ .x * DF))%>%
  #pivot longer to make calculations and plotting easier
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm",
    values_drop_na = TRUE)%>%
  #Combine with duplicate samples 
  right_join(dups,by=c("Sample.Id","Element"))%>%
  #calculate percent difference
  mutate(dup_percentdiff = calc_percent_diff(ppm_dup,ppm))



p<-dups_conc%>%
  pivot_longer(cols = contains("ppm"),
    names_to = "Type",
    values_to = "ppm",
    values_drop_na = TRUE)%>%
  ggplot(aes(x=rep,y=ppm))+
  facet_wrap(~Element,scales = "free_y")+
  geom_bar(stat="identity",position = "dodge",aes(fill=Type))+
  scale_fill_manual(values = c("ppm" = "#6c91c2", "ppm_dup" = "#ff934f"),
                    name = " ",
                    labels = c("Original","Duplicate"))+
  labs(x="",y="Concentration (mg/L)",title = "Duplicate Samples")



print(p)


dup_percent <- dups_conc%>%
  mutate(Element = str_remove(Element,fixed(".(mg/L)")))%>%
  group_by(Element)%>%
  summarise(percentdiff_mean = mean(abs(dup_percentdiff)),
            Fail_percent = mean(abs(dup_percentdiff)>15)*100,
            Pass_percent = 100 - Fail_percent)%>%
  mutate(dup_flag = ifelse(Pass_percent >= 80,"PASS","FAIL"))%>%
  arrange(dup_flag)
 



cat("\n\n**LIST OF ELEMENTS WITH <80% OF DUPLICATE SAMPLES WITHIN 15% OF SAMPLE CONCENTRATION - REASSESS**")


dup_percent %>%
  filter(dup_flag == "FAIL")%>%
  select(Element,percentdiff_mean)%>%
  mutate(percentdiff_mean =round(percentdiff_mean,digits = 2))%>%
  rename("Avg % Difference"=percentdiff_mean)%>%
  kbl()%>%
  kable_styling()
  
  


```

### Sample Spikes

*Samples spiked with 0.2 mL of LFB (M10)*

```{r,echo=FALSE, warning = FALSE,fig.width=10,results='asis'}
#pull out LFB used for sample spike
# only metals

spk.mL <-0.2
total.mL <-7
spk.ppm <-10


LFB<-dat2 %>%
  filter(Sample.Id == "LFB")%>%
  select(-contains("(IS)"), -contains("S."), -contains("P."), -contains("Si."))%>%
  #CALCULATE LFB CONCENTRATION *NO DF* 
  summarise(across(contains("(mg/L)"),mean))%>%
  #pivot longer to make calculations and plotting easier
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm_LFB",
    values_drop_na = TRUE)%>%
  #EXPECTED CONCENTRAITONS BASED ON LFB CONCENTRATION 
  mutate(expected_conc_LFB = (spk.mL * ppm_LFB)/total.mL)



#pull out spikes
spks<-dat2%>%
  filter(str_detect(Sample.Id,"_spk"))%>%
  select(-contains("(IS)"), -contains("S."), -contains("P."), -contains("Si."))%>%
  #REMOVE _spk FROM SAMPLE IDS
  mutate(Sample.Id = gsub("_spk","",Sample.Id),
         rep = row_number())%>%
  #CALCULATE CONCENTRATIONS WITH DILUTION FACTOR
  mutate(across(contains("(mg/L"), ~ .x * DF))%>%
  #pivot longer to make calculations and plotting easier
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm_spk",
    values_drop_na = TRUE)


#remove dupes from dat2
spk_conc<-dat2%>%
  filter(!str_detect(Sample.Id,"_dup|_spk"))%>%
  select(!contains("(IS)"))%>%
  #ACCOUNT FOR DILUTION FACTOR IN CONCENTRATION
  mutate(across(contains("(mg/L)"),~ .x * DF))%>%
  #pivot longer to make calculations and plotting easier
  pivot_longer(cols = contains("(mg/L)"),
    names_to = "Element",
    values_to = "ppm", #SAMPLE CONCENTRATION
    values_drop_na = TRUE)%>%
  #Combine with duplicate samples 
  right_join(spks,by=c("Sample.Id","Element"))%>%
  #Combine with mean LFB values
  left_join(LFB,by=c("Element"))%>%
  #PERCENT RECOVERY OF SPIKE
  mutate(spk_recovery = ((ppm_spk-ppm)/ppm_LFB)*100)%>%
  #ADD FLAGS 
  mutate(flag = case_when(
    spk_recovery < 80 | spk_recovery > 120 ~ "FAIL",
    spk_recovery > 80 & spk_recovery < 90 ~ "OKAY",
    spk_recovery <= 120 & spk_recovery > 110 ~ "OKAY",
    spk_recovery >= 90 & spk_recovery <= 110 ~ "PASS"
  ))





p<-spk_conc%>%
  ggplot(aes(x=rep,y=spk_recovery))+
  geom_hline(yintercept = c(80,120),color='black',linetype='dashed',linewidth=1)+
  facet_wrap(~Element,scales = "free_y")+
  geom_bar(stat="identity",position = "dodge",aes(fill=flag))+
  scale_fill_manual(values = c("FAIL" = "darkred", 
                               "OKAY" = "goldenrod2",
                               "PASS" = "chartreuse4"),
                    name = "% recovery")+
  labs(x="",y="Recovery (%)",title = "Spiked Samples")



print(p)



spk_percent <- spk_conc%>%
  mutate(Element = str_remove(Element,fixed(".(mg/L)")))%>%
  group_by(Element)%>%
  summarise(percentrecovery_mean = mean(abs(spk_recovery)),
            Fail_percent = mean(flag == "FAIL")*100,
            Pass_percent = 100 - Fail_percent,
            RSD = (abs(sd(ppm_spk))/mean(ppm_spk))*100)%>%
  mutate(spk_flag = ifelse(Pass_percent >= 80,"PASS","FAIL"),
         RSD_flag = ifelse(RSD<=15,"PASS","FAIL"))
 





cat("\n\n**LIST OF ELEMENTS WITH <80% OF SPIKED SAMPLES WITH 100 (+/- 20%) PERCENT RECOVERY - REASSESS**")


spk_percent %>%
  filter(spk_flag == "FAIL")%>%
  select(Element,percentrecovery_mean)%>%
  mutate(percentrecovery_mean =round(percentrecovery_mean,digits = 2))%>%
  rename("Avg % Recovery"=percentrecovery_mean)%>%
  kbl()%>%
  kable_styling()
  



```

# **Summary of QA/QC Results**

 *Red elements did not pass QA/QC Checks*

```{r,echo=FALSE, warning = FALSE,fig.width=10,results='asis'}

#check standards
chk<-rbind(chks_percent,P_chks_percent)%>%
  dplyr::select(Element,PE_flag)%>%
  rename(chk_flag = PE_flag)%>%
  #JOIN COUNT RANGE
  left_join(cps_percent,by = "Element")%>%
  #Join NIST
  left_join(NIST_percent,by="Element")%>%
  #JOIN QCS
  left_join(QCS_percent,by="Element")%>%
  #DUPLICATES
  left_join(dup_percent,by="Element")%>%
  #SPIKES
  left_join(spk_percent,by="Element")%>%
  dplyr::select(Element,range_flag,chk_flag,NIST_flag,QCS_flag,dup_flag,spk_flag)%>%
  pivot_longer(cols = contains("flag"),
    names_to = "Flag_Type",
    values_to = "Flag",
    values_drop_na = TRUE)%>%
  mutate(Flag_Type = gsub("_flag","",Flag_Type))%>%
  group_by(Element)%>%
  #percent of checks for each element that PASS 
  summarise(num_checks = n_distinct(Flag_Type),
            Fail_percent = round(mean(Flag == "FAIL")*100),
            Passed_checks = list(Flag_Type[Flag == "PASS"]),
            Failed_checks = list(Flag_Type[Flag =="FAIL"]),
            
            #MUST PASS EITHER NIST OR QCS CHECK
            QAQC = ifelse(str_detect(Passed_checks, "NIST|QCS") & str_detect(Passed_checks, "chk"),"PASS","FAIL"))
  


chk%>%
  mutate(Element = cell_spec(Element,
                  color=ifelse(QAQC == "FAIL","red","black"),
                  bold =ifelse(QAQC == "FAIL",TRUE,FALSE)))%>%
  rename("Fail (%)" = Fail_percent,
         "Number of Checks" = num_checks)%>%
  kbl(caption = "Element must pass either NIST or QCS checks AND check standards",
      escape= FALSE)%>%
  kable_styling()



```

# Sample Processing Steps

1. Sample concentrations were compared to blank concentrations. If sample concentration < blank concentration, then the sample concentration was set to zero. 
2. Dilution factor was applied to each sample. In most cases, DF=5
3. Flags added to elments that had a poor standard curve. 
4. Elements that did not pass QAQC were removed from the dataset 
5. Removed samples that had low internal standards (<90%)
6. Column added to data listing elements that did not pass QAQC
7. Data exported to .csv



```{r,include = FALSE}
# Establish Detection Limits

blank.dat <- dat2%>%
  filter(str_detect(Sample.Id, "MB"))%>%
  #average of blanks for all elements
  summarise(across(contains("(mg/L)"), ~mean(.x)))%>%
  #set any negative values to zero
  mutate(across(contains("(mg/L)"), ~ ifelse(.x > 0, .x, 0)))%>%
  pivot_longer(cols = contains(c("(mg/L)","(IS)")),
    names_to = "Element",
    values_to = "ppm_blank", #SAMPLE CONCENTRATION
    values_drop_na = TRUE)
  



```




```{r,include=FALSE}

sample.dat <- dat2 %>%
  #filter out samples
  filter(str_detect(Sample.Id,"TMP"))%>%
  
    #remove spikes and duplicate samples
  filter(!str_detect(Sample.Id,"spk|dup"))%>%
  
   #data manipulation
  pivot_longer(cols = contains(c("(mg/L)","(IS)")),
    names_to = "Element",
    values_to = "ppm", #SAMPLE CONCENTRATION
    values_drop_na = TRUE)%>%
  #join with blanks
  left_join(blank.dat,by = "Element",multiple = "all")%>%

  #Flag samples/elements that were lower than blank averages
  mutate(bdl_Flag = ifelse(ppm < ppm_blank,"bdl",NA), 
         #all samples bdl = 0 
         ppm = ifelse(str_detect(Element,"(mg/L)")&ppm < ppm_blank, 0 , ppm),
         #apply dilution factor
         
         ppm = ifelse(str_detect(Element,"(mg/L)"),ppm*DF,ppm))%>%
  
    # remove mg.L part of Element names
  mutate(Element = str_remove(Element,fixed(".(mg/L)")))%>%
  #remove unnecessary columns 
  select(-c(R:Method.File,X1:`Total.Volume.(mL)`,spk.vol))%>%
  
 
  mutate(Element = str_replace(Element,fixed(".(IS).(%)"),".percent"))%>%
  left_join(stds_curve,by = "Element")%>%
  left_join(chk,by= "Element")%>%
  
  
  #ADD FLAGS TO DATA
  mutate(R2_Flag =if_else(R2 < 0.999, "Data not reliable",NA))%>%
  
    #FILTER OUT ELEMENTS THAT DID NOT PASS QAQC
  filter((str_detect(Element,".percent")) | is.na(R2_Flag) & QAQC=="PASS")%>%
  
  #more data manipulation
  select(-c(Notes,slope:Failed_checks))%>%
  pivot_wider(names_from = Element,values_from = c(ppm,R2_Flag,QAQC,ppm_blank,bdl_Flag),names_vary = "slowest")%>%

  
  #INTERNAL STANDARDS FLAG
  mutate(Flag_Sc.axial.percent = if_else(ppm_Sc.axial.percent < 90, "Internal Standard Low", NA),
         Flag_Y.radial.percent = if_else(ppm_Y.radial.percent < 90, "Internal Standard Low",NA),
         Flag_Ga.radial.percent = if_else(ppm_Ga.radial.percent < 90, "Internal Standard Low", NA))%>%
  rename(Sc.axial.percent = ppm_Sc.axial.percent,
         Y.radial.percent = ppm_Y.radial.percent,
         Ga.radial.percent = ppm_Ga.radial.percent)%>%
  
  #FILTER OUT SAMPLES WHEN INTERNAL STANDARDS IS BAD
  filter(is.na(Flag_Sc.axial.percent),
         is.na(Flag_Y.radial.percent),
         is.na(Flag_Ga.radial.percent))


#Make column of elements that did not pass QAQC
sample.dat$bad_elements <- list(chk$Element[chk$QAQC=="FAIL"])



# only important columns
dat.clean <- sample.dat%>%
  select(Sample.Id,starts_with("ppm_"),DF,bad_elements)%>%
  #remove blank columns
  select(-contains("_blank_"))%>%
  separate_wider_delim(Sample.Id,cols_remove = FALSE,names = c("Project","Treatment","Grid","Collection_Date","Collection_Time"),delim = "_",too_few = "align_start")%>%
  mutate(bad_elements = map_chr(bad_elements, paste, collapse = ", "),
         Collection_Date = as.Date(Collection_Date, format = "%Y%m%d" ))%>%
  #ADD RUN DATE AS COLUMN
  mutate(Run.Date = run_date)


#Export

write.csv(dat.clean,file = "C:/Users/amontg35/OneDrive - Arizona State University/Data/TEMPEST/Porewater/02_CleanMicroplateData/TMP_20260415_ICP_Clean.csv")


```


# Preliminary Plot of Data


```{r,echo = FALSE, message = FALSE, warning = FALSE,results='asis'}


p<-dat.clean%>%
  pivot_longer(cols = contains(c("ppm")),
    names_to = "Element",
    values_to = "ppm", #SAMPLE CONCENTRATION
    values_drop_na = TRUE)%>%
    ggplot(aes(x=Collection_Date,y=ppm))+
  geom_point(position = position_dodge(1),aes(color = Treatment))+
  facet_wrap(~Element,scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45,hjust =1))+
    scale_color_manual(values = c("#00AC52",
                               '#00EEEE',
                               "#BC105E"),
                      labels=c("Control",
                               "Freshwater",
                               "Saltwater"))


print(p)


```



