---
title: "TEMPEST: Porewater Nutrients"
author: "2023-2025 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true

output_dir: "To Be Reviewed/PDF"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(readxl)
library(tidyr)
library(tibble)
library(openxlsx)
library(ggh4x)
library(kableExtra)
```

## Grab 2022 and 2023 All Data Files
```{r Pull in 2022-2024 data files, message=FALSE, include=FALSE}

#read in 2023 Nutrient Data:
dat23 <- read.csv("2023/COMPASS_TEMPEST_NUTR_2023.csv", colClasses = "character")
colnames(dat23)
dat23 <- dat23 %>% 
  rename(
    NOx_Conc_mgL = NOx_mgL, 
    NOx_Conc_uM = NOx_uM, 
    NOx_Conc_Flag = NOx_range, 
    NH3_Conc_mgL = NH3_mgL, 
    NH3_Conc_uM = NH3_uM, 
    NH3_Conc_Flag = NH3_range, 
    PO4_Conc_mgL = PO4_mgL, 
    PO4_Conc_uM = PO4_uM, 
    PO4_Conc_Flag = PO4_range, 
  ) %>% 
  mutate(Plot = ifelse(Plot == "CON", "C", Plot))

#read in 2024 Nutrient Data:
dat24 <- read.csv("2024/COMPASS_TEMPEST_NUTR_2024.csv", colClasses = "character")
colnames(dat24)

#read in 2025 Nutrient Data:
dat25 <- read.csv("2025/COMPASS_TEMPEST_NUTR_2025.csv", colClasses = "character")
colnames(dat24)

all_dat <- bind_rows(dat23, dat24, dat25)

##Convert Collection Date into Date format
all_dat$Collection_Date <- as.character(all_dat$Collection_Date)
all_dat$Collection_Date_formatted <- as.Date(all_dat$Collection_Date, format = "%Y%m%d")

##Extract Year and month from Collection Date
all_dat$Year <- format(all_dat$Collection_Date_formatted, format = "%Y")
all_dat$Month <- format(all_dat$Collection_Date_formatted, format = "%m")
all_dat$Day <- format(all_dat$Collection_Date_formatted, format = "%d")


final_dat <- all_dat %>%
  mutate(
    Project = "COMPASS", 
    Site = "TEMPEST", 
    EST_EDT = NA) %>%
  rename( 
    Evacuation_Date_YYYMMDD = Evacuation_Date, 
    Collection_Date_YYYYMMDD = Collection_Date) %>%
  dplyr::select(
    Project, Site, Plot, Grid, Depth_cm, Event_Time, Time_of_day, Sample_ID, 
    Volume_mL, Evacuation_Date_YYYMMDD, Collection_Date_YYYYMMDD, 
    Collection_Start_Time_24hrs, Collection_End_Time_24hrs, EST_EDT, 
    Deployment_Date, Deployment_Time_24hrs, Collection_Time_24hrs, 
    NOx_Conc_mgL, NOx_Conc_uM, NOx_Conc_Flag, NOx_QAQC_Flag, 
    NH3_Conc_mgL, NH3_Conc_uM, NH3_Conc_Flag, NH3_QAQC_Flag, 
    PO4_Conc_mgL, PO4_Conc_uM, PO4_Conc_Flag, PO4_QAQC_Flag,
    Analysis_rundate,  Run_notes, Field_notes
        # list columns in the order you want them
  )

##Set 4 decimal points max
final_dat <- final_dat %>%
  mutate(across(where(is.numeric), round, digits = 4)) ##This does make all salnities 0 when Cl_ppm is 0 instead of 0.000026

```


##Make Relevant Metadata Sheet
```{r make metadata, include = FALSE}

#pull the column names from the all_dat dataframe and create a new dataframe with them 
metadat <- data.frame(Column_Names = colnames(final_dat))
metadat

metadat$Description <- ""
metadat$Units <- ""
metadat$Instrument_ifapplicable <- ""

colnames(final_dat)
#add information about each line: 

metadat$Description <- c("Project name", 
                         "Site where sample collected", 
                         "Plot where sample collected (SW = Saltwater, FW = Freshwater, C = 
                         Control, Estuary (Sourcewater))", 
                         "Grid where sample collected (Letter + Number)", 
                         "Depth where sample collected", 
                         "Event sampling time (T1 through T6)", 
                         "Time of day of sample collection (AM, PM, Midday)",
                         "Sample Identifier", 
                         "Volume of sample collected (mL)", 
                         "Date lysimeter evacuated", 
                         "Date lysimeter sample collected", 
                         "Time lysimeter evacuated", 
                         "Time sample collected", 
                         "Time zone of sample collection", 
                         "Date sourcewater sample collector deployed", 
                         "Time sourcewater sample collector deployed", 
                         "Time sourcewater sample collected", 
                         "NOx concentration in mg/L", 
                         "NOx concentration in uM", 
                         "NOx concentration flag base on standard curve (bdl=below detection limit (0), adl=above detection limit (highest   
                         standard), Within_Range=within detection range)", 
                         "NOx QAQC Flags generated during checks of standards, blanks, dups, and spks",
                         "NH3 concentration in mg/L", 
                         "NH3 concentration in uM", 
                         "NH3 concentration flag base on standard curve (bdl=below detection limit (0), adl=above detection limit (highest   
                         standard), Within_Range=within detection range)", 
                         "NH3 QAQC Flags generated during checks of standards, blanks, dups, and spks",
                         "PO4 concentration in mg/L", 
                         "PO4 concentration in uM", 
                         "PO4 concentration flag base on standard curve (bdl=below detection limit (0), adl=above detection limit (highest   
                         standard), Within_Range=within detection range)", 
                         "PO4 QAQC Flags generated during checks of standards, blanks, dups, and spks",
                         "The date that samples were run on the Nutrient IC", 
                         "Any notes generated during QAQC", 
                         "Notes generated during field sample collection")

metadat$Units <-       c("NA", 
                         "NA", 
                         "NA",
                         "NA",
                         "NA", 
                         "NA",
                         "NA",
                         "NA", 
                         "mL", 
                         "NA",
                         "NA",
                         "NA",
                         "NA",
                         "NA",
                         "NA",
                         "NA",
                         "NA", 
                         "mg/L", 
                         "uM", 
                         "NA", 
                         "NA", 
                         "mg/L", 
                         "uM", 
                         "NA", 
                         "NA", 
                         "mg/L", 
                         "uM", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA",
                         "NA")

metadat$Instrument_ifapplicable <-  
  
                       c("NA", 
                         "NA", 
                         "NA",
                         "NA",
                         "NA", 
                         "NA",
                         "NA",
                         "NA", 
                         "mL", 
                         "NA",
                         "NA",
                         "NA",
                         "NA",
                         "NA",
                         "NA",
                         "NA",
                         "NA", 
                         "AQ300 Discrete Analyzer (SEAL)", 
                         "AQ300 Discrete Analyzer (SEAL)", 
                         "NA", 
                         "NA", 
                         "AQ300 Discrete Analyzer (SEAL)", 
                         "AQ300 Discrete Analyzer (SEAL)", 
                         "NA", 
                         "NA", 
                         "AQ300 Discrete Analyzer (SEAL)", 
                         "AQ300 Discrete Analyzer (SEAL)", 
                         "NA", 
                         "NA", 
                         "NA",
                         "NA", 
                         "NA")


```

## Write out files
```{r write out combined data and metadata, include=FALSE}

#write out a csv of all the data to the main folder: 
write.csv(final_dat, "COMPASS_TEMPEST_Nutrient_AllData.csv", row.names = FALSE)


#write out a csv of the metadata associated with the data: 
write.csv(metadat, "COMPASS_TEMPEST_Nutrient_Metadata.csv", row.names = FALSE)

```

## Visualize Data by Plot   
```{r Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=20, dpi = 300}

##Remove Sourcewater
data_plotting <- all_dat %>%
  filter(!str_detect(Grid, "SOURCEWATER"), ignore.case = TRUE)%>%
  filter(!is.na(Grid))

data_plotting <- data_plotting %>%
  mutate(
    NOx_Conc_mgL = as.numeric(NOx_Conc_mgL),
    NH3_Conc_mgL = as.numeric(NH3_Conc_mgL),
    PO4_Conc_mgL = as.numeric(PO4_Conc_mgL)
  )

data_plotting <- data_plotting %>% 
    mutate(
    plot_1 = case_when(
      Plot == "C" ~ "Control",
      Plot == "FW" ~ "Freshwater",
      Plot == "SW" ~ "Saltwater",
      TRUE ~ Plot  # keep everything else unchanged
    )
  ) %>%
  mutate(plot_1 = factor(plot_1, levels = c("Control", "Freshwater", "Saltwater")))

#Plot samples to get a first look at concentrations (sanity check)
viz_NOx_plot <- ggplot(data_plotting, aes(x = as.character(Year), y = NOx_Conc_mgL, fill = plot_1)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ plot_1, scales="free") +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = " ", y = "NOx (mg/L)", title = "TEMPEST Porewater: NOx") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))+ 
  theme(legend.position = "none")

# viz_NOx_plot


viz_NH3_plot <- ggplot(data_plotting, aes(x = as.character(Year), y = NH3_Conc_mgL, fill = plot_1)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ plot_1, scales="free") +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = " ", y = "NH3 (mg/L)", title = "TEMPEST Porewater: NH3") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))+ 
  theme(legend.position = "none")

# viz_NH3_plot

viz_PO4_plot <- ggplot(data_plotting, aes(x = as.character(Year), y = PO4_Conc_mgL, fill = plot_1)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ plot_1, scales="free") +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = " ", y = "PO4 (mg/L)", title = "TEMPEST Porewater: PO4") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))+ 
  theme(legend.position = "none")

# viz_PO4_plot

ggarrange(viz_NOx_plot, viz_NH3_plot, viz_PO4_plot, nrow=3, ncol=1)


```


## Summarized data for Site and Zone 
```{r Summarize and Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=22, dpi = 300}

##Summarize SO4 and Cl by Date and Plot
Nutrient_avg <- data_plotting %>%
  group_by(Collection_Date_formatted, plot_1) %>%
  dplyr::summarize(mean_NOx = mean(NOx_Conc_mgL, na.rm=TRUE), 
            sd_NOx = sd(NOx_Conc_mgL, na.rm=TRUE), 
            mean_NH3 = mean(NH3_Conc_mgL, na.rm=TRUE), 
            sd_NH3 = sd(NH3_Conc_mgL, na.rm=TRUE), 
            mean_PO4 = mean(PO4_Conc_mgL, na.rm=TRUE), 
            sd_PO4 = sd(PO4_Conc_mgL, na.rm=TRUE))

# Define the specific date breaks you want to display
specific_breaks <- as.Date(c("2023-01-01", "2024-01-01", "2025-01-01"))

viz_NOx_plot_avgs <- ggplot(Nutrient_avg, aes(x = Collection_Date_formatted, y = mean_NOx, color = plot_1)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_NOx - sd_NOx, ymax = mean_NOx + sd_NOx), width = 0.2) + 
  geom_line() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) + 
  theme_classic() +
  labs(x = "Date", y = "NOx (mg/L)", title = "Samples: NOx") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") + # Format x-axis labels
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# viz_NOx_plot_avgs

viz_NH3_plot_avgs <- ggplot(Nutrient_avg, aes(x = Collection_Date_formatted, y = mean_NH3, color = plot_1)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_NH3 - sd_NH3, ymax = mean_NH3 + sd_NH3), width = 0.2) + 
  geom_line() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) + 
  theme_classic() +
  labs(x = "Date", y = "NH3 (mg/L)", title = "Samples: NH3") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") + # Format x-axis labels
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# viz_NH3_plot_avgs

viz_PO4_plot_avgs <- ggplot(Nutrient_avg, aes(x = Collection_Date_formatted, y = mean_PO4, color = plot_1)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_PO4 - sd_PO4, ymax = mean_PO4 + sd_PO4), width = 0.2) + 
  geom_line() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) + 
  theme_classic() +
  labs(x = "Date", y = "PO4 (mg/L)", title = "Samples: PO4") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") + # Format x-axis labels
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# viz_PO4_plot_avgs

ggarrange(viz_NOx_plot_avgs, viz_NH3_plot_avgs, viz_PO4_plot_avgs, nrow=3, ncol=1)

```

## Boxplot summary
```{r Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Summarize by Date, Site, and Zone
viz_NOx_boxplot <- ggplot(data_plotting, aes(x = Collection_Date, y = NOx_Conc_mgL, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) +
  theme_classic() +
  labs(x = " ", y = "NOx (mg/L)", title = "TEMPEST Porewater: NOx") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_NOx_boxplot


viz_NH3_boxplot <- ggplot(data_plotting, aes(x = Collection_Date, y = NH3_Conc_mgL, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) +
  theme_classic() +
  labs(x = " ", y = "NH3 (mg/L)", title = "TEMPEST Porewater: NH3") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")


viz_NH3_boxplot

viz_PO4_boxplot <- ggplot(data_plotting, aes(x = Collection_Date, y = PO4_Conc_mgL, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) +
  theme_classic() +
  labs(x = " ", y = "PO4 (mg/L)", title = "TEMPEST Porewater: PO4") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_PO4_boxplot

# ggarrange(viz_cl_plot, viz_so4_plot, nrow=2, ncol=1)

# # Save as a PNG
# png("COMPASS_TEMPEST_NOx_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_NOx_boxplot)  # For ggplot objects, use print()
# dev.off() 
# 
# # Save as a PNG
# png("COMPASS_TEMPEST_NH3_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_NH3_boxplot)  # For ggplot objects, use print()
# dev.off() 
# 
# # Save as a PNG
# png("COMPASS_TEMPEST_PO4_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_PO4_boxplot)  # For ggplot objects, use print()
# dev.off() 

```

## Boxplot summary for Sourcewater
```{r Sourcewater Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Only Sourcewater
data_plotting_source <- all_dat %>%
  filter(str_detect(Grid, "SOURCEWATER")) %>%
  mutate(
    NOx_Conc_mgL = as.numeric(NOx_Conc_mgL),
    NH3_Conc_mgL = as.numeric(NH3_Conc_mgL),
    PO4_Conc_mgL = as.numeric(PO4_Conc_mgL)
  )

data_plotting_source <- data_plotting_source %>% 
    mutate(
    plot_1 = case_when(
      Plot == "FW" ~ "Freshwater",
      Plot == "SW" ~ "Saltwater",
      Plot == "ESTUARY" ~ "Estuary",
      TRUE ~ Plot  # keep everything else unchanged
    )
  ) %>%
  mutate(plot_1 = factor(plot_1, levels = c("Control", "Freshwater", "Saltwater", "Estuary")))


##Boxplot Summarize SO4 and Cl by Date, Site, and Plot
viz_NOx_boxplot <- ggplot(data_plotting_source, aes(x = Collection_Date,  y = NOx_Conc_mgL, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  # scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2", 
    "Estuary" = "darkred"
  ), labels = c("Control", "Freshwater", "Saltwater", "Estuary")) +
  theme_classic() +
  labs(x = " ", y = "NOx (mg/L)", title = "TEMPEST Sourcewater: NOx") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_NOx_boxplot


viz_NH3_boxplot <- ggplot(data_plotting_source, aes(x = Collection_Date,  y = NH3_Conc_mgL, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  # scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2", 
    "Estuary" = "darkred"
  ), labels = c("Control", "Freshwater", "Saltwater", "Estuary")) +
  theme_classic() +
  labs(x = " ", y = "NH3 (mg/L)", title = "TEMPEST Sourcewater: NH3") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_NH3_boxplot


viz_PO4_boxplot <- ggplot(data_plotting_source, aes(x = Collection_Date,  y = PO4_Conc_mgL, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  # scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2", 
    "Estuary" = "darkred"
  ), labels = c("Control", "Freshwater", "Saltwater", "Estuary")) +
  theme_classic() +
  labs(x = " ", y = "PO4 (mg/L)", title = "TEMPEST Sourcewater: PO4") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_PO4_boxplot


# ggarrange(viz_cl_boxplot, viz_so4_boxplot, nrow=1, ncol=2)

# # Save as a PNG
# png("COMPASS_SynopticCB_Cl_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_cl_boxplot)  # For ggplot objects, use print()
# dev.off() 
# 
# # Save as a PNG
# png("COMPASS_SynopticCB_SO4_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_so4_boxplot)  # For ggplot objects, use print()
# dev.off() 

```

## Calculate summary metrics
```{r Additional summary metrics, include=FALSE, dpi = 300}
# It would also be great to add to our collation codes a summary chunk that will calculate value ranges, means, and standard deviations (and errors) for each site, zone, depth for each year; also for each site zone, depth overall; and for each site and zone overall so that we have some easy access to metrics. Was just thinking this might be helpful today when re-collating the DIC data.

colnames(all_dat)

##Metrics for each site, zone, depth for each year
summary_by_year <- data_plotting %>%
  group_by(Year, plot_1) %>%
   reframe(
     min_NOx_mgL = min(NOx_Conc_mgL, na.rm=TRUE),
     max_NOx_mgL = max(NOx_Conc_mgL, na.rm=TRUE), 
     mean_NOx_mgL = mean(NOx_Conc_mgL, na.rm=TRUE), 
     sd_NOx_mgL = sd(NOx_Conc_mgL, na.rm=TRUE), 

     min_NH3_mgL = min(NH3_Conc_mgL, na.rm=TRUE),
     max_NH3_mgL = max(NH3_Conc_mgL, na.rm=TRUE), 
     mean_NH3_mgL = mean(NH3_Conc_mgL, na.rm=TRUE), 
     sd_NH3_mgL = sd(NH3_Conc_mgL, na.rm=TRUE), 
     
     min_PO4_mgL = min(PO4_Conc_mgL, na.rm=TRUE),
     max_PO4_mgL = max(PO4_Conc_mgL, na.rm=TRUE), 
     mean_PO4_mgL = mean(PO4_Conc_mgL, na.rm=TRUE), 
     sd_PO4_mgL = sd(PO4_Conc_mgL, na.rm=TRUE))

##Metrics for each site, zone, depth for all years
summary_all_years <- data_plotting %>%
  group_by(plot_1) %>%
   reframe(
     min_NOx_mgL = min(NOx_Conc_mgL, na.rm=TRUE),
     max_NOx_mgL = max(NOx_Conc_mgL, na.rm=TRUE), 
     mean_NOx_mgL = mean(NOx_Conc_mgL, na.rm=TRUE), 
     sd_NOx_mgL = sd(NOx_Conc_mgL, na.rm=TRUE), 

     min_NH3_mgL = min(NH3_Conc_mgL, na.rm=TRUE),
     max_NH3_mgL = max(NH3_Conc_mgL, na.rm=TRUE), 
     mean_NH3_mgL = mean(NH3_Conc_mgL, na.rm=TRUE), 
     sd_NH3_mgL = sd(NH3_Conc_mgL, na.rm=TRUE), 
     
     min_PO4_mgL = min(PO4_Conc_mgL, na.rm=TRUE),
     max_PO4_mgL = max(PO4_Conc_mgL, na.rm=TRUE), 
     mean_PO4_mgL = mean(PO4_Conc_mgL, na.rm=TRUE), 
     sd_PO4_mgL = sd(PO4_Conc_mgL, na.rm=TRUE))

```

##Calculate summary metrics for Sourcewater
```{r Sourcewater additional summary metrics, include=FALSE, dpi = 300}
# It would also be great to add to our collation codes a summary chunk that will calculate value ranges, means, and standard deviations (and errors) for each site, Grid, depth for each year; also for each site Grid, depth overall; and for each site and Grid overall so that we have some easy access to metrics. Was just thinking this might be helpful today when re-collating the DIC data.

colnames(data_plotting_source)

##Metrics for each Grid for each year
summary_by_year_source <- data_plotting_source %>%
  group_by(Year, plot_1) %>%
   reframe(
     min_NOx_mgL = min(NOx_Conc_mgL, na.rm=TRUE),
     max_NOx_mgL = max(NOx_Conc_mgL, na.rm=TRUE), 
     mean_NOx_mgL = mean(NOx_Conc_mgL, na.rm=TRUE), 
     sd_NOx_mgL = sd(NOx_Conc_mgL, na.rm=TRUE), 

     min_NH3_mgL = min(NH3_Conc_mgL, na.rm=TRUE),
     max_NH3_mgL = max(NH3_Conc_mgL, na.rm=TRUE), 
     mean_NH3_mgL = mean(NH3_Conc_mgL, na.rm=TRUE), 
     sd_NH3_mgL = sd(NH3_Conc_mgL, na.rm=TRUE), 
     
     min_PO4_mgL = min(PO4_Conc_mgL, na.rm=TRUE),
     max_PO4_mgL = max(PO4_Conc_mgL, na.rm=TRUE), 
     mean_PO4_mgL = mean(PO4_Conc_mgL, na.rm=TRUE), 
     sd_PO4_mgL = sd(PO4_Conc_mgL, na.rm=TRUE)) %>% 
  mutate(across(where(is.numeric), round, digits = 4)) 

```

## Summary Tables for Porewater
```{r Porewater summary table metrics, echo = FALSE, message = FALSE, dpi = 100}

library(kableExtra)

kable(
  summary_by_year[, c("Year", "plot_1", "min_NOx_mgL",	"max_NOx_mgL",	"mean_NOx_mgL", "min_NH3_mgL",	"max_NH3_mgL",	"mean_NH3_mgL", "min_PO4_mgL",	"max_PO4_mgL",	"mean_PO4_mgL")],  
  format = "latex",
  booktabs = TRUE,
  caption = "Porewater Summary Statistics by Year"
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)

kable(
  summary_all_years[, c("plot_1", "min_NOx_mgL",	"max_NOx_mgL",	"mean_NOx_mgL", "min_NH3_mgL",	"max_NH3_mgL",	"mean_NH3_mgL", "min_PO4_mgL",	"max_PO4_mgL",	"mean_PO4_mgL")],  
  format = "latex",
  booktabs = TRUE,
  caption = "Porewater Summary Statistics by Plot"
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)

```

## Summary Tables for Sourcewater
```{r Sourcewater summary table metrics, echo = FALSE, message = FALSE, dpi = 100}

kable(
  summary_by_year_source[, c("Year", "plot_1", "min_NOx_mgL",	"max_NOx_mgL",	"mean_NOx_mgL", "min_NH3_mgL",	"max_NH3_mgL",	"mean_NH3_mgL", "min_PO4_mgL",	"max_PO4_mgL",	"mean_PO4_mgL")],  
  format = "latex",
  booktabs = TRUE,
  caption = "Sourcewater Summary Statistics by Year"
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)

```

##Calculate number of samples 
```{r Number of samples, include=FALSE, dpi = 300}
#  I am thinking it might be cool to add to our porewater collation files a table at the end that has total number of samples then also some info about # of samples per zone, site, depth etc.

##Metrics for each year
sample_numbers_years <- data_plotting %>%
  group_by(Year) %>%
   reframe(
     Number_of_samples = n())

##Metrics for each site, zone, depth for all years
sample_numbers <- data_plotting %>%
  group_by(plot_1) %>%
   reframe(
     Number_of_samples = n())
     

```

## Number of Samples
```{r number of samples tables, echo = FALSE, message = FALSE, dpi = 100}

kable(
  sample_numbers_years[, c("Year", "Number_of_samples")],  
  format = "latex",
  booktabs = TRUE,
  caption = paste0("Total Samples: ", nrow(all_dat))
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)

```

```{r Number of Samples Barplots, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}
##Remove SW samples
# sample_numbers_pw <- sample_numbers %>% 
#   filter(Zone != "SW")

##Number of Samples Plots
sample_num_barplot <- ggplot(sample_numbers, aes(x = plot_1, y = Number_of_samples, fill = plot_1)) +
  geom_col(position = "dodge", color = "black", width = 0.75) + # 'color' adds a border to the bars
  # facet_grid(~plot_1 , scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) + 
  theme_classic() +
  labs(x = "Depth (cm)", y = "Number of Samples", title = "TEMPEST Porewater Sample Numbers") +
  theme(
    panel.spacing = unit(0.5, "lines"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+
  # theme(legend.position = "none")

sample_num_barplot
```

## Write out summarized data
```{r write out summarized data, include=FALSE}

#write out a csv of all the data to the main folder: 
write.csv(summary_by_year, "COMPASS_TEMPEST_Nutrient_Summarized_By_Year.csv")


#write out a csv of the metadata associated with the data: 
write.csv(summary_all_years, "COMPASS_TEMPEST_Nutrient_Summarized_Years_Combined.csv")

```



