---
title: "TMP_PW_Inventory_TrialCode"
author: "S.Wilson"
date: "2026-06-09"
output: pdf_document
---

Set paths for the datafiles you are planning to add to the inventory: 
```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(openxlsx)
library(googledrive)
library(googlesheets4)


#TEMPEST Event Path for just entered data sheets:
pw_drive_path <- "https://drive.google.com/drive/u/0/folders/1rgLRWzPJWol3McRebYK6zM3DwcHsnH5Z"

sce_drive_path <- "https://drive.google.com/drive/u/0/folders/1YjeqzmBnWLg8U_-fbNzzQnL_la-TGfSb"

ro_drive_path <- "https://drive.google.com/drive/u/0/folders/1gzAa97H8ukTfbjioSMfGeZWcsD7LwIX1"


#Full Inventory Path 
inventory_path <- "https://docs.google.com/spreadsheets/d/1sFWq-WKhemPzbOFInqhCu_Lx0lsO6a_Z/edit?gid=1740738201#gid=1740738201"

#pH file path 
pH_path <- "https://docs.google.com/spreadsheets/d/1KoskjXkHgPmSyFERvZLCX5ZBuV2cyev4-UEBuT0y-Jw/edit?gid=0#gid=0"

#Folder to upload the new files to
folder_path = as_id("1oEeiU48OULo0UOu0np3l1JhYYEQcnw7e")

```


PARSE ENTERED DATA FILES: RUN POREWATER CHUNKS, SOURCEWATER & RUNOFF FOR EVENTS  

Porewater
## Read in the POREWATER datasheets and parse
```{r, pull in entered datasheets}

#for just one folder 
#googledrive::drive_ls(path = drive_path, pattern = "Field") -> files

#to look within multipel folders 
files <- googledrive::drive_ls(
  path = pw_drive_path,
  recursive = TRUE,
  pattern = "Field"
)


source("parse_event_porewater_file.R")

parse_files <- list()

for(i in seq_len(nrow(files))) {

    cat("Processing:", files$name[i], "\n")

    drive_download(files$id[i],
                   path = "temp.xlsx",
                   overwrite = TRUE)

    pw_raw <- read_excel("temp.xlsx",
                         col_names = FALSE)

    parse_files[[i]] <- tryCatch(
        parse_event_porewater_file(pw_raw),
        error = function(e) {
            message("Failed on: ", files$name[i])
            message(e$message)
            return(NULL)
        }
    )

    unlink("temp.xlsx")
}

# I think we can use either purrr:map or lapply here but I couldn't figure it out, 
# for now we'll use a for loop

parse_files <- list()

for(i in 1:nrow(files)) {
    
    drive_download(files$id[i], path = "temp.xlsx", overwrite = TRUE)
    read_excel("temp.xlsx", col_names = FALSE) -> pw_raw
    parse_event_porewater_file(pw_raw) -> parse_files[[i]]
    unlink("temp.xlsx")

} 

# When you're ready to upload you can use googledrive::drive_upload

```

# Reorganize POREWATER data to match inventory
```{r}

source("reorg_event_porewater_file.R")

pw_inventory_files <- lapply(parse_files, process_event_inventory)

pw_inventory_all <- bind_rows(pw_inventory_files)


```

Surface Water
## Read in the SOURCEWATER datasheets and parse 
```{r}

#to look within multiple folders 
sce_files <- googledrive::drive_ls(
  path = sce_drive_path,
  recursive = TRUE,
  pattern = "Field"
)

source("parse_sourcewater_file.R")

parse_sce_files <- list()

for(i in seq_len(nrow(sce_files))) {

  cat("Processing:", sce_files$name[i], "\n")

  drive_download(
    sce_files$id[i],
    path = "sce_temp.xlsx",
    overwrite = TRUE
  )

  parse_sce_files[[i]] <- tryCatch(
    parse_sourcewater_file("sce_temp.xlsx"),
    error = function(e) {
      message("Failed on: ", sce_files$name[i])
      message(e$message)
      NULL
    }
  )

  unlink("sce_temp.xlsx")
}


```

# Reorganize SOURCEWATER data to match inventory
```{r}

source("reorg_sourcewater_file.R")

sce_inventory_files <- lapply(parse_sce_files, process_sce_inventory)

sce_inventory_all <- bind_rows(sce_inventory_files)


```


# Reorganize POREWATER and SOURCEWATER pH data to pH entering datasheets. 
Runoff pH was measured by the Aquatroll. 
```{r}

#Porewater
source("reorg_pH_file.R")
pw_pH_files <- lapply(parse_files, process_event_pw_pH)
pw_pH_all <- bind_rows(pw_pH_files)

#Sourcewater
source("reorg_pH_file.R")
sce_pH_files <- lapply(parse_sce_files, process_sce_pH)
sce_pH_all <- bind_rows(sce_pH_files)

pH_all <- bind_rows(pw_pH_all, sce_pH_all)

```

NOW COMBINE THE NEW pH DATA WITH THE OLD
# Prep pH to be appended to the pH file 
```{r, process data file to match inventory}

# Run this to pull the TEMPEST porewater metadata into GitHub if not already there
drive_download(pH_path, overwrite = TRUE)

#Read in the master inventory
pH_master <- read_excel(
  "TEMPEST_Porewater_pH.xlsx",
  sheet = "pH"
)

pH_master <- pH_master  %>%
    mutate(
      Depth_cm = as.numeric(Depth_cm), 
      Cond = as.numeric(Cond),
      Temp_C = as.numeric(Temp_C)
    )

# Align the columns
pH_new <- pH_all %>%
  select(any_of(names(pH_master)))

missing_cols <- setdiff(names(pH_master), names(pH_new))
missing_cols

for (col in missing_cols) {
  pH_new[[col]] <- NA
}

# Append to the inventory
pH_new <- pH_new %>%
  select(names(pH_master))

pH_final <- bind_rows(pH_new, pH_master)

```

# Replace the POREWATER sheet in the inventory file - not doing this because not pushing to google yet 
```{r, add the new inventory file}

# Load existing workbook
wb <- loadWorkbook("TEMPEST_Porewater_pH.xlsx")

# Remove existing sheet if it exists
if ("pH Updated" %in% names(wb)) {
  removeWorksheet(wb, "pH Updated")
}

# Add replacement sheet
addWorksheet(wb, "pH Updated")

# Write inventory data below it
header_style <- createStyle(
  textDecoration = "bold"
)

writeData(
  wb,
  "pH Updated",
  pH_final, 
  headerStyle = header_style
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_Porewater_pH_updated.xlsx", overwrite = TRUE)

drive_upload(
  media = "TEMPEST_Porewater_pH_updated.xlsx",
  path = folder_path, 
  name = "TEMPEST_Porewater_pH_updated",
  type = "spreadsheet"                    # Triggers conversion to Google Sheets
)

```

Runoff
## Read in the RUNOFF datasheets and parse 
```{r}

#to look within multipel folders 
ro_files <- googledrive::drive_ls(
  path = ro_drive_path,
  recursive = TRUE,
  pattern = "Field"
)

source("parse_runoff_file.R")

parse_ro_files <- list()

for(i in seq_len(nrow(ro_files))) {

  cat("Processing:", ro_files$name[i], "\n")

  drive_download(
    ro_files$id[i],
    path = "ro_temp.xlsx",
    overwrite = TRUE
  )

  parse_ro_files[[i]] <- tryCatch(
    parse_runoff_file("ro_temp.xlsx"),
    error = function(e) {
      message("Failed on: ", ro_files$name[i])
      message(e$message)
      NULL
    }
  )

  unlink("ro_temp.xlsx")
}


```
# Reorganize RUNOFF data to match inventory
```{r}

source("reorg_runoff_file.R")

ro_inventory_files <- lapply(parse_ro_files, process_ro_inventory)

ro_inventory_all <- bind_rows(ro_inventory_files)


```



NOW COMBINE THE NEW INVENTORY WITH THE OLD
Porewater
# Prep POREWATER to be appended to the inventory file 
```{r, process data file to match inventory}

# Run this to pull the TEMPEST porewater metadata into GitHub if not already there
drive_download(inventory_path, overwrite = TRUE)

#Read in the master inventory
master <- read_excel(
  "TEMPEST_PorewaterInventory_May2022_Present.xlsx",
  sheet = "Porewater - Individual",
  skip = 3
)

#clean out any empty sample_ID rows
master <- master %>%
  filter(!is.na(Sample_ID))

# Align the columns
inventory_new <- pw_inventory_all %>%
  select(any_of(names(master)))

missing_cols <- setdiff(names(master), names(inventory_new))
missing_cols

for (col in missing_cols) {
  inventory_new[[col]] <- NA
}

# Append to the inventory
inventory_new <- inventory_new %>%
  select(names(master))

pw_inventory <- bind_rows(inventory_new, master)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS: TEMPEST",
    "Site: SERC",
    "Content: Lysimeter Porewater Inventory",
    "" ))

```

# Replace the POREWATER sheet in the inventory file - not doing this because not pushing to google yet 
```{r, add the new inventory file}

# Run this to pull the TEMPEST porewater metadata into GitHub if not already there
drive_download(inventory_path, overwrite = TRUE)

# Load existing workbook
wb <- loadWorkbook("TEMPEST_PorewaterInventory_May2022_Present.xlsx")

# Remove existing sheet if it exists
if ("Porewater - Individual - Updated" %in% names(wb)) {
  removeWorksheet(wb, "Porewater - Individual Updated")
}

# Add replacement sheet
addWorksheet(wb, "Porewater - Individual Updated")

# Write header block
writeData(wb, "Porewater - Individual Updated", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "Porewater - Individual Updated",
  pw_inventory,
  startRow = nrow(header_rows) + 1
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", overwrite = TRUE)

```

Source Water 
# Prep SOURCEWATER to be appended to the inventory file  
```{r, process data file to match inventory}

#Read in the master inventory 
sce_master <- read_excel(
  #path = inventory_path,
  "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
  sheet = "Source Water",
  skip = 4
)

#clean out any empty sample_ID rows
sce_master <- sce_master %>%
  filter(!is.na(Sample_ID))

# Align the columns
sce_inventory_new <- sce_inventory_all %>%
  select(any_of(names(sce_master)))

missing_cols <- setdiff(names(sce_master), names(sce_inventory_new))
missing_cols

for (col in missing_cols) {
  sce_inventory_new[[col]] <- NA
}

# Append to the inventory
sce_inventory_new <- sce_inventory_new %>%
  select(names(sce_master))

sce_inventory <- bind_rows(sce_inventory_new, sce_master)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS: TEMPEST",
    "Site: SERC",
    "Event: Source Water",
    "" ))

```

# Replace the SOURCEWATER sheet in the inventory file 
```{r, add the new inventory file}

# Load existing workbook
wb <- loadWorkbook("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx")

# Remove existing sheet if it exists
if ("Source Water Updated" %in% names(wb)) {
  removeWorksheet(wb, "Source Water Updated")
}

# Add replacement sheet
addWorksheet(wb, "Source Water Updated")

# Write header block
writeData(wb, "Source Water Updated", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "Source Water Updated",
  sce_inventory,
  startRow = nrow(header_rows) + 1
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", overwrite = TRUE)

```


Runoff
# Prep RUNOFF to be appended to the inventory file  
There is currently no runoff in the original inventory file, so commented out some parts of this section for now
```{r, process data file to match inventory}

#Read in the master inventory 
# ro_master <- read_excel(
#   #path = inventory_path,
#   "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
#   sheet = "Source Water",
#   skip = 4
# )
# 
# #clean out any empty sample_ID rows
# ro_master <- ro_master %>%
#   filter(!is.na(Sample_ID))

# Align the columns
# ro_inventory_new <- ro_inventory_all %>%
#   select(any_of(names(sce_master)))

# missing_cols <- setdiff(names(ro_master), names(ro_inventory_new))
# missing_cols
# 
# for (col in missing_cols) {
#   ro_inventory_new[[col]] <- NA
# }

# ro_inventory_new <- ro_inventory_new %>%
#   select(names(ro_master))

# Append to the inventory
# ro_inventory <- bind_rows(ro_inventory_new, ro_master)

ro_inventory <- ro_inventory_all %>% 
  select(-analyte_code)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS: TEMPEST",
    "Site: SERC",
    "Event: Runoff",
    "" ))

```

# Replace the Runoff sheet in the inventory file 
```{r, add the new inventory file}

# Load existing workbook
wb <- loadWorkbook("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx")

# Remove existing sheet if it exists
if ("Runoff Updated" %in% names(wb)) {
  removeWorksheet(wb, "Runoff Updated")
}

# Add replacement sheet
addWorksheet(wb, "Runoff Updated")

# Write header block
writeData(wb, "Runoff Updated", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "Runoff Updated",
  ro_inventory,
  startRow = nrow(header_rows) + 1
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", overwrite = TRUE)

```

-Need to do the same thing for pooled SPE samples and remove those from the individual porewater sample list, also need to output the pooled samples with the volumes from each grid somehow 

# Upload the new inventory to Google Drive
```{r}

drive_upload(
  media = "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
  path = folder_path, 
  name = "TEMPEST_PorewaterInventory_May2022_Present_updated",
  type = "spreadsheet"                    # Triggers conversion to Google Sheets
)

```


# Create sample lists from the new inventory or whole inventory

```{r, create sample lists}

#Input the start date, end date, and the analyte you want the list for 
start_date <- as.numeric("20260501")
end_date   <- as.numeric("20260612")
target_analyte <- "SO4/Cl/H2S" #"SO4/Cl/H2S"
target_file_name <- "SO4"


#create a function that builds a sample list 
build_sample_list <- function(dat, target_analyte, start_date, end_date) {

  dat %>%
    filter(
      Analyte == target_analyte,
      Collection_Date_YYYYMMDD >= start_date,
      Collection_Date_YYYYMMDD <= end_date
    ) %>%
    mutate(
      Plot = if ("Plot" %in% names(.)) {
        factor(Plot, levels = c("Control", "Saltwater", "Freshwater", "Estuary"))
      } else {
        NA_character_
      }
    ) %>%
    arrange(Collection_Date_YYYYMMDD, Timepoint, Plot) %>%
    transmute(
      Sample_ID,
      Plot,
      Collection_Date_YYYYMMDD,
      Timepoint,
      Analyte
    )
}

#apply the function to each inventory we made 
pw_sample_list <- build_sample_list(
  pw_inventory_all,
  target_analyte,
  start_date,
  end_date
)

sce_sample_list <- build_sample_list(
  sce_inventory_all,
  target_analyte,
  start_date,
  end_date
)

ro_sample_list <- build_sample_list(
  ro_inventory_all,
  target_analyte,
  start_date,
  end_date
)




```

```{r, combine and order sample lists}

#add a type column 
pw_sample_list$type <- "PW"
sce_sample_list$type <- "SCE"
ro_sample_list$type <- "RO"


#order the pw list 
pw_sample_list <- pw_sample_list %>%
  mutate(
    Grid = stringr::str_extract(Sample_ID, "(?<=_)[A-Z]\\d"),
    Depth_cm = stringr::str_extract(Sample_ID, "\\d+(?=cm)") %>% as.numeric()
  )

grid_order <- c("B4","C3","C6", "D5", "E3","F4","H3","H6", "I5")

pw_sample_list <- pw_sample_list %>%
  mutate(
    Grid = factor(Grid, levels = grid_order)
  )


#order the SCE list 
sce_sample_list <- sce_sample_list %>%
  mutate(
    time_order = paste(Collection_Date_YYYYMMDD, Timepoint)
  )

#order the ro list 
ro_sample_list <- ro_sample_list %>%
  mutate(
    time_order = paste(Collection_Date_YYYYMMDD, Timepoint)
  )

#combine the dataframes
all_samples <- bind_rows(
  pw_sample_list,
  sce_sample_list,
  ro_sample_list
)

all_samples <- all_samples %>%
  mutate(
    time_block = Timepoint
  )


all_samples <- all_samples %>%
  mutate(
    type_order = case_when(
      type == "PW" ~ 1,
      type == "SCE" ~ 2,
      type == "RO" ~ 3,
      TRUE ~ 99
    )
  ) %>%
  arrange(
    Collection_Date_YYYYMMDD,
    time_block,
    type_order,
    Plot,
    Grid,
    Depth_cm
  ) %>%
  mutate(
    Sample_Number = row_number()
  )

sample_list <- all_samples %>%
  select(
    Sample_Number,
    Sample_ID,
    Plot,
    Collection_Date_YYYYMMDD,
    Timepoint, 
    type
  )
  


```

```{r, read out sample list  }

#Export to print the list 
printable_list <- createWorkbook()
addWorksheet(printable_list, "Sample List")

#create a header 
writeData(
  printable_list,
  "Sample List",
  paste("TEMPEST Sample List:", target_analyte, start_date, "-", end_date),
  startRow = 1,
  startCol = 1,
  colNames = FALSE
)

writeData(
  printable_list,
  "Sample List",
  sample_list,
  startRow = 2)

#Save the workbook
file_name <- paste0(
  "TEMPEST_",
  target_file_name, "_",
  start_date, "-", end_date,
  ".xlsx")

#export the workbook 
saveWorkbook(printable_list, file_name, overwrite = TRUE)

```


