---
title: "TMP_PW_Inventory_MonMon_Code"
author: "S.Wilson, Z. Read"
date: "2026-06-09"
output: pdf_document
---

Set paths for the datafiles you are planning to add to the inventory: 
```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(openxlsx)
library(googledrive)
library(googlesheets4)

####CHANGE FOR THE MONTH AS NEEDED: ####

#Monthly path 
drive_path <- "https://drive.google.com/drive/folders/1mCUes-7sBt0PAWxQ5mPz5wn2E1w558aw"

#Full Inventory Path 
inventory_path <- "https://docs.google.com/spreadsheets/d/1WuuUiIieTfmYlZUuQkEBkTkx4bumNjNEGKBcxTJ2viw/edit?gid=834818575#gid=834818575"

#pH file path 
pH_path <- "https://docs.google.com/spreadsheets/d/1-KTABMU-2e6ZxQXeY1yMahHB5lX9HuvN-NF-kWinLTA/edit?gid=1955793989#gid=1955793989"

#Folder to upload the new files to
folder_path = as_id("1oEeiU48OULo0UOu0np3l1JhYYEQcnw7e")

```


PARSE ENTERED DATA FILES: RUN POREWATER CHUNKS

Porewater
## Read in the POREWATER datasheets and parse
```{r, pull in entered datasheets}

#for just one folder 
googledrive::drive_ls(path = drive_path, pattern = "Field") -> files

source("parse_monthly_porewater_file.R")

parse_files <- list()

for(i in seq_len(nrow(files))) {

    cat("Processing:", files$name[i], "\n")

    drive_download(files$id[i],
                   path = "temp.xlsx",
                   overwrite = TRUE)

    pw_raw <- read_excel("temp.xlsx",
                         col_names = FALSE, 
                         col_types = "text")

    parse_files[[i]] <- tryCatch(
        parse_monthly_porewater_file(pw_raw),
        error = function(e) {
            message("Failed on: ", files$name[i])
            message(e$message)
            return(NULL)
        }
    )

    unlink("temp.xlsx")
}

```

# Reorganize POREWATER data to match inventory
```{r}

source("reorg_monthly_porewater_file.R")

pw_inventory_files <- lapply(parse_files, process_monthly_inventory)

pw_inventory_all <- bind_rows(pw_inventory_files)

```

# Reorganize POOLED POREWATER data to match inventory
```{r}

source("reorg_monthly_pooled_porewater_file.R")

pooled_inventory_files <- lapply(parse_files, process_monthly_pooled_inventory)

pooled_inventory_all <- bind_rows(pooled_inventory_files)

```



# Reorganize pH data from POREWATER and SOURCEWATER 
Runoff pH was measured by the Aquatroll. 
```{r}

#Porewater
source("reorg_monthly_pH_file.R")
pw_pH_files <- lapply(parse_files, process_monthly_pw_pH)
pw_pH_all <- bind_rows(pw_pH_files)

pH_all <- pw_pH_all %>%
  mutate(Plot = factor(Plot, levels = c("Control", "Saltwater", "Freshwater"))) %>%
  arrange(Date_YYYYMMDD, Timepoint, Plot, Location, Depth_cm)

```

NOW COMBINE THE NEW pH DATA WITH THE OLD
# Prep pH to be appended to the pH file 
```{r, process data file to match inventory}

# Run this to pull the TEMPEST porewater metadata into GitHub if not already there
drive_download(pH_path, overwrite = TRUE)

#Read in the master inventory
pH_master <- read_excel(
  "TEMPEST_Porewater_pH.xlsx",
  sheet = "pH"
)

pH_master <- pH_master  %>%
  mutate(
      Depth_cm = as.numeric(Depth_cm), 
      Cond = as.numeric(Cond),
      Temp_C = as.numeric(Temp_C)
    ) %>% 
  filter(!if_all(c(pH, Cond, Temp_C), is.na))

# Align the columns
pH_new <- pH_all %>%
  select(any_of(names(pH_master)))

missing_cols <- setdiff(names(pH_master), names(pH_new))
missing_cols

for (col in missing_cols) {
  pH_new[[col]] <- NA
}

# Append to the inventory
pH_new <- pH_new %>%
  select(names(pH_master))

pH_final <- bind_rows(pH_master, pH_new)

```

# Replace the pH sheet in the pH file
```{r, add the new pH file}

# Load existing workbook
wb <- loadWorkbook("TEMPEST_Porewater_pH.xlsx")

# Remove existing sheet if it exists
if ("pH" %in% names(wb)) {
  removeWorksheet(wb, "pH")
}

# Add replacement sheet
addWorksheet(wb, "pH")

# Write inventory data below it
header_style <- createStyle(
  textDecoration = "bold"
)

writeData(
  wb,
  "pH",
  pH_final, 
  headerStyle = header_style
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_Porewater_pH_updated.xlsx", overwrite = TRUE)

```



NOW COMBINE THE NEW INVENTORY WITH THE OLD
Porewater
# Prep POREWATER to be appended to the inventory file 
```{r, process data file to match inventory}

# Run this to pull the TEMPEST porewater metadata into GitHub if not already there
drive_download(inventory_path, overwrite = TRUE)

#Read in the master inventory
master <- read_excel(
  "TEMPEST_PorewaterInventory_May2022_Present.xlsx",
  sheet = "Porewater - Individual",
  skip = 3
)

#clean out any empty sample_ID rows
master <- master %>%
  filter(!is.na(Sample_ID))

# Align the columns
inventory_new <- pw_inventory_all %>%
  select(any_of(names(master)))

missing_cols <- setdiff(names(master), names(inventory_new))
missing_cols

for (col in missing_cols) {
  inventory_new[[col]] <- NA
}

# Append to the inventory
inventory_new <- inventory_new %>%
  select(names(master))

pw_inventory <- bind_rows(inventory_new, master)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS",
    "Site: TEMPEST",
    "Content: Lysimeter Porewater Inventory",
    "" ))

```

# Replace the POREWATER sheet in the inventory file
```{r, add the new inventory file}

# Run this to pull the TEMPEST porewater metadata into GitHub if not already there
drive_download(inventory_path, overwrite = TRUE)

# Load existing workbook
wb <- loadWorkbook("TEMPEST_PorewaterInventory_May2022_Present.xlsx")

# Remove existing sheet if it exists
if ("Porewater - Individual" %in% names(wb)) {
  removeWorksheet(wb, "Porewater - Individual")
}

# Add replacement sheet
addWorksheet(wb, "Porewater - Individual")

# Write header block
writeData(wb, "Porewater - Individual", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "Porewater - Individual",
  pw_inventory,
  startRow = nrow(header_rows) + 1
)

##Bold the headers
header_style <- createStyle(
  textDecoration = "bold",
  halign = "center",
  valign = "center"
)
addStyle(
  wb,
  sheet = "Porewater - Individual",
  style = header_style,
  rows = 1:5,
  cols = 1:ncol(pw_inventory),
  gridExpand = TRUE
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", overwrite = TRUE)

```


POOLED
# Prep POOLED porewater to be appended to the inventory file 
```{r, process data file to match inventory}

#Read in the master inventory 
pooled_master <- read_excel(
  #path = inventory_path,
  "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
  sheet = "Porewater - Pooled",
  skip = 4
)

#clean out any empty sample_ID rows
pooled_master <- pooled_master %>%
  filter(!is.na(Sample_ID))

# Align the columns
inventory_new <- pooled_inventory_all %>%
  select(any_of(names(pooled_master)))

missing_cols <- setdiff(names(pooled_master), names(inventory_new))
missing_cols

for (col in missing_cols) {
  inventory_new[[col]] <- NA
}

# Append to the inventory
inventory_new <- inventory_new %>%
  select(names(pooled_master))

pooled_inventory <- bind_rows(inventory_new, pooled_master)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS",
    "Site: TEMPEST",
    "Event: POOLED Porewater Inventory July 2022 - present",
    "" ))

```

# Replace the POOLED porewater sheet in the inventory file
```{r, add the new inventory file}

# Load existing workbook
wb <- loadWorkbook("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx")

# Remove existing sheet if it exists
if ("Porewater - Pooled" %in% names(wb)) {
  removeWorksheet(wb, "Porewater - Pooled")
}

# Add replacement sheet
addWorksheet(wb, "Porewater - Pooled")

# Write header block
writeData(wb, "Porewater - Pooled", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "Porewater - Pooled",
  pooled_inventory,
  startRow = nrow(header_rows) + 1
)

# Write the header text into G4 (column 7, row 4)
writeData(
  wb,
  sheet = "Porewater - Pooled",
  x = "Breakdown_by_volume_mL",
  startCol = 7,
  startRow = 4,
  colNames = FALSE
)

# Merge G4:Q4 (columns 7-17)
mergeCells(
  wb,
  sheet = "Porewater - Pooled",
  cols = 7:16,
  rows = 4
)

##Bold the headers
header_style <- createStyle(
  textDecoration = "bold",
  halign = "center",
  valign = "center"
)
addStyle(
  wb,
  sheet = "Porewater - Pooled",
  style = header_style,
  rows = 1:5,
  cols = 1:21,
  gridExpand = TRUE
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", overwrite = TRUE)

```


# Upload the new pH and inventory files to Google Drive
```{#r}

##pH
drive_upload(
  media = "TEMPEST_Porewater_pH_updated.xlsx",
  path = folder_path, 
  name = "TEMPEST_Porewater_pH_updated",
  type = "spreadsheet"                    # Triggers conversion to Google Sheets
)

##inventory
drive_upload(
  media = "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
  path = folder_path, 
  name = "TEMPEST_PorewaterInventory_May2022_Present_updated",
  type = "spreadsheet"                    # Triggers conversion to Google Sheets
)

```


# Create sample lists from the new inventory or whole inventory

```{#r, create sample lists}

#Input the start date, end date, and the analyte you want the list for 
start_date <- as.numeric("20260501")
end_date   <- as.numeric("20260612")
target_analyte <- "SO4/Cl/H2S" #"SO4/Cl/H2S"
target_file_name <- "SO4"


#create a function that builds a sample list 
build_sample_list <- function(dat, target_analyte, start_date, end_date) {

  dat %>%
    filter(
      Analyte == target_analyte,
      Collection_Date_YYYYMMDD >= start_date,
      Collection_Date_YYYYMMDD <= end_date
    ) %>%
    mutate(
      Plot = if ("Plot" %in% names(.)) {
        factor(Plot, levels = c("Control", "Saltwater", "Freshwater", "Estuary"))
      } else {
        NA_character_
      }
    ) %>%
    arrange(Collection_Date_YYYYMMDD, Timepoint, Plot) %>%
    transmute(
      Sample_ID,
      Plot,
      Collection_Date_YYYYMMDD,
      Timepoint,
      Analyte
    )
}

#apply the function to each inventory we made 
pw_sample_list <- build_sample_list(
  pw_inventory_all,
  target_analyte,
  start_date,
  end_date
)

pooled_sample_list <- build_sample_list(
  pooled_inventory_all,
  target_analyte,
  start_date,
  end_date
)

sce_sample_list <- build_sample_list(
  sce_inventory_all,
  target_analyte,
  start_date,
  end_date
)

ro_sample_list <- build_sample_list(
  ro_inventory_all,
  target_analyte,
  start_date,
  end_date
)




```

```{#r, combine and order sample lists}

#add a type column 
pw_sample_list$type <- "PW"
sce_sample_list$type <- "SCE"
ro_sample_list$type <- "RO"


#order the pw list 
pw_sample_list <- pw_sample_list %>%
  mutate(
    Grid = stringr::str_extract(Sample_ID, "(?<=_)[A-Z]\\d"),
    Depth_cm = stringr::str_extract(Sample_ID, "\\d+(?=cm)") %>% as.numeric()
  )

grid_order <- c("B4","C3","C6", "D5", "E3","F4","H3","H6", "I5")

pw_sample_list <- pw_sample_list %>%
  mutate(
    Grid = factor(Grid, levels = grid_order)
  )


#order the SCE list 
sce_sample_list <- sce_sample_list %>%
  mutate(
    time_order = paste(Collection_Date_YYYYMMDD, Timepoint)
  )

#order the ro list 
ro_sample_list <- ro_sample_list %>%
  mutate(
    time_order = paste(Collection_Date_YYYYMMDD, Timepoint)
  )

#combine the dataframes
all_samples <- bind_rows(
  pw_sample_list,
  sce_sample_list,
  ro_sample_list
)

all_samples <- all_samples %>%
  mutate(
    time_block = Timepoint
  )


all_samples <- all_samples %>%
  mutate(
    type_order = case_when(
      type == "PW" ~ 1,
      type == "SCE" ~ 2,
      type == "RO" ~ 3,
      TRUE ~ 99
    )
  ) %>%
  arrange(
    Collection_Date_YYYYMMDD,
    time_block,
    type_order,
    Plot,
    Grid,
    Depth_cm
  ) %>%
  mutate(
    Sample_Number = row_number()
  )

sample_list <- all_samples %>%
  select(
    Sample_Number,
    Sample_ID,
    Plot,
    Collection_Date_YYYYMMDD,
    Timepoint, 
    type
  )
  


```

```{#r, read out sample list  }

#Export to print the list 
printable_list <- createWorkbook()
addWorksheet(printable_list, "Sample List")

#create a header 
writeData(
  printable_list,
  "Sample List",
  paste("TEMPEST Sample List:", target_analyte, start_date, "-", end_date),
  startRow = 1,
  startCol = 1,
  colNames = FALSE
)

writeData(
  printable_list,
  "Sample List",
  sample_list,
  startRow = 2)

#Save the workbook
file_name <- paste0(
  "TEMPEST_",
  target_file_name, "_",
  start_date, "-", end_date,
  ".xlsx")

#export the workbook 
saveWorkbook(printable_list, file_name, overwrite = TRUE)

```


