---
title: "TMP_PW_Inventory_TrialCode"
author: "S.Wilson"
date: "2026-06-09"
output: pdf_document
---

Set paths for the datafiles you are planning to add to the inventory: 
```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(openxlsx)
library(googledrive)
library(googlesheets4)



#TEMPEST Event Path for just entered data sheets:
pw_drive_path <- "https://drive.google.com/drive/u/0/folders/1rgLRWzPJWol3McRebYK6zM3DwcHsnH5Z"

sce_drive_path <- "https://drive.google.com/drive/u/0/folders/1YjeqzmBnWLg8U_-fbNzzQnL_la-TGfSb"

ro_drive_path <- "https://drive.google.com/drive/u/0/folders/1gzAa97H8ukTfbjioSMfGeZWcsD7LwIX1"



#Full Inventory Path 
inventory_path <- "https://docs.google.com/spreadsheets/d/1sFWq-WKhemPzbOFInqhCu_Lx0lsO6a_Z/edit?gid=1740738201#gid=1740738201"


```


PARSE ENTERED DATA FILES: RUN POREWATER CHUNKS, SOURCEWATER & RUNOFF FOR EVENTS  

Porewater
## Read in the POREWATER datasheets and parse
```{r, pull in entered datasheets}

#for just one folder 
#googledrive::drive_ls(path = drive_path, pattern = "Field") -> files

#to look within multipel folders 
files <- googledrive::drive_ls(
  path = pw_drive_path,
  recursive = TRUE,
  pattern = "Field"
)


source("parse_event_porewater_file.R")

parse_files <- list()

for(i in seq_len(nrow(files))) {

    cat("Processing:", files$name[i], "\n")

    drive_download(files$id[i],
                   path = "temp.xlsx",
                   overwrite = TRUE)

    pw_raw <- read_excel("temp.xlsx",
                         col_names = FALSE)

    parse_files[[i]] <- tryCatch(
        parse_event_porewater_file(pw_raw),
        error = function(e) {
            message("Failed on: ", files$name[i])
            message(e$message)
            return(NULL)
        }
    )

    unlink("temp.xlsx")
}

# I think we can use either purrr:map or lapply here but I couldn't figure it out, 
# for now we'll use a for loop

parse_files <- list()

for(i in 1:nrow(files)) {
    
    drive_download(files$id[i], path = "temp.xlsx", overwrite = TRUE)
    read_excel("temp.xlsx", col_names = FALSE) -> pw_raw
    parse_event_porewater_file(pw_raw) -> parse_files[[i]]
    unlink("temp.xlsx")

} 

# When you're ready to upload you can use googledrive::drive_upload

```
# Reorganize POREWATER data to match inventory
```{r}

source("reorg_event_porewater_file.R")

pw_inventory_files <- lapply(parse_files, process_event_inventory)

pw_inventory_all <- bind_rows(pw_inventory_files)


```

Surface Water
## Read in the SOURCEWATER datasheets and parse 
```{r}

#to look within multiple folders 
sce_files <- googledrive::drive_ls(
  path = sce_drive_path,
  recursive = TRUE,
  pattern = "Field"
)

source("parse_sourcewater_file.R")

parse_sce_files <- list()

for(i in seq_len(nrow(sce_files))) {

  cat("Processing:", sce_files$name[i], "\n")

  drive_download(
    sce_files$id[i],
    path = "sce_temp.xlsx",
    overwrite = TRUE
  )

  parse_sce_files[[i]] <- tryCatch(
    parse_sourcewater_file("sce_temp.xlsx"),
    error = function(e) {
      message("Failed on: ", sce_files$name[i])
      message(e$message)
      NULL
    }
  )

  unlink("sce_temp.xlsx")
}


```
# Reorganize SOURCEWATER data to match inventory
```{r}

source("reorg_sourcewater_file.R")

sce_inventory_files <- lapply(parse_sce_files, process_sce_inventory)

sce_inventory_all <- bind_rows(sce_inventory_files)


```

Runoff
## Read in the RUNOFF datasheets and parse 
```{r}

#to look within multipel folders 
ro_files <- googledrive::drive_ls(
  path = ro_drive_path,
  recursive = TRUE,
  pattern = "Field"
)

source("parse_runoff_file.R")

parse_ro_files <- list()

for(i in seq_len(nrow(ro_files))) {

  cat("Processing:", ro_files$name[i], "\n")

  drive_download(
    ro_files$id[i],
    path = "ro_temp.xlsx",
    overwrite = TRUE
  )

  parse_ro_files[[i]] <- tryCatch(
    parse_runoff_file("ro_temp.xlsx"),
    error = function(e) {
      message("Failed on: ", ro_files$name[i])
      message(e$message)
      NULL
    }
  )

  unlink("ro_temp.xlsx")
}


```
# Reorganize RUNOFF data to match inventory
```{r}

source("reorg_runoff_file.R")

ro_inventory_files <- lapply(parse_ro_files, process_ro_inventory)

ro_inventory_all <- bind_rows(ro_inventory_files)


```



NOW COMBINE THE NEW INVENTORY WITH THE OLD - just using a downloaded inventory for now but needs changed to pull from google when we are comfy with this

Porewater
# Prep POREWATER to be appended to the inventory file 
```{#r, process data file to match inventory}

#Read in the master inventory 
master <- read_excel(
  "TEMPEST_PorewaterInventory_May2022_Present.xlsx",
  sheet = "Porewater - Individual",
  skip = 3
)

#clean out any empty sample_ID rows
master <- master %>%
  filter(!is.na(Sample_ID))

# Align the columns
inventory_new <- pw_inventory_all %>%
  select(any_of(names(master)))

missing_cols <- setdiff(names(master), names(inventory_new))
missing_cols

for (col in missing_cols) {
  inventory_new[[col]] <- NA
}

# Append to the inventory
inventory_new <- inventory_new %>%
  select(names(master))

inventory <- bind_rows(inventory_new, master)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS: TEMPEST",
    "Site: SERC",
    "Content: Lysimeter Porewater Inventory",
    "" ))

```
# Replace the POREWATER sheet in the inventory file - not doing this because not pushing to google yet 
```{#r, add the new inventory file}

# Load existing workbook
wb <- loadWorkbook(inventory_path)

# Remove existing sheet if it exists
if ("Inventory" %in% names(wb)) {
  removeWorksheet(wb, "Inventory")
}

# Add replacement sheet
addWorksheet(wb, "Inventory")

# Write header block
writeData(wb, "Inventory", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "Inventory",
  inventory,
  startRow = nrow(header_rows) + 1
)

# Save changes back to the same file

```

Source Water 
# Prep SOURCEWATER to be appended to the inventory file  
```{#r, process data file to match inventory}

#Read in the master inventory 
sce_master <- read_excel(
  #path = inventory_path,
  "TEMPEST_PorewaterInventory_May2022_Present.xlsx",
  sheet = "Source Water",
  skip = 4
)

#clean out any empty sample_ID rows
sce_master <- sce_master %>%
  filter(!is.na(Sample_ID))

# Align the columns
sce_inventory_new <- sce_inventory_all %>%
  select(any_of(names(sce_master)))

missing_cols <- setdiff(names(sce_master), names(sce_inventory_new))
missing_cols

for (col in missing_cols) {
  sce_inventory_new[[col]] <- NA
}

# Append to the inventory
sce_inventory_new <- sce_inventory_new %>%
  select(names(master))

sce_inventory <- bind_rows(sce_inventory_new, master)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS: TEMPEST",
    "Site: SERC",
    "Event: Source Water",
    "" ))

```
# Replace the SOURCEWATER sheet in the inventory file 
```{#r, add the new inventory file}

# Load existing workbook
wb <- loadWorkbook(inventory_path)

# Remove existing sheet if it exists
if ("sce-Inventory" %in% names(wb)) {
  removeWorksheet(wb, "sce-Inventory")
}

# Add replacement sheet
addWorksheet(wb, "sce-Inventory")

# Write header block
writeData(wb, "sce-Inventory", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "sce-Inventory",
  sce_inventory,
  startRow = nrow(header_rows) + 1
)

# Save changes back to the same file

```

Runoff - doesn't have a page in the inventory yet... 



# Create sample lists from the new inventory or whole inventory

```{r, create sample lists}

#Input the start date, end date, and the analyte you want the list for 
start_date <- as.numeric("20260501")
end_date   <- as.numeric("20260612")
target_analyte <- "SO4/Cl/H2S" #"SO4/Cl/H2S"
target_file_name <- "SO4"


#create a function that builds a sample list 
build_sample_list <- function(dat, target_analyte, start_date, end_date) {

  dat %>%
    filter(
      Analyte == target_analyte,
      Collection_Date_YYYYMMDD >= start_date,
      Collection_Date_YYYYMMDD <= end_date
    ) %>%
    mutate(
      Plot = if ("Plot" %in% names(.)) {
        factor(Plot, levels = c("Control", "Saltwater", "Freshwater", "Estuary"))
      } else {
        NA_character_
      }
    ) %>%
    arrange(Collection_Date_YYYYMMDD, Timepoint, Plot) %>%
    transmute(
      Sample_ID,
      Plot,
      Collection_Date_YYYYMMDD,
      Timepoint,
      Analyte
    )
}

#apply the function to each inventory we made 
pw_sample_list <- build_sample_list(
  pw_inventory_all,
  target_analyte,
  start_date,
  end_date
)

sce_sample_list <- build_sample_list(
  sce_inventory_all,
  target_analyte,
  start_date,
  end_date
)

ro_sample_list <- build_sample_list(
  ro_inventory_all,
  target_analyte,
  start_date,
  end_date
)




```

```{r, combine and order sample lists}

#add a type column 
pw_sample_list$type <- "PW"
sce_sample_list$type <- "SCE"
ro_sample_list$type <- "RO"


#order the pw list 
pw_sample_list <- pw_sample_list %>%
  mutate(
    Grid = stringr::str_extract(Sample_ID, "(?<=_)[A-Z]\\d"),
    Depth_cm = stringr::str_extract(Sample_ID, "\\d+(?=cm)") %>% as.numeric()
  )

grid_order <- c("B4","C3","C6", "D5", "E3","F4","H3","H6", "I5")

pw_sample_list <- pw_sample_list %>%
  mutate(
    Grid = factor(Grid, levels = grid_order)
  )


#order the SCE list 
sce_sample_list <- sce_sample_list %>%
  mutate(
    time_order = paste(Collection_Date_YYYYMMDD, Timepoint)
  )

#order the ro list 
ro_sample_list <- ro_sample_list %>%
  mutate(
    time_order = paste(Collection_Date_YYYYMMDD, Timepoint)
  )

#combine the dataframes
all_samples <- bind_rows(
  pw_sample_list,
  sce_sample_list,
  ro_sample_list
)

all_samples <- all_samples %>%
  mutate(
    time_block = Timepoint
  )


all_samples <- all_samples %>%
  mutate(
    type_order = case_when(
      type == "PW" ~ 1,
      type == "SCE" ~ 2,
      type == "RO" ~ 3,
      TRUE ~ 99
    )
  ) %>%
  arrange(
    Collection_Date_YYYYMMDD,
    time_block,
    type_order,
    Plot,
    Grid,
    Depth_cm
  ) %>%
  mutate(
    Sample_Number = row_number()
  )

sample_list <- all_samples %>%
  select(
    Sample_Number,
    Sample_ID,
    Plot,
    Collection_Date_YYYYMMDD,
    Timepoint, 
    type
  )
  


```

```{r, read out sample list  }

#Export to print the list 
printable_list <- createWorkbook()
addWorksheet(printable_list, "Sample List")

#create a header 
writeData(
  printable_list,
  "Sample List",
  paste("TEMPEST Sample List:", target_analyte, start_date, "-", end_date),
  startRow = 1,
  startCol = 1,
  colNames = FALSE
)

writeData(
  printable_list,
  "Sample List",
  sample_list,
  startRow = 2)

#Save the workbook
file_name <- paste0(
  "TEMPEST_",
  target_file_name, "_",
  start_date, "-", end_date,
  ".xlsx")

#export the workbook 
saveWorkbook(printable_list, file_name, overwrite = TRUE)

```


