---
title: "TEMPEST: Porewater CDOM"
author: "2022-2023 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true

---

##This code pulls in CDOM Abs and EEM processed data files from Alan's folders on the Google Drive and renames them based on Alan's Sample Log also from the Google Drive so that the names correspond to sample IDs. 
##In addition, this outputs metadata for each sample ID, including the integration time, dilution, RSU_area_1s columns from the sample log. 

# Need to double check that these sample ID's match all the ones in the collated data for all other analytes and check against metadata/flag data that Jieun said is missing 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F)

# load packages
require(pacman)
pacman::p_load(tidyverse, # keep things tidy
               janitor, # useful for simplifying column names
               googlesheets4, # read_sheet 
               googledrive, # drive_ functions
               plotrix,
               here, 
               readxl, 
               readr, 
               purrr) 

#Try to deal with the timezone demons 
common_tz = "Etc/GMT+5"

Sys.setenv(TZ = "America/New_York")

# Authenticate with your Google account
# drive_auth()

```

# Bring in the sample log:
```{r read in sample log, include=FALSE}

##Read in and combine Alan's sample logs
#create folder
# log_folder_path <- "Sample_Logs"
# dir.create(log_folder_path)
# 
# #folder link to id
# log_folder = "https://drive.google.com/drive/folders/1yEDeKKGTEToyfHYTbs7ZF7363afSu2cQ?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto"
# folder_id = drive_get(as_id(log_folder))
# 
# #find files in folder
# files = drive_ls(folder_id)
# files <- files %>% 
#   filter(name != "20230825_Mapping_COMPASS_TEMPEST_CDOM.xlsx")
# 
# #download files
# for (i in seq_along(files$name)) {
#     #fails if already exists
#       drive_download(
#         as_id(files$id[i]),
#         path = paste0("Sample_Logs/", files$name[i]), 
#         overwrite = TRUE
#       )
# }

##Edited Sample Logs from AMP

#list all sample logs
log_files <- list.files(path = "Edited Sample Logs/logs", 
                    pattern = "*.xlsx", 
                    full.names = TRUE)

#read, convert all columns to character, and combine
sample_log <- log_files %>%
  map(~ {
    
    df <- read_excel(.x) %>%
      mutate(across(everything(), as.character))
    
    # Rename columns only if they exist
    names(df)[names(df) == 
      "Sample Description \r\n(60 Character Max) - Will display on final EEM figure"] <-
      "Sample Description (60 Character Max) - Will display on final EEM figure"
    
    names(df)[names(df) == 
      "Sample Description \n(60 Character Max) - Will display on final EEM figure"] <-
      "Sample Description (60 Character Max) - Will display on final EEM figure"
    
    names(df)[names(df) == 
      "Project Sample ID"] <-
      "Sample Description (60 Character Max) - Will display on final EEM figure"
    
    names(df)[names(df) == 
      "Dilution Factor\r\n*Decimal Format…e.g. 1 part sample and 1 part water = 0.50"] <-
      "Dilution Factor (Decimal Format; e.g. 0.5)"

    names(df)[names(df) == 
      "Absorbance \r\nPathlength"] <-
      "Absorbance Pathlength"
    
    names(df)[names(df) == 
      "Absorbance \nPathlength"] <-
      "Absorbance Pathlength"
    
    df
  }) %>%
  bind_rows()

colnames(sample_log)

sample_log <- sample_log %>% 
  rename(Sample_Description = "Sample Description (60 Character Max) - Will display on final EEM figure", 
         Dilution_Factor = "Dilution Factor (Decimal Format; e.g. 0.5)")

##Remove "EXP" and "Tea" files
sample_log <- sample_log %>% 
  filter(!str_detect(Sample_Description, "EXP")) %>% 
  filter(!str_detect(Sample_Description, "Tea"))

##Add Sample ID for ones that were NA 
sample_log <- sample_log %>% 
  rename(Sample_ID = 'Sample ID') %>% 
  mutate(Sample_ID = ifelse(is.na(Sample_ID), Sample_Description, Sample_ID))

##Fix sample ID for ones that are actually renamed in file ID 
sample_log <- sample_log %>% 
  mutate(Sample_ID = case_when(
      Sample_ID == "TSD51002 2s" ~ Sample_Description,
      Sample_ID == "TSF61002 1s" ~ Sample_Description,
      Sample_ID == "TSC31002 2s" ~ Sample_Description,
      Sample_ID == "TSH31002 2s" ~ Sample_Description,
      Sample_ID == "TFH31002 2s" ~ Sample_Description,
      Sample_ID == "TFI51005 1s" ~ Sample_Description,
      Sample_ID == "TCI51002 2s" ~ Sample_Description,
      TRUE ~ Sample_ID)) # fallback for other values

##Fix some Sample Descriptions
sample_log <- sample_log %>% 
  mutate(Sample_Description = toupper(Sample_Description)) %>% 
  mutate(Sample_Description = gsub("POOLED", "POOL", Sample_Description)) %>% 
  mutate(Sample_Description = gsub("BARGE", "SOURCE", Sample_Description)) %>% 
  mutate(Sample_Description = gsub("SW_ESTUARY", "ESTUARY_SOURCE", Sample_Description)) %>% 
  mutate(Sample_ID = gsub(" ", "", Sample_ID)) 

```


##Pull processed absorbance data from google drive and rename files to save into github 
```{r read in and rename processed absorbance data}
##Processed absorbance data: https://drive.google.com/drive/folders/1_wsS1tiujYuFDlmEOjQZPSEOJKfitWh3?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto

# #Remove the old Abs folder
unlink("Abs_Data", recursive = TRUE)

##Files in Abs folder
abs_folder_path <- "Abs_Data"
dir.create(abs_folder_path)

#folder link to id
abs_folder = "https://drive.google.com/drive/folders/1_wsS1tiujYuFDlmEOjQZPSEOJKfitWh3?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto"
abs_folder_id = drive_get(as_id(abs_folder))

#find files in folder
files = drive_ls(abs_folder_id)
files

#loop dirs and download files inside them
for (i in seq_along(files$name)) {
  #list files
  i_dir = drive_ls(files[i, ])

  #download files
  for (file_i in seq_along(i_dir$name)) {
    #fails if already exists
      drive_download(
        as_id(i_dir$id[file_i]),
        path = paste0("Abs_Data/", i_dir$name[file_i]),
        overwrite = TRUE
      )
  }
}

##Remove "EXP" files and May 23 test files and well data files as these are not part of the TEMPEST project data
files_to_delete <- list.files(
  path = abs_folder_path,
  pattern = "EXP|TMPC23May|TMPFW23May|TMPSW23May",
  full.names = TRUE
)

files_to_delete

file.remove(files_to_delete)


##Remove well samples
samples_to_delete <- sample_log5 %>% 
  filter(str_detect(Sample_Description, "WELL|SW11|FW11"))
    
samples_to_delete_path_abs = paste0("Abs_Data/",samples_to_delete$original_file_name_abs)

samples_to_delete_path_abs

file.remove(samples_to_delete_path_abs)


##List all files
files_abs <- list.files(path = abs_folder_path)
files_abs <- data.frame(files_abs)  
files_abs <- files_abs %>% 
  rename(original_file_name_abs = files_abs)

##Remove extra words so it's just the sample ID 
files_abs <- files_abs %>% 
  mutate(Sample_ID = gsub("_DilCorr_Abs", "", original_file_name_abs)) %>% 
  mutate(Sample_ID = gsub("_ABS", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub(".dat", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub(" ", "", Sample_ID))



```


##Pull processed eem data from google drive and rename files to save into github 
```{r read in and rename processed eem data}
##Processed EEM data: https://drive.google.com/drive/folders/1g_NoLv1qc-VsQZ_1L4Lmy5cgSjNte1sC?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto

#Remove the old EEM folder
unlink("EEM_Data", recursive = TRUE)

##Files in EEM folder
eem_folder_path <- "EEM_Data"
dir.create(eem_folder_path)

#folder link to id
eem_folder = "https://drive.google.com/drive/folders/1g_NoLv1qc-VsQZ_1L4Lmy5cgSjNte1sC?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto"
eem_folder_id = drive_get(as_id(eem_folder))

#find files in folder
files = drive_ls(eem_folder_id)
files

#loop dirs and download files inside them
for (i in seq_along(files$name)) {
  #list files
  i_dir = drive_ls(files[i, ])

  #download files
  for (file_i in seq_along(i_dir$name)) {
    #fails if already exists
      drive_download(
        as_id(i_dir$id[file_i]),
        path = paste0("EEM_Data/", i_dir$name[file_i]),
        overwrite = TRUE
      )
  }
}

##Remove "EXP" files and May 23 test files as these are not part of the TEMPEST project data
files_to_delete <- list.files(
  path = eem_folder_path,
  pattern = "EXP|TMPC23May|TMPFW23May|TMPSW23May",
  full.names = TRUE
)

files_to_delete

file.remove(files_to_delete)

##Remove well samples
samples_to_delete <- sample_log5 %>%
  filter(str_detect(Sample_Description, "WELL|SW11|FW11"))

samples_to_delete_path_eem = paste0("EEM_Data/",samples_to_delete$original_file_name_eem)

samples_to_delete_path_eem

file.remove(samples_to_delete_path_eem)


##List all files
files_eem <- list.files(path = eem_folder_path)
files_eem <- data.frame(files_eem)  
files_eem <- files_eem %>% 
  rename(original_file_name_eem = files_eem)

##Remove extra words so it's just the sample ID 
files_eem <- files_eem %>% 
  mutate(Sample_ID = gsub("_DilCorr_IFE_RamNorm", "", original_file_name_eem)) %>% 
  mutate(Sample_ID = gsub("_EEM", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub(".dat", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub(" ", "", Sample_ID))



```

##Combine Abs and EEM file dataframes
```{r}
files <- merge(files_abs, files_eem, by = "Sample_ID", all = TRUE)
```


##Make new file ID
```{r}
##Combine with sample log
sample_log1 <- files %>% 
  left_join(sample_log, by = "Sample_ID")

unique(sample_log1$Sample_Description)
subset <- subset(sample_log1, str_detect(sample_log$Sample_Description, "TEMPEST"))
pool <- subset(sample_log1, str_detect(sample_log$Sample_Description, "POOL"))

##Fix sample description
sample_log2 <- sample_log1 %>%
  ##TEMPEST pooled samples for 2022
  filter(Sample_Description != "TEMPEST") %>% 
  mutate(Sample_Description = gsub("_CDOM", "", Sample_Description)) %>% 
  ##Remove part of ID that don't need 
  # mutate(Sample_Description = gsub("_SW11", "", Sample_Description)) %>% 
  # mutate(Sample_Description = gsub("_FW11", "", Sample_Description)) %>% 
  ##Date for 2022 Source water samples 
  mutate(Sample_Description = ifelse(str_detect(Sample_Description, "SOURCE_HR"), paste0(Sample_Description, "_20220622"), Sample_Description)) %>% 
  ##Fix PreW based on AMP's code
  mutate(Sample_Description = stringr::str_replace(Sample_Description, "PREW", "T0")) %>% 
  ##Dates for 2022 Porewater samples
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T0", "T0_20220620", Sample_Description), Sample_Description)) %>% 
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T1", "T1_20220622", Sample_Description), Sample_Description)) %>% 
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T2", "T2_20220622", Sample_Description), Sample_Description)) %>% 
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T3", "T3_20220623", Sample_Description), Sample_Description)) %>% 
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T4", "T4_20220624", Sample_Description), Sample_Description)) %>% 
  ##Fix Estuary time
  mutate(Sample_Description = stringr::str_replace(Sample_Description, "TMP_ESTUARY_SOURCE_HR8_20220622", "TMP_ESTUARY_SOURCE_HR7_20220622")) %>% 
  mutate(Sample_Description = stringr::str_replace(Sample_Description, "202306627", "20230627"))




##Separate into columns to make into correct sample IDs
# TMP_FW_C3_20220622_T1_15CM
sample_log3 <- sample_log2 %>%
  mutate(Plot = stringr::str_extract(Sample_Description, "TMP_C|TMP_FW|TMP_SW|TMP_ESTUARY") ,
         Grid = stringr::str_extract(Sample_Description, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5|SOURCE|POOL") ,
         Date = stringr::str_extract(Sample_Description, "[0-9]{8}"), 
         T_num = stringr::str_extract(Sample_Description, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7|HR8"), 
         Time = stringr::str_extract(Sample_Description, "(?<=_)[0-9]{4}$"), 
         Plot = gsub("TMP_", "", Plot))


##T1-3 for sourcewater
sce_2023 <- sample_log3 %>% 
  filter(!is.na(Time)) %>% 
  filter(str_detect(Sample_Description, "SOURCE")) %>% 
  group_by(Plot, Date) %>% 
  arrange(Time, .by_group = TRUE) %>% # Sort by Time ascending
  mutate(T_num = paste0("T", row_number())) %>% 
  ungroup()

##AM/PM for 2023 porewater/pooled samples
pw_2023 <- sample_log3 %>% 
  filter(str_detect(Sample_Description, "20230607")) %>% 
  filter(!str_detect(Sample_Description, "SOURCE")) %>% 
  mutate(TOD = case_when(
    Time < 1200  ~ "AM",
    Time > 1200  ~ "PM",
    TRUE    ~ NA))


##Recombine data
sample_log4 <- sample_log3 %>%
  filter(is.na(Time))

sample_log5 <- bind_rows(sample_log4, sce_2023, pw_2023)


##Make new Sample IDs
##Desired Sample IDs 
#TMP_FW_PW_H6_15cm_CDOM_20250509
#TMP_Source_Estuary_CDOM_20240611_PM

sample_log6 <- sample_log5 %>% 
  mutate(Sample_ID = 
           case_when(
             str_detect(Sample_Description, "SOURCE") & !str_detect(Sample_Description, "HR") ~  paste("TMP", "Source", Plot, T_num, "POOLED", Date, TOD, sep = "_"), 
             str_detect(Sample_Description, "SOURCE") & str_detect(Sample_Description, "HR") ~  paste("TMP", "Source", Plot, T_num, Date, TOD, sep = "_"), 
             str_detect(Sample_Description, "POOL") ~  paste("TMP", Plot, "PW", "15cm", "POOLED", T_num, Date, TOD, sep = "_"), 
             TRUE ~ paste("TMP", Plot, "PW", Grid, "15cm", "CDOM", T_num, Date, TOD, sep = "_"))) %>% ##Make sample ID to match others
  mutate(Sample_ID = gsub("_NA", "", Sample_ID)) %>% 
  
  mutate(Sample_ID = toupper(Sample_ID)) %>% 
  
  ####Sample ID Fixes####
  
  mutate(Sample_ID = gsub("TMP_C_PW_D5_15CM_CDOM_20230627", "TMP_C_PW_D5_15CM_20230628", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_C_PW_15CM_POOLED_20221019", "TMP_C_PW_15CM_POOLED_20221019-24", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_FW_PW_15CM_POOLED_20221019", "TMP_FW_PW_15CM_POOLED_20221019-24", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("POOLED_20221128", "POOLED_202211-12", Sample_ID)) %>%  
  mutate(Sample_ID = gsub("POOLED_20230202", "POOLED_202302", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("POOLED_20230707", "POOLED_202307", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_SW_PW_H6_15CM_CDOM_20230612", "TMP_SW_PW_B4_15CM_20230612", Sample_ID)) %>% 
  
  ##Remove well samples
  filter (!str_detect(Sample_Description, "WELL|SW11|FW11")) %>% 

  mutate(
  Sample_ID = case_when(
    Sample_Description == "TEMPEST CONTROL PLOT - 09/09/22" ~ "TMP_C_PW_15cm_POOLED_20220912",
    Sample_Description == "TEMPEST CONTROL PLOT - 08/08/22" ~ "TMP_C_PW_15cm_POOLED_20220808",
    Sample_Description == "TEMPEST CONTROL PLOT - 07/18/22" ~ "TMP_C_PW_15cm_POOLED_20220718",
    Sample_Description == "TEMPEST CONTROL PLOT - 07/21/22" ~ "TMP_C_PW_15cm_POOLED_20220721",
    Sample_Description == "TEMPEST FRESHWATER PLOT - 09/09/22" ~ "TMP_FW_PW_15cm_POOLED_20220912",
    Sample_Description == "TEMPEST FRESHWATER PLOT -  08/08/22" ~ "TMP_FW_PW_15cm_POOLED_20220808",
    Sample_Description == "TEMPEST FRESHWATER PLOT - 07/18/22" ~ "TMP_FW_PW_15cm_POOLED_20220718",
    Sample_Description == "TEMPEST FRESHWATER PLOT - 07/21/22" ~ "TMP_FW_PW_15cm_POOLED_20220721",
    Sample_Description == "TEMPEST SEAWATER PLOT - 09/09/22" ~ "TMP_SW_PW_15cm_POOLED_20220912",
    Sample_Description == "TEMPEST SEAWATER PLOT - 08/08/22" ~ "TMP_SW_PW_15cm_POOLED_20220808",
    Sample_Description == "TEMPEST SEAWATER PLOT - 07/18/22" ~ "TMP_SW_PW_15cm_POOLED_20220718",
    Sample_Description == "TEMPEST SEAWATER PLOT - 07/21/22" ~ "TMP_SW_PW_15cm_POOLED_20220721", 
    TRUE ~ Sample_ID
  )
) %>% 
  
  mutate(Sample_ID = ifelse(str_detect(Sample_ID, "PW"), gsub("_CDOM", "", Sample_ID), Sample_ID) )%>% 
  mutate(Sample_ID = ifelse(str_detect(Sample_ID, "PW"), gsub("_DOC", "", Sample_ID), Sample_ID))


```

##Create the sample metadata
```{r output metadata}


##Get sample metadata from Sample ID
sample_log7 <- sample_log6 %>%
  mutate(Sample_ID = toupper(Sample_ID)) %>% 
  mutate(Plot = stringr::str_extract(Sample_ID, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5|POOLED") ,
         Date = stringr::str_extract(Sample_ID, "\\d{6}(?:\\d{2}(?:-\\d{2})?)?"), 
         Depth = stringr::str_extract(Sample_ID, "15CM|30CM|50CM"),
         T_num = stringr::str_extract(Sample_ID, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID, "AM|MID|PM"), 
         Sample_Type = stringr::str_extract(Sample_ID, "SOURCE"), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         Grid = ifelse(Sample_Type == "SOURCE", NA, Grid), 
         
         Sample_ID_formatted = paste("TMP", Sample_Type, Plot, Grid, Depth, Date, T_num, TOD, sep = "_"), 
         Sample_ID_formatted = gsub("_NA", "", Sample_ID_formatted)) %>% 
  # select(-Sample_ID) %>% 
  distinct()


#Create desired file ID
sample_log8 <- sample_log7 %>%
  mutate(new_abs_file_id = paste0(Sample_ID_formatted, "_ABS_DilCorr.dat")) %>%
  mutate(new_eem_file_id = paste0(Sample_ID_formatted, "_EEM_DilCorr_IFE_RamNorm.dat"))

sample_log9 <- sample_log8 %>% 
  rename(abs_file_id = new_abs_file_id, 
         eem_file_id = new_eem_file_id, 
         integration_time_s = `Integration Time`, 
         dilution_factor = Dilution_Factor, 
         RSU_area_1s = `Raman 1s Area`) %>% 
  select(c(Sample_ID, Sample_ID_formatted, integration_time_s, dilution_factor, RSU_area_1s, abs_file_id, eem_file_id)) %>% 
  select(-Sample_ID)

```

## Pull in active porewater tracking inventory sheet from Google Drive: 
```{r rest of metadata, include=FALSE}

#Run these to pull the TEMPEST porewater metadata into GitHub if not already there
inventory_directory <- "https://docs.google.com/spreadsheets/d/1WuuUiIieTfmYlZUuQkEBkTkx4bumNjNEGKBcxTJ2viw/edit?gid=260638594#gid=260638594"

drive_download(inventory_directory, overwrite = TRUE)

sheet_names <- excel_sheets("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx")
sheet_names

raw_metadata_lys <- read_excel("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", sheet = "Porewater - Individual", skip = 4)

raw_metadata_sce <- read_excel("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", sheet = "Source Water", skip = 4)

raw_metadata_pool <- read_excel("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", sheet = "Porewater - Pooled", skip = 4)

raw_metadata <- bind_rows(raw_metadata_lys, raw_metadata_sce, raw_metadata_pool)

#Select CDOM Data
raw_metadata_DOC <- subset(raw_metadata, str_detect(Sample_ID, "CDOM|DOC/CDOM|of|POOLED|HR|DOC"))

#Select 2022/2023 Data
raw_metadata_DOC_22_23 <- subset(raw_metadata_DOC, str_detect(Sample_ID, "2022|2023"))

#Create IDs from what was collected for comparison later
raw_metadata_DOC_formatted <- raw_metadata_DOC_22_23 %>%
  mutate(Sample_ID = toupper(Sample_ID)) %>% 
  mutate(Sample_ID = gsub("DOC/CDOM", "CDOM", Sample_ID)) %>% 
  mutate(Sample_ID = case_when(str_detect(Sample_ID, "Source") & str_detect(Sample_ID, "HR") ~ gsub("_POOLED", "", Sample_ID), TRUE ~ Sample_ID)) %>% 
  mutate(Sample_ID = ifelse(str_detect(Sample_ID, "of"), gsub("cm_", "cm_CDOM_", Sample_ID), Sample_ID)) %>% 
  mutate(Sample_ID = sub("_[0-9]{4}$", "", Sample_ID)) %>% 
  select(-c(Plot, Sample_location, Analyte)) %>% 
  
  mutate(Sample_ID = gsub("_1of1", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_1of2", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_2of2", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_1of3", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_2of3", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_3of3", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_3of3", "", Sample_ID)) %>% 
  mutate(Vial_ID = gsub("_1of1", "", Vial_ID)) %>% 
  mutate(Vial_ID = gsub("_1of2", "", Vial_ID)) %>% 
  mutate(Vial_ID = gsub("_2of2", "", Vial_ID)) %>% 
  mutate(Vial_ID = gsub("_1of3", "", Vial_ID)) %>% 
  mutate(Vial_ID = gsub("_2of3", "", Vial_ID)) %>% 
  mutate(Vial_ID = gsub("_3of3", "", Vial_ID)) %>% 
  mutate(Vial_ID = gsub("_2", "", Vial_ID)) %>% 
  
  mutate(Sample_ID = ifelse(str_detect(Sample_ID, "PW"), gsub("_CDOM", "", Sample_ID), Sample_ID))%>% 
  mutate(Sample_ID = ifelse(str_detect(Sample_ID, "PW"), gsub("_DOC", "", Sample_ID), Sample_ID))


raw_metadata_DOC_formatted1 <- raw_metadata_DOC_formatted %>%
  mutate(Plot = stringr::str_extract(Sample_ID, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5|POOLED") ,
         Date = stringr::str_extract(Sample_ID, "\\d{6}(?:\\d{2}(?:-\\d{2})?)?"), 
         Depth = stringr::str_extract(Sample_ID, "15CM|30CM|50CM"),
         T_num = stringr::str_extract(Sample_ID, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID, "AM|MID|PM"), 
         # TOD = case_when(
         #   Date == "20240612" & T_num == "T2" ~ "AM", 
         #   Date == "20240612" & T_num == "T3" ~ "PM", 
         #   Date == "20240613" & T_num == "T4" ~ "AM", 
         #   Date == "20240613" & T_num == "T5" ~ "PM", 
         #   TRUE ~ TOD), 
         Sample_Type = stringr::str_extract(Sample_ID, "SOURCE"), 
         # Sample_Type = gsub("SOURCE_RAINWATER", "RAINWATER", Sample_Type), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         Grid = ifelse(Sample_Type == "SOURCE", NA, Grid), 
         
         Sample_ID_formatted = paste("TMP", Sample_Type, Plot, Grid, Depth, Date, T_num, TOD, sep = "_"), 
         Sample_ID_formatted = gsub("_NA", "", Sample_ID_formatted)) %>% 
  # select(-Sample_ID) %>% 
  distinct()


##Merge pooled rows as needed 
colnames(raw_metadata_DOC_formatted1)

raw_metadata_DOC_formatted2 <- raw_metadata_DOC_formatted1 %>%
  filter(!str_detect(Sample_ID_formatted, "SOURCE")) %>%
  group_by(Sample_ID_formatted) %>%
  summarise(
    Grid_B4 = sum(Grid_B4, na.rm = TRUE),
    Grid_C3 = sum(Grid_C3, na.rm = TRUE),
    Grid_C6 = sum(Grid_C6, na.rm = TRUE),
    Grid_D5 = sum(Grid_D5, na.rm = TRUE),
    Grid_E3 = sum(Grid_E3, na.rm = TRUE),
    Grid_F4 = sum(Grid_F4, na.rm = TRUE),
    Grid_F6 = sum(Grid_F6, na.rm = TRUE),
    Grid_H3 = sum(Grid_H3, na.rm = TRUE),
    Grid_H6 = sum(Grid_H6, na.rm = TRUE),
    Grid_I5 = sum(Grid_I5, na.rm = TRUE),
    Total_Volume_mL = sum(Total_Volume_mL, na.rm = TRUE),

    # Earliest evacuation date + start time from that same row
    Evacuation_date_YYYYMMDD = min(Evacuation_date_YYYYMMDD, na.rm = TRUE),
    Collection_Start_Time_24hrs = Collection_Start_Time_24hrs[which.min(Evacuation_date_YYYYMMDD)],

    # Latest collection date + end time from that same row
    Collection_Date_YYYYMMDD = max(Collection_Date_YYYYMMDD, na.rm = TRUE),
    Collection_End_Time_24hrs = Collection_End_Time_24hrs[which.max(Collection_Date_YYYYMMDD)],

    # First value for every other column
    across(
      -c(
    Grid_B4,
    Grid_C3,
    Grid_C6,
    Grid_D5,
    Grid_E3,
    Grid_F4,
    Grid_F6,
    Grid_H3,
    Grid_H6,
    Grid_I5,
    Evacuation_date_YYYYMMDD, 
    Collection_Start_Time_24hrs,
    Collection_Date_YYYYMMDD,
    Collection_End_Time_24hrs
      
    ),
      first
    ),

    .groups = "drop"
  ) %>%
  bind_rows(
    raw_metadata_DOC_formatted1 %>% filter(str_detect(Sample_ID_formatted, "SOURCE"))
  )


##Remove duplicate sample IDs
raw_metadata_DOC_formatted3 <- raw_metadata_DOC_formatted2 %>% 
  distinct(Sample_ID_formatted, .keep_all = TRUE) %>% 
  select(-Sample_ID)

```


## Check to see if samples run match metadata & merge info
```{r check sample ids with metadata, echo=FALSE, warning = FALSE}

sample_log9$Sample_ID_formatted <- toupper(sample_log9$Sample_ID_formatted) 
raw_metadata_DOC_formatted3$Sample_ID_formatted <- toupper(raw_metadata_DOC_formatted3$Sample_ID_formatted)

all_present <- all(sample_log9$Sample_ID_formatted %in% raw_metadata_DOC_formatted3$Sample_ID_formatted)

if (all_present) {
  message("All sample IDs are present in metadata.")
} else {
  message("Some sample IDs are missing from metadata.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(sample_log9$Sample_ID_formatted, raw_metadata_DOC_formatted3$Sample_ID_formatted)
  print(missing_ids)
}

missing_ids <- data.frame("Missing.IDs" = missing_ids)

write.csv(missing_ids, "Samples_missing_from_metadata.csv", row.names = FALSE)


missing <- sample_log9 %>% 
  filter(Sample_ID_formatted == "TMP_FW_PW_I5_15CM_20230614")

####No metadata####
# "TMP_SOURCE_ESTUARY_HR4_20220622" - no metadata, time should be 110000    
# "TMP_SOURCE_ESTUARY_HR7_20220622" - no metadata, time should be 150000 
# "TMP_C_PW_C3_15CM_20230628" - no metadata, already checked with AMP, no time available  

# "TMP_SW_PW_15CM_T3_20220623" - well sample          
# "TMP_FW_PW_15CM_T2_20220622" - well sample               
# "TMP_FW_PW_15CM_T1_20220622" - well sample              
# "TMP_SW_PW_15CM_T2_20220622" - well sample               
# "TMP_FW_PW_15CM_T3_20220623" - well sample               
# "TMP_FW_PW_15CM_POOLED_20230607" - from FW11 well
# "TMP_SW_PW_15CM_POOLED_20230607" - from SW11 well    


####Fixed####
# "TMP_FW_PW_I5_15CM_20230662" - 20230627
# "TMP_C_PW_D5_15CM_20230627" - should be TMP_C_PW_D5_15CM_20230627       
# "TMP_C_PW_15CM_POOLED_20221019" - should be TMP_C_PW_15CM_POOLED_20221019-24       
# "TMP_C_PW_15CM_POOLED_20230202" - should be TMP_C_PW_15CM_POOLED_202302       
# "TMP_FW_PW_15CM_POOLED_20221019" - should be TMP_FW_PW_15CM_POOLED_20221019-24     
# "TMP_FW_PW_15CM_POOLED_20230202" - should be TMP_FW_PW_15CM_POOLED_202302   
# "TMP_FW_PW_15CM_POOLED_20230707"- should be TMP_FW_PW_15CM_POOLED_202307   
# "TMP_SW_PW_H6_15CM_20230612" - could be TMP_SW_PW_B4_15CM_20230612 since missing that one          
# "TMP_SW_PW_15CM_POOLED_20230202" - should be TMP_SW_PW_15CM_POOLED_202302    
# "TMP_SW_PW_15CM_POOLED_20230707" - should be TMP_SW_PW_15CM_POOLED_202307  

# "TMP_PW_15CM": 
# TEMPEST CONTROL PLOT - 09/09/22 - should be TMP_C_PW_15cm_POOLED_20220912
# TEMPEST CONTROL PLOT - 08/08/22 - should be TMP_C_PW_15cm_POOLED_20220808
# TEMPEST CONTROL PLOT - 07/18/22 - should be TMP_C_PW_15cm_POOLED_20220718
# TEMPEST CONTROL PLOT - 07/21/22 - should be TMP_C_PW_15cm_POOLED_20220721    
# TEMPEST FRESHWATER PLOT - 09/09/22 - should be TMP_FW_PW_15cm_POOLED_20220912
# TEMPEST FRESHWATER PLOT -  08/08/22"  - should be TMP_FW_PW_15cm_POOLED_20220808
# TEMPEST FRESHWATER PLOT - 07/18/22"   - should be TMP_FW_PW_15cm_POOLED_20220718
# TEMPEST FRESHWATER PLOT - 07/21/22"  - should be TMP_FW_PW_15cm_POOLED_20220721    
# TEMPEST SEAWATER PLOT - 09/09/22 - should be TMP_SW_PW_15cm_POOLED_20220912
# TEMPEST SEAWATER PLOT -  08/08/22"  - should be TMP_SW_PW_15cm_POOLED_20220808
# TEMPEST SEAWATER PLOT - 07/18/22"   - should be TMP_SW_PW_15cm_POOLED_20220718
# TEMPEST SEAWATER PLOT - 07/21/22"  - should be TMP_SW_PW_15cm_POOLED_20220721   

##Replace CDOM with "POOLED"
# "TMP_SOURCE_ESTUARY_T1_CDOM_20230606"
# "TMP_SOURCE_ESTUARY_T2_CDOM_20230606" 
# "TMP_SOURCE_ESTUARY_T3_CDOM_20230606" 
# "TMP_SOURCE_ESTUARY_T1_CDOM_20230607"  
# "TMP_SOURCE_ESTUARY_T2_CDOM_20230607" 
# "TMP_SOURCE_ESTUARY_T3_CDOM_20230607" 
# "TMP_SOURCE_FW_T1_CDOM_20230606"     
# "TMP_SOURCE_FW_T2_CDOM_20230606"      
# "TMP_SOURCE_FW_T3_CDOM_20230606"      
# "TMP_SOURCE_FW_T1_CDOM_20230607"      
# "TMP_SOURCE_FW_T2_CDOM_20230607"     
# "TMP_SOURCE_FW_T3_CDOM_20230607"      
# "TMP_SOURCE_SW_T1_CDOM_20230606"      
# "TMP_SOURCE_SW_T2_CDOM_20230606"      
# "TMP_SOURCE_SW_T3_CDOM_20230606"     
# "TMP_SOURCE_SW_T1_CDOM_20230607"      
# "TMP_SOURCE_SW_T2_CDOM_20230607"      
# "TMP_SOURCE_SW_T3_CDOM_20230607"     
```

##Merge the metadata and export
```{r}

#merge metadata with sample run data 
merged_data <- sample_log9 %>%
  left_join(raw_metadata_DOC_formatted3, by = "Sample_ID_formatted")


# "TMP_SOURCE_ESTUARY_HR4_20220622" - no metadata, time should be 110000    
# "TMP_SOURCE_ESTUARY_HR7_20220622" - no metadata, time should be 150000 
# TMP_PW_C_C3_15CM_20230628 - no metadata, no time

merged_data1 <- merged_data %>%
  mutate(Plot = stringr::str_extract(Sample_ID_formatted, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID_formatted, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5|POOLED") ,
         Depth = stringr::str_extract(Sample_ID_formatted, "15CM|30CM|50CM"),
         T_num = stringr::str_extract(Sample_ID_formatted, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID_formatted, "AM|MID|PM"), 
         Sample_Type = stringr::str_extract(Sample_ID_formatted, "SOURCE"), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         Grid = ifelse(Sample_Type == "SOURCE", NA, Grid)) %>% 
  mutate(Collection_Time_24hrs = case_when(
    Sample_ID_formatted == "TMP_SOURCE_ESTUARY_20220622_HR4" ~ 1100, 
    Sample_ID_formatted == "TMP_SOURCE_ESTUARY_20220622_HR7" ~ 1500, 
    Sample_ID_formatted == "TMP_SOURCE_SW_20220622_HR0" ~ 0645, 
    Sample_ID_formatted == "TMP_SOURCE_FW_20220622_HR0" ~ 0654, 
    Sample_ID_formatted == "TMP_SOURCE_SW_20220622_HR1" ~ 0800, 
    Sample_ID_formatted == "TMP_SOURCE_FW_20220622_HR1" ~ 0800, 
    Sample_ID_formatted == "TMP_SOURCE_SW_20220622_HR2" ~ 0930, 
    Sample_ID_formatted == "TMP_SOURCE_FW_20220622_HR2" ~ 0930, 
    Sample_ID_formatted == "TMP_SOURCE_SW_20220622_HR3" ~ 1000, 
    Sample_ID_formatted == "TMP_SOURCE_FW_20220622_HR3" ~ 1000, 
    Sample_ID_formatted == "TMP_SOURCE_SW_20220622_HR4" ~ 1100, 
    Sample_ID_formatted == "TMP_SOURCE_FW_20220622_HR4" ~ 1100, 
    Sample_ID_formatted == "TMP_SOURCE_SW_20220622_HR5" ~ 1200, 
    Sample_ID_formatted == "TMP_SOURCE_FW_20220622_HR5" ~ 1200, 
    Sample_ID_formatted == "TMP_SOURCE_SW_20220622_HR7" ~ 1448, 
    Sample_ID_formatted == "TMP_SOURCE_FW_20220622_HR7" ~ 1448, 
    TRUE ~ Collection_Time_24hrs
  )) %>%
  mutate(Collection_Date_YYYYMMDD = case_when(
    str_detect(Sample_ID_formatted, "20220622") ~ 20220622, 
    Sample_ID_formatted == "TMP_PW_C_C3_15CM_20230628" ~ 20230628, 
    TRUE ~ Collection_Date_YYYYMMDD
  )) %>% 
  mutate(Notes = case_when(
    Sample_ID_formatted == "TMP_C_PW_C3_15CM_20230628" ~ "No metadata", 
    TRUE ~ Notes
  )) %>% 
  #manual entry of missing evacuation information from the notes/inventory sheet: 
  mutate(Evacuation_date_YYYYMMDD = case_when(
    Collection_Date_YYYYMMDD == 20220620 ~ 20220618,
    Collection_Date_YYYYMMDD == 20220622 ~ 20220622,
    Collection_Date_YYYYMMDD == 20220623 ~ 20220622,
    Collection_Date_YYYYMMDD == 20220624 ~ 20220623,
    Collection_Date_YYYYMMDD == 20220909 ~ 20220909,
    Collection_Date_YYYYMMDD == 20230606 ~ 20230605,
    Collection_Date_YYYYMMDD == 20230607 & str_detect(Sample_ID_formatted, "AM") ~ 20230606,
    Collection_Date_YYYYMMDD == 20230607 & str_detect(Sample_ID_formatted, "PM") ~ 20230607,
    Collection_Date_YYYYMMDD == 20230612 ~ 20230608,
    Collection_Date_YYYYMMDD == 20230627 ~ 20230623,
    Collection_Date_YYYYMMDD == 20230628 ~ 20230623,
    Collection_Date_YYYYMMDD == 20230707 ~ 20230705,
    TRUE ~ Evacuation_date_YYYYMMDD))

merged_data2 <- merged_data1 %>% 
  rename(Sample_ID = Sample_ID_formatted) %>% 
  mutate(Collection_End_Time_24hrs = ifelse(is.na(Collection_End_Time_24hrs), Collection_Time_24hrs, Collection_End_Time_24hrs)) %>% 
  select("Sample_ID", "integration_time_s", "dilution_factor", "RSU_area_1s", "abs_file_id", "eem_file_id", 
         "Evacuation_date_YYYYMMDD", "Collection_Date_YYYYMMDD", "Collection_Start_Time_24hrs", "Collection_End_Time_24hrs", "Notes", "Deployment_Date_YYYYMMDD", 
         "Deployment_Time_24hrs", "EST_EDT")


write.csv(merged_data2, "CDOM_metadata_2022-2023.csv", row.names = FALSE)

```

##Rename all absorbance files
```{r}

##Rename files
for(i in 1:nrow(sample_log8)){
    
    # Create file paths with folder and file name
    old_file <- file.path(abs_folder_path, sample_log8$original_file_name_abs[i])
    new_file <- file.path(abs_folder_path, sample_log8$new_abs_file_id[i])
    
    # Check if the original file actually exists
  if (file.exists(old_file)) {
    
    # Rename the file
    file.rename(from = old_file, to = new_file)
    
    cat("Success: Renamed", sample_log8$original_file_name_abs[i], "to", sample_log8$new_abs_file_id[i], "\n")
  } else {
    cat("Warning: File", sample_log8$original_file_name_abs[i], "not found.\n")
  }
}

##Convert files from .dat to .csv

# Get a list of all .dat files in that folder
abs_files <- list.files(path = abs_folder_path, pattern = "\\.dat$", full.names = TRUE, ignore.case = TRUE)

# Loop through each file and convert it
for (file in abs_files) {
  
  # 1. Read the file (Change 'sep' depending on your delimiter)
  # Use sep = "" for spaces, sep = "\t" for tabs, or sep = "," for commas
  temp_data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
  # 2. Create the new file path by changing the extension to .csv
  csv_file <- sub("\\.dat$", ".csv", file, ignore.case = TRUE)
  
  # 3. Write out the data frame as a CSV
  write.csv(temp_data, file = csv_file, row.names = FALSE)
  
  # 4. Remove old .dat file
  file.remove(file)

}

print("Conversion complete!")
```


##Rename all EEM files
```{r}

##Rename files
for(i in 1:nrow(sample_log8)){
    
    # Create file paths with folder and file name
    old_file <- file.path(eem_folder_path, sample_log8$original_file_name_eem[i])
    new_file <- file.path(eem_folder_path, sample_log8$new_eem_file_id[i])
    
    # Check if the original file actually exists
  if (file.exists(old_file)) {
    
    # Rename the file
    file.rename(from = old_file, to = new_file)
    
    cat("Success: Renamed", sample_log8$original_file_name_eem[i], "to", sample_log8$new_eem_file_id[i], "\n")
  } else {
    cat("Warning: File", sample_log8$original_file_name_eem[i], "not found.\n")
  }
}


##Convert files from .dat to .csv
# Get a list of all .dat files in that folder
eem_files <- list.files(path = eem_folder_path, pattern = "\\.dat$", full.names = TRUE, ignore.case = TRUE)

# Loop through each file and convert it
for (file in eem_files) {
  
  # 1. Read the file (Change 'sep' depending on your delimiter)
  # Use sep = "" for spaces, sep = "\t" for tabs, or sep = "," for commas
  temp_data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
  # 2. Create the new file path by changing the extension to .csv
  csv_file <- sub("\\.dat$", ".csv", file, ignore.case = TRUE)
  
  # 3. Write out the data frame as a CSV
  write.csv(temp_data, file = csv_file, row.names = FALSE)
  
  # 4. Remove old .dat file
  file.remove(file)

}

print("Conversion complete!")
```


##Rename the abs file columns 
```{r}
library(readr)
library(dplyr)
library(purrr)

abs_files <- list.files(path = abs_folder_path, pattern = "\\.csv$", full.names = TRUE, ignore.case = TRUE)

walk(abs_files, function(file) {

  # Read assuming there IS a header first
  dat <- read_csv(file, show_col_types = FALSE)

  # If first column name looks like wavelength, file had a header
  if (grepl("Wavelength", names(dat)[1], ignore.case = TRUE)) {

    names(dat)[1:2] <- c("Wavelength_nm", "Dilution_Corrected_Absorbance")

  } else {

    # No header: re-read so first data row isn't lost
    dat <- read_csv(
      file,
      col_names = c("Wavelength_nm", "Dilution_Corrected_Absorbance"),
      show_col_types = FALSE
    )
  }

  # Overwrite original file
  write_csv(dat, file)
})

####EEM files
eem_files <- list.files(path = eem_folder_path, pattern = "\\.csv$", full.names = TRUE, ignore.case = TRUE)

walk(eem_files, function(file) {

  # Read assuming there IS a header first
  dat <- read_csv(file, show_col_types = FALSE)
  
  ##Remove X from Column names
  names(dat) <- gsub("X", "", names(dat))
  names(dat)[1] <- c("Wavelength_nm")

  # Overwrite original file
  write_csv(dat, file)
})

# dat <- read_csv(eem_files[1], show_col_types = FALSE)
# names(dat)
```

