---
title: "TEMPEST: Porewater CDOM"
author: "2022-2023 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true

---

##This code pulls in CDOM Abs and EEM processed data files from Alan's folders on the Google Drive and renames them based on Alan's Sample Log also from the Google Drive so that the names correspond to sample IDs. 
##In addition, this outputs metadata for each sample ID, including the integration time, dilution, RSU_area_1s columns from the sample log. 

# Need to double check that these sample ID's match all the ones in the collated data for all other analytes and check against metadata/flag data that Jieun said is missing 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F)

# load packages
require(pacman)
pacman::p_load(tidyverse, # keep things tidy
               janitor, # useful for simplifying column names
               googlesheets4, # read_sheet 
               googledrive, # drive_ functions
               plotrix,
               here, 
               readxl, 
               readr, 
               purrr) 

#Try to deal with the timezone demons 
common_tz = "Etc/GMT+5"

Sys.setenv(TZ = "America/New_York")

# Authenticate with your Google account
drive_auth()

```

# Bring in the sample log:
```{r read in sample log, include=FALSE}

##Read in and combine Alan's sample logs
#create folder
log_folder_path <- "Sample_Logs"
dir.create(log_folder_path)

#folder link to id
log_folder = "https://drive.google.com/drive/folders/1yEDeKKGTEToyfHYTbs7ZF7363afSu2cQ?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto"
folder_id = drive_get(as_id(log_folder))

#find files in folder
files = drive_ls(folder_id)
files <- files %>% 
  filter(name != "20230825_Mapping_COMPASS_TEMPEST_CDOM.xlsx")

#download files
for (i in seq_along(files$name)) {
    #fails if already exists
      drive_download(
        as_id(files$id[i]),
        path = paste0("Sample_Logs/", files$name[i]), 
        overwrite = TRUE
      )
}

#list all sample logs
log_files <- list.files(path = "Sample_Logs", 
                    pattern = "*.xlsx", 
                    full.names = TRUE)

#read, convert all columns to character, and combine
sample_log <- log_files %>%
  map(~ {
    
    df <- read_excel(.x) %>%
      mutate(across(everything(), as.character))
    
    # Rename columns only if they exist
    names(df)[names(df) == 
      "Sample Description \r\n(60 Character Max) - Will display on final EEM figure"] <-
      "Sample Description (60 Character Max) - Will display on final EEM figure"
    
    names(df)[names(df) == 
      "Sample Description \n(60 Character Max) - Will display on final EEM figure"] <-
      "Sample Description (60 Character Max) - Will display on final EEM figure"
    
    names(df)[names(df) == 
      "Project Sample ID"] <-
      "Sample Description (60 Character Max) - Will display on final EEM figure"
    
    names(df)[names(df) == 
      "Dilution Factor\r\n*Decimal Format…e.g. 1 part sample and 1 part water = 0.50"] <-
      "Dilution Factor (Decimal Format; e.g. 0.5)"

    names(df)[names(df) == 
      "Absorbance \r\nPathlength"] <-
      "Absorbance Pathlength"
    
    names(df)[names(df) == 
      "Absorbance \nPathlength"] <-
      "Absorbance Pathlength"
    
    df
  }) %>%
  bind_rows()

colnames(sample_log)

sample_log <- sample_log %>% 
  rename(Sample_Description = "Sample Description (60 Character Max) - Will display on final EEM figure", 
         Dilution_Factor = "Dilution Factor (Decimal Format; e.g. 0.5)")

##Remove "EXP" and "Tea" files
sample_log <- sample_log %>% 
  filter(!str_detect(Sample_Description, "EXP")) %>% 
  filter(!str_detect(Sample_Description, "Tea"))

##Add Sample ID for ones that were NA 
sample_log <- sample_log %>% 
  rename(Sample_ID = 'Sample ID') %>% 
  mutate(Sample_ID = ifelse(is.na(Sample_ID), Sample_Description, Sample_ID))

##Fix sample ID for ones that are actually renamed in file ID 
sample_log <- sample_log %>% 
  mutate(Sample_ID = case_when(
      Sample_ID == "TSD51002 2s" ~ Sample_Description,
      Sample_ID == "TSF61002 1s" ~ Sample_Description,
      Sample_ID == "TSC31002 2s" ~ Sample_Description,
      Sample_ID == "TSH31002 2s" ~ Sample_Description,
      Sample_ID == "TFH31002 2s" ~ Sample_Description,
      Sample_ID == "TFI51005 1s" ~ Sample_Description,
      Sample_ID == "TCI51002 2s" ~ Sample_Description,
      TRUE ~ Sample_ID)) # fallback for other values

##Fix some Sample Descriptions
sample_log <- sample_log %>% 
  mutate(Sample_Description = toupper(Sample_Description)) %>% 
  mutate(Sample_Description = gsub("POOLED", "POOL", Sample_Description)) %>% 
  mutate(Sample_Description = gsub("BARGE", "SOURCE", Sample_Description)) %>% 
  mutate(Sample_Description = gsub("SW_ESTUARY", "ESTUARY_SOURCE", Sample_Description)) %>% 
  mutate(Sample_ID = gsub(" ", "", Sample_ID)) 

```


##Pull processed absorbance data from google drive and rename files to save into github 
```{r read in and rename processed absorbance data}
##Processed absorbance data: https://drive.google.com/drive/folders/1_wsS1tiujYuFDlmEOjQZPSEOJKfitWh3?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto

#Remove the old Abs folder
unlink("Abs_Data", recursive = TRUE)

##Files in Abs folder
abs_folder_path <- "Abs_Data"
dir.create(abs_folder_path)

#folder link to id
abs_folder = "https://drive.google.com/drive/folders/1_wsS1tiujYuFDlmEOjQZPSEOJKfitWh3?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto"
abs_folder_id = drive_get(as_id(abs_folder))

#find files in folder
files = drive_ls(abs_folder_id)
files

#loop dirs and download files inside them
for (i in seq_along(files$name)) {
  #list files
  i_dir = drive_ls(files[i, ])
  
  #download files
  for (file_i in seq_along(i_dir$name)) {
    #fails if already exists
      drive_download(
        as_id(i_dir$id[file_i]),
        path = paste0("Abs_Data/", i_dir$name[file_i]), 
        overwrite = TRUE
      )
  }
}

##Remove "EXP" files and May 23 test files as these are not part of the TEMPEST project data
files_to_delete <- list.files(
  path = abs_folder_path, 
  pattern = "EXP|TMPC23May|TMPFW23May|TMPSW23May", 
  full.names = TRUE
)

files_to_delete

file.remove(files_to_delete)


##List all files
files_abs <- list.files(path = abs_folder_path)
files_abs <- data.frame(files_abs)  
files_abs <- files_abs %>% 
  rename(original_file_name_abs = files_abs)

##Remove extra words so it's just the sample ID 
files_abs <- files_abs %>% 
  mutate(Sample_ID = gsub("_DilCorr_Abs", "", original_file_name_abs)) %>% 
  mutate(Sample_ID = gsub("_ABS", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub(".dat", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub(" ", "", Sample_ID))



```


##Pull processed eem data from google drive and rename files to save into github 
```{r read in and rename processed eem data}
##Processed EEM data: https://drive.google.com/drive/folders/1g_NoLv1qc-VsQZ_1L4Lmy5cgSjNte1sC?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto

#Remove the old EEM folder
unlink("EEM_Data", recursive = TRUE)

##Files in EEM folder
eem_folder_path <- "EEM_Data"
dir.create(eem_folder_path)

#folder link to id
eem_folder = "https://drive.google.com/drive/folders/1g_NoLv1qc-VsQZ_1L4Lmy5cgSjNte1sC?dmr=1&ec=wgc-drive-%5Bmodule%5D-goto"
eem_folder_id = drive_get(as_id(eem_folder))

#find files in folder
files = drive_ls(eem_folder_id)
files

#loop dirs and download files inside them
for (i in seq_along(files$name)) {
  #list files
  i_dir = drive_ls(files[i, ])
  
  #download files
  for (file_i in seq_along(i_dir$name)) {
    #fails if already exists
      drive_download(
        as_id(i_dir$id[file_i]),
        path = paste0("EEM_Data/", i_dir$name[file_i]), 
        overwrite = TRUE
      )
  }
}

##Remove "EXP" files and May 23 test files as these are not part of the TEMPEST project data
files_to_delete <- list.files(
  path = eem_folder_path, 
  pattern = "EXP|TMPC23May|TMPFW23May|TMPSW23May", 
  full.names = TRUE
)

files_to_delete

file.remove(files_to_delete)


##List all files
files_eem <- list.files(path = eem_folder_path)
files_eem <- data.frame(files_eem)  
files_eem <- files_eem %>% 
  rename(original_file_name_eem = files_eem)

##Remove extra words so it's just the sample ID 
files_eem <- files_eem %>% 
  mutate(Sample_ID = gsub("_DilCorr_IFE_RamNorm", "", original_file_name_eem)) %>% 
  mutate(Sample_ID = gsub("_EEM", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub(".dat", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub(" ", "", Sample_ID))



```

##Combine Abs and EEM file dataframes
```{r}
files <- merge(files_abs, files_eem, by = "Sample_ID", all = TRUE)
```


##Make new file ID
```{r}
##Combine with sample log
sample_log <- files %>% 
  left_join(sample_log, by = "Sample_ID")


##Desired Sample IDs 
#TMP_FW_H6_20220620_T0_15CM
#TMP_FW_SOURCE_20220622_T0

##Fix sample description
sample_log1 <- sample_log %>%
  mutate(Sample_Description = gsub("TEMPEST CONTROL PLOT", "TMP_C_SOURCE", Sample_Description)) %>% ##Fix sourcewater formats
  mutate(Sample_Description = gsub("TEMPEST FRESHWATER PLOT", "TMP_FW_SOURCE", Sample_Description)) %>%
  mutate(Sample_Description = gsub("TEMPEST SEAWATER PLOT", "TMP_SW_SOURCE", Sample_Description)) %>%
  mutate(Sample_Description = gsub(" - 09/09/22", "_20220909", Sample_Description)) %>% ##Fix date formats
  mutate(Sample_Description = gsub(" - 08/08/22", "_20220808", Sample_Description)) %>%
  mutate(Sample_Description = gsub(" -  08/08/22", "_20220808", Sample_Description)) %>%
  mutate(Sample_Description = gsub(" - 07/18/22", "_20220718", Sample_Description)) %>%
  mutate(Sample_Description = gsub(" - 07/21/22", "_20220721", Sample_Description)) %>% 
  # mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), paste0(Sample_Description, "_2022"), Sample_Description)) %>% ##Add year to 2022 samples
  mutate(Sample_Description = gsub("_CDOM", "", Sample_Description)) %>% ##Remove CDOM since these are all CDOM samples
  mutate(Sample_Description = ifelse(str_detect(Sample_Description, "SOURCE_HR"), paste0(Sample_Description, "_20220622"), Sample_Description)) %>% ##Date for 2022 Source water samples 
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T0", "T0_20220620", Sample_Description), Sample_Description)) %>% ##Date for 2022 Porewater samples
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T1", "T1_20220622", Sample_Description), Sample_Description)) %>% 
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T2", "T2_20220622", Sample_Description), Sample_Description)) %>% 
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), gsub("T3", "T3_20220623", Sample_Description), Sample_Description)) %>% 
  mutate(Sample_Description = ifelse(str_detect(Sample_ID, "2022_CDOM"), sub("T4", "T4_20220624", Sample_Description), Sample_Description))


##Separate into columns to make into correct sample IDs
# TMP_FW_C3_20220622_T1_15CM
sample_log2 <- sample_log1 %>%
  separate(
    col = Sample_Description,
    sep = "_",
    into = c("Project", "Plot", "Grid", "Date", "Timepoint"),
    remove = FALSE) %>% 
  mutate(temp_Date = Timepoint) %>% ##Fix Dates and Timepoints
  mutate(Timepoint = ifelse(str_detect(Date, "T|HR"), Date, Timepoint)) %>% 
  mutate(Date = ifelse(str_detect(Date, "T|HR"), temp_Date, Date)) %>% 
  mutate(Timepoint = ifelse(str_detect(Timepoint, "2022"), NA, Timepoint)) %>% 
  mutate(Depth = ifelse(Grid == "SOURCE"|Grid == "WELL", NA, "15cm"))  %>% ##Add 15 cm depth for all porewater samples
  mutate(Time_of_day = ifelse(str_detect(Timepoint, "T|HR"), NA, Timepoint))

##T1-3 for sourcewater
sw_2023 <- sample_log2 %>% 
  filter(!is.na(Time_of_day)) %>% 
  filter(str_detect(Sample_Description, "SOURCE")) %>% 
  group_by(Plot, Date) %>% 
  arrange(Time_of_day, .by_group = TRUE) %>% # Sort by Time ascending
  mutate(Timepoint = paste0("T", row_number())) %>% 
  ungroup()

##AM/PM for 2023 porewater/pooled samples
pw_2023 <- sample_log2 %>% 
  filter(str_detect(Sample_Description, "20230607")) %>% 
  filter(!str_detect(Sample_Description, "SOURCE")) %>% 
  mutate(Timepoint = case_when(
    Time_of_day < 1200  ~ "AM",
    Time_of_day > 1200  ~ "PM",
    TRUE    ~ Timepoint))


##Recombine data
sample_log3 <- sample_log2 %>%
  filter(is.na(Time_of_day))

sample_log4 <- bind_rows(sample_log3, sw_2023, pw_2023)



##Make new Sample IDs
sample_log5 <- sample_log4 %>% 
  mutate(Sample_ID = paste(Project, Plot, Grid, Date, Timepoint, Depth, sep = "_")) %>% ##Make sample ID to match others
  mutate(Sample_ID = gsub("_NA", "", Sample_ID))


#Create desired file ID
sample_log5 <- sample_log5 %>%
  mutate(new_abs_file_id = paste0(Sample_ID, "_ABS_DilCorr.dat")) %>%
  mutate(new_eem_file_id = paste0(Sample_ID, "_EEM_DilCorr_IFE_RamNorm.dat"))


```

##Rename all absorbance files
```{r}

##Rename files
for(i in 1:nrow(sample_log5)){
    
    # Create file paths with folder and file name
    old_file <- file.path(abs_folder_path, sample_log5$original_file_name_abs[i])
    new_file <- file.path(abs_folder_path, sample_log5$new_abs_file_id[i])
    
    # Check if the original file actually exists
  if (file.exists(old_file)) {
    
    # Rename the file
    file.rename(from = old_file, to = new_file)
    
    cat("Success: Renamed", sample_log5$original_file_name_abs[i], "to", sample_log5$new_abs_file_id[i], "\n")
  } else {
    cat("Warning: File", sample_log5$original_file_name_abs[i], "not found.\n")
  }
}

##Convert files from .dat to .csv

# Get a list of all .dat files in that folder
abs_files <- list.files(path = abs_folder_path, pattern = "\\.dat$", full.names = TRUE, ignore.case = TRUE)

# Loop through each file and convert it
for (file in abs_files) {
  
  # 1. Read the file (Change 'sep' depending on your delimiter)
  # Use sep = "" for spaces, sep = "\t" for tabs, or sep = "," for commas
  temp_data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
  # 2. Create the new file path by changing the extension to .csv
  csv_file <- sub("\\.dat$", ".csv", file, ignore.case = TRUE)
  
  # 3. Write out the data frame as a CSV
  write.csv(temp_data, file = csv_file, row.names = FALSE)
  
  # 4. Remove old .dat file
  file.remove(file)

}

print("Conversion complete!")
```


##Rename all EEM files
```{r}

##Rename files
for(i in 1:nrow(sample_log5)){
    
    # Create file paths with folder and file name
    old_file <- file.path(eem_folder_path, sample_log5$original_file_name_eem[i])
    new_file <- file.path(eem_folder_path, sample_log5$new_eem_file_id[i])
    
    # Check if the original file actually exists
  if (file.exists(old_file)) {
    
    # Rename the file
    file.rename(from = old_file, to = new_file)
    
    cat("Success: Renamed", sample_log5$original_file_name_eem[i], "to", sample_log5$new_eem_file_id[i], "\n")
  } else {
    cat("Warning: File", sample_log5$original_file_name_eem[i], "not found.\n")
  }
}


##Convert files from .dat to .csv
# Get a list of all .dat files in that folder
eem_files <- list.files(path = eem_folder_path, pattern = "\\.dat$", full.names = TRUE, ignore.case = TRUE)

# Loop through each file and convert it
for (file in eem_files) {
  
  # 1. Read the file (Change 'sep' depending on your delimiter)
  # Use sep = "" for spaces, sep = "\t" for tabs, or sep = "," for commas
  temp_data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
  # 2. Create the new file path by changing the extension to .csv
  csv_file <- sub("\\.dat$", ".csv", file, ignore.case = TRUE)
  
  # 3. Write out the data frame as a CSV
  write.csv(temp_data, file = csv_file, row.names = FALSE)
  
  # 4. Remove old .dat file
  file.remove(file)

}

print("Conversion complete!")
```



##Create the metadata
```{r output metadata}
colnames(sample_log5)
metadata <- sample_log5 %>% 
  rename(abs_file_id = new_abs_file_id, 
         eem_file_id = new_eem_file_id, 
         integration_time_s = `Integration Time`, 
         dilution_factor = Dilution_Factor, 
         RSU_area_1s = `Raman 1s Area`) %>% 
  select(c(Sample_ID, integration_time_s, dilution_factor, RSU_area_1s, abs_file_id, eem_file_id))

write.csv(metadata, "CDOM_metadata_2022-2023.csv")

```

