---
title: "COMPASS_TEMPEST_Collate_Dionex_2022_Data"
author: "Zoe Read"
date: "2026-07-29"
output: html_document
---

# From tempest_pw_june_event_pooled.R script: 
```{r}
## This script imports the metadata for pooled pore water collection during
## the June 2022 TEMPEST event and organizes it in a format that
## can be used for further data manipulation.
## Raw Data are read in from GoogleDrive
## 
## Created: 2023-01 by Julia McElhinny for TEMPEST, with some code edited from
## processing scripts written by Allison Myers-Pigg.
## Updated for systeme level analysis on 10/5/23

# 1. Load Packages -------------------------------------------------------------

require(pacman)
pacman::p_load(tidyverse,
               broom,
               janitor,
               googlesheets4,
               googledrive)


# 2. Load Data -----------------------------------------------------------------

# check working directory - should be the root of the repository
getwd()

# link directory to google drive location
directory <- "https://drive.google.com/drive/folders/1l6ChfOs6wU82ujsoAUw9dGNfoOVL4E8R"

# check list of files found at the link
drive_ls(directory)

# filter to necessary files
# June 2022 event
files_pulse <- drive_ls(directory) %>%
  filter(grepl("TMP Data", name))

# download files to local drive
lapply(c(files_pulse$name), drive_download, 
       overwrite = TRUE)

lapply(split(files_pulse, seq_len(nrow(files_pulse))), 
       drive_download, overwrite = TRUE)


# some files were not xlsx format in file names but downloaded as such - updating  
# names to reflect this
for (n in 1:nrow(files_pulse)) {
  
  # if the name doesn't have the excel extension, add it
   if (!grepl(".xlsx", files_pulse$name[n], fixed = TRUE)) {
     files_pulse$name[n] <- paste0(files_pulse$name[n], ".xlsx")
   }
}

# read in data by named tabs - creates lists of tibbles
control_pulse <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                         sheet = "CONTROL", 
                                                                          range = cell_rows(7:17)))

freshwater_pulse <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                            sheet = "FRESHWATER", 
                                                                            range = cell_rows(7:17)))

seawater_pulse <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                           sheet = "SEAWATER", 
                                                                           range = cell_rows(7:17)))


# 3. Pulse Data Edits ----------------------------------------------------------

## edit each data frame in the lists to have the collection time period
# pull time periods out of spreadsheet names
pulse_times <- as.data.frame(substr(files_pulse$name, 1,2))

# make sure the number date frame of each list matches that same number time
# in the pulse_times list
b = 1

for (a in 1:nrow(pulse_times)) {
  
  control_pulse[[b]] <- mutate(control_pulse[[b]], timepoint = pulse_times[a,1])
  freshwater_pulse[[b]] <- mutate(freshwater_pulse[[b]], timepoint = pulse_times[a,1])
  seawater_pulse[[b]] <- mutate(seawater_pulse[[b]], timepoint = pulse_times[a,1])
  
  # set index value for next round
  b = b + 1 
}

# make list of new column names for better readability
pulse_column_names <- c("Plot", "Grid", "DOC", "Ions", "Water_Isotopes", "HR-MS", 
                        "Saccharides", "Amino_Acids", "Pooled", "CDOM_Subsample",
                        "Time_Notes", "Notes", "Total_Volume", "Evacuated_cb", "Timepoint")

for (c in 1:length(control_pulse)) {

    # change column names  
    colnames(control_pulse[[c]]) <- pulse_column_names
      # make all columns character for future merging
      control_pulse[[c]] <- mutate(control_pulse[[c]], across(everything(), as.character))
  
    colnames(freshwater_pulse[[c]]) <- pulse_column_names
      freshwater_pulse[[c]] <- mutate(freshwater_pulse[[c]], across(everything(), as.character))
      
    colnames(seawater_pulse[[c]]) <- pulse_column_names
      seawater_pulse[[c]] <- mutate(seawater_pulse[[c]], across(everything(), as.character))
}

# merge all 15 tibbles together into one data frame
list <- c(control_pulse, freshwater_pulse, seawater_pulse)

pulse_all_plots <- bind_rows(list)
pulse_all_plots <- as.data.frame(pulse_all_plots)

# edit different forms of NA values
pulse_all_plots[pulse_all_plots == "—"] <- NA
pulse_all_plots[pulse_all_plots == "NA"] <- NA
pulse_all_plots[pulse_all_plots == "--"] <- NA

# total volume column contains some notes - separate those out
total_volume_notes <- pulse_all_plots %>%
  filter(grepl("discarded|keep|threw", Total_Volume)) # can put these notes into the Notes column

# total volume cells with notes replaced with NA (to indicate collection but no use for data purposes)
pulse_all_plots$Notes[grepl("discarded|keep|threw", pulse_all_plots$Total_Volume)] <- total_volume_notes$Total_Volume
pulse_all_plots$Total_Volume[grepl("discarded|keep|threw", pulse_all_plots$Total_Volume)] <- NA
pulse_all_plots$Pooled[grepl("discarded|keep|threw", pulse_all_plots$Notes)] <- NA


# 4. Manipulate Lysimeter Collection Time/Date Information ---------------------
# this code is unnecessary if read_excel reads in the times the right way in the column
# key is that if one tab has some times entered, all rows need to have one, even if 00:00 needs to be entered

# read in the time information from the sheets for individual lysimeters
#SW_T4 <- readxl::read_excel(path = files_pulse$name[5],
                   #sheet = "SEAWATER", 
                   #skip = 6) %>%
              #mutate(Time_Notes = format(Time_Notes, "%I:%M:%S %p")) %>%
              #mutate(Timepoint = "T4")

#SW_T2 <- readxl::read_excel(path = files_pulse$name[4],
                            #sheet = "SEAWATER", 
                            #skip = 6) %>%
  #mutate(Time_Notes = format(Time_Notes, "%I:%M:%S %p")) %>%
  #mutate(Timepoint = "T2")

#FW_T1 <- readxl::read_excel(path = files_pulse$name[1],
                            #sheet = "FRESHWATER", 
                            #skip = 6) %>%
  #mutate(Time_Notes = format(Time_Notes, "%I:%M:%S %p")) %>%
  #mutate(Timepoint = "T1")

# join the three data frames together where this data was collected
#colnames(SW_T4) <- pulse_column_names
  #SW_T4<- mutate(SW_T4, across(everything(), as.character))

#colnames(SW_T2) <- pulse_column_names
  #SW_T2<- mutate(SW_T2, across(everything(), as.character))
  
#colnames(FW_T1) <- pulse_column_names
  #FW_T1<- mutate(FW_T1, across(everything(), as.character))
  #FW_T1$Time_Notes[6] <- NA #entered as 00:00 in the parent google sheet to avoid read-in errors
  #FW_T1 <- FW_T1[-11,]
  
#collection_times <- rbind(SW_T4, SW_T2)
#collection_times <- rbind(collection_times, FW_T1)


# 5. Finalize Compiled Data Frame of Collection Volumes ------------------------

# replace incorrectly loaded time information
#pulse_all_plots$Time_Notes[!is.na(pulse_all_plots$Time_Notes)] <- collection_times$Time_Notes

# edit time formats read in automatically by read_excel
pulse_all_plots$Time_Notes[pulse_all_plots$Plot == "FRESHWATER" & pulse_all_plots$Timepoint == "T1" & pulse_all_plots$Grid == "B4"] <- NA

# fix variable types
pulse_all_plots[,c(3:10,13,14)] <- as.numeric(unlist(pulse_all_plots[,c(3:10,13,14)]))
pulse_all_plots[,"Plot"] <- factor(pulse_all_plots[,"Plot"], levels = c("CONTROL", "FRESHWATER", "SEAWATER"))
pulse_all_plots$Grid <- as.factor(pulse_all_plots$Grid)
pulse_all_plots$Timepoint <- factor(pulse_all_plots$Timepoint, levels = c("T0", 
                                                                          "T1", "T2",
                                                                          "T3", "T4"))

# make R recognize datetime and extract just time information as character values
pulse_all_plots$Time_Notes <- as.POSIXct(pulse_all_plots$Time_Notes, format = "%Y-%m-%d %H:%M:%S")
pulse_all_plots$Time_Notes <- format(pulse_all_plots$Time_Notes, format = "%H:%M:%S")


# 6. Read In Date/Time Info for Each Timepoint ---------------------------------

# make list of cells to extract
sheet_cells <- c("B3", "D3", "F3")

# make list of tab names
tab_names <- c("CONTROL", "FRESHWATER", "SEAWATER")

# pull individual cells out of excel spreadsheets
# control date
control_dates <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                         sheet = tab_names[1], 
                                                                         range = sheet_cells[1]))
# control start time
control_start <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                         sheet = tab_names[1], 
                                                                         range = sheet_cells[2]))
# control end time
control_end <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                         sheet = tab_names[1], 
                                                                         range = sheet_cells[3]))
# fw date
fw_dates <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                         sheet = tab_names[2], 
                                                                         range = sheet_cells[1]))
# fw start time
fw_start <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                         sheet = tab_names[2], 
                                                                         range = sheet_cells[2]))
# fw end time
fw_end <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                         sheet = tab_names[2], 
                                                                         range = sheet_cells[3]))

sw_dates <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                    sheet = tab_names[3], 
                                                                    range = sheet_cells[1]))

sw_start <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                    sheet = tab_names[3], 
                                                                    range = sheet_cells[2]))

sw_end <- lapply(files_pulse$name, function(x) readxl::read_excel(path = x,
                                                                  sheet = tab_names[3], 
                                                                  range = sheet_cells[3]))

dates_list <- list(control_dates, fw_dates, sw_dates)
start_list <- list(control_start, fw_start, sw_start)
end_list <- list(control_end, fw_end, sw_end)

## extract values from the lists and add timepoint/plot information
# make data frame to drop those values in
dates_times <- data.frame(matrix(NA,
                                     nrow=15,
                                     ncol =5))
colnames(dates_times) <- c("Plot", "Timepoint", "Date", "Start_time", "End_time")

# set index value for the row number where data is going to be inserted
d = 1

# cycle through plots
for (f in 1:length(tab_names)) {
  
  # for each plot cycle through the 5 timepoints
  for (e in 1:nrow(pulse_times)) {
  
    # check if a value is present
    # if present value, pull it and put it in the "date" column
    if (!is.na(colnames(dates_list[[f]][[e]])[1])) {
    dates_times[d,3] <- colnames(dates_list[[f]][[e]])[1]
      }

    # start time column
    if (!is.na(colnames(start_list[[f]][[e]])[1])) {
    dates_times[d,4] <- colnames(start_list[[f]][[e]])[1]
      }

    # end time column  
    if (!is.na(colnames(end_list[[f]][[e]])[1])) {
    dates_times[d,5] <- colnames(end_list[[f]][[e]])[1]
    }
    
    # add time point information
    dates_times[d,2] <- pulse_times[e,1]
    
    # add grid information
    dates_times[d,1] <- tab_names[f]
  
    # increase row number so no overwriting    
    d = d + 1
  }
}

## further data frame edits
# edit NA values
dates_times[dates_times == "—"] <- NA

# edit variable types
dates_times[,1:2] <- lapply(dates_times[,1:2], function(x) as.factor(x))
dates_times[,3:5] <- lapply(dates_times[,3:5], function(x) as.numeric(x))

# make excel serial dates into MDY
dates_times$Date <- excel_numeric_to_date(dates_times$Date)

# make decimal times into HMS
dates_times[,4:5] <- lapply(dates_times[,4:5], function(x) convert_to_datetime(x))
dates_times[,4:5] <- lapply(dates_times[,4:5], function(x) format(x, format = "%H:%M:%S"))

# replace NA dates with dates listed for that time point on other plots
dates_tofill <- dates_times %>%
  filter(!is.na(Date)) %>%
  select(Timepoint, Date)
dates_tofill <- unique(dates_tofill)

for (g in 1:nrow(dates_times)) {
  if (is.na(dates_times$Date[g])) {
    
    # get the date for replacing based on match between timepoint
    missing_date <- dates_tofill$Date[dates_tofill$Timepoint == dates_times$Timepoint[g]]
    
    # do the replacement
    dates_times$Date[g] <- missing_date
  
  }
}


# 7. Combine Data Frames for Final Export --------------------------------------

TMP_Event_June2022_Metadata <- left_join(pulse_all_plots, dates_times)

#reorder columns for easier reading
TMP_Event_June2022_Metadata <- TMP_Event_June2022_Metadata[,c(1,15:18,2:10,13,14,11,12)]

# edit plot names to acronyms
# TMP_Event_June2022_Metadata <- TMP_Event_June2022_Metadata %>%
#   mutate(Plot = case_when(
#     Plot == "CONTROL" ~ "C",
#     Plot == "FRESHWATER" ~ "FW",
#     Plot == "SEAWATER" ~ "SW"
#   ))

# fill in start time gaps - pulling the earliest start time reported based on other plots at same timepoint
# T1 control plot set to 12:45
TMP_Event_June2022_Metadata$Start_time[is.na(TMP_Event_June2022_Metadata$Start_time) & TMP_Event_June2022_Metadata$Timepoint == "T1"] <- "12:45:00"

# T2 freshwater set to 16:14
TMP_Event_June2022_Metadata$Start_time[is.na(TMP_Event_June2022_Metadata$Start_time) & TMP_Event_June2022_Metadata$Timepoint == "T2"] <- "16:14:00"

# T3 freshwater set to 9:38
TMP_Event_June2022_Metadata$Start_time[is.na(TMP_Event_June2022_Metadata$Start_time) & TMP_Event_June2022_Metadata$Timepoint == "T3"] <- "9:38:00"


# write out as CSV with all volumes collected by grid by plot by time for future records
write.csv(TMP_Event_June2022_Metadata, "TMP_Event_June2022_gridlevel_volumes.csv",
        row.names = FALSE)


# # # 8. Create CSV File for Further Pooled Sample Calculations --------------------
# 
# # trim down columns
# tmp_event_pooled_volumes <- TMP_Event_June2022_Metadata[,-c(7:12,14:17)]
# 
# # replace NA with 0 for calculations
# tmp_event_pooled_volumes$Pooled[is.na(tmp_event_pooled_volumes$Pooled)] <- 0
# 
# # calculate total pooled volume for each of the plots at each time period
# tmp_event_pooled_sum <- tmp_event_pooled_volumes %>%
#   group_by(Plot, Timepoint) %>%
#   summarize(total_pooled_volume = sum(Pooled))
# 
# # merge the total volume back with original data frame
# tmp_event_pooled_volumes_tidy <- left_join(tmp_event_pooled_volumes, tmp_event_pooled_sum)
# 
# # write out RDS file
# # write_rds(tmp_event_pooled_volumes_tidy, file = "tmp_2022_event_pooled_pw_volumes_tidy.rds")
# 
# # alter format for easier viewing
# tmp_event_pooled_volumes1 <- tmp_event_pooled_volumes %>%
# spread(Grid, Pooled)
# # would need more edits to make sure this function works to only have one row per plot/time combo
# 
# # write out as CSV with all volumes collected by grid by plot by time for future records
# write.csv(tmp_event_pooled_volumes1, "TMP_Event_June2022_pooled_volumes.csv",
#         row.names = FALSE)
# 
# 9. Edit Source Water Metadata ------------------------------------------------

# load source water info
files_source_water <- drive_ls(directory) %>%
  filter(grepl("Source", name))

# download files to local drive
lapply(c(files_source_water$name), drive_download,
       overwrite = TRUE)

# add ".xlsx" to file name
files_source_water$name <- paste0(files_source_water$name, ".xlsx")

# make data frame
source_water_meta <- readxl::read_excel(files_source_water$name)
source_water_meta <- source_water_meta %>%
  clean_names()
 
# # save complete source water data as csv file
# #write.csv(source_water_meta, "./data/raw/DOC meta/June2022_Event_SourceWater_Metadata.csv")
# 
# edit water type column
source_water_meta_edit <- source_water_meta %>%
  rename(Plot = source_water_type,
         Date = date,
         Start_time = time) %>%
  mutate(Plot = case_when(Plot == "Freshwater" ~ "FW",
                          Plot == "Seawater" ~ "SW"),
         Grid = "SOURCE",
         Timepoint = gsub(".*SOURCE_", "", label_name))

# edit the time format
source_water_meta_edit$Start_time <- as.POSIXct(source_water_meta_edit$Start_time, format = "%Y-%m-%d %H:%M:%S")
source_water_meta_edit$Start_time <- format(source_water_meta_edit$Start_time, format = "%H:%M:%S")

# cut columns to match the porewater metadata
source_water_datetime <- source_water_meta_edit %>%
  select(Plot, Grid, Timepoint, Date, Start_time) %>%
  mutate(End_time = NA)
 
# write out as CSV with all volumes collected by grid by plot by time for future records
write.csv(source_water_datetime, "TMP_Event_June2022_sourcewater_volumes.csv", row.names = FALSE)


# # 10. Make Final Date/Time Metadata File for June 2022 Event -------------------
# 
# # separate out the unique values of start time and timepoint
# pw_timepoint_times <- TMP_Event_June2022_Metadata %>%
#   ungroup() %>%
#   group_by(Timepoint) %>%
#   select(Date, Start_time) %>%
#   unique()
# 
# # make a copy of the full metadata data frame
# event_pw_datetime <- TMP_Event_June2022_Metadata
# 
# # fill in start time gaps - pulling the earliest start time reported based on other plots at same timepoint
# # T1 control plot set to 12:45
# event_pw_datetime$Start_time[is.na(event_pw_datetime$Start_time) & event_pw_datetime$Timepoint == "T1"] <- "12:45:00"
# 
# # T2 freshwater set to 16:14
# event_pw_datetime$Start_time[is.na(event_pw_datetime$Start_time) & event_pw_datetime$Timepoint == "T2"] <- "16:14:00"
# 
# # T3 freshwater set to 9:38
# event_pw_datetime$Start_time[is.na(event_pw_datetime$Start_time) & event_pw_datetime$Timepoint == "T3"] <- "9:38:00"
# 
# # reduce number of columns to merge
# event_pw_datetime <- event_pw_datetime %>%
#   select(Plot, Timepoint, Date, Start_time, End_time) %>%
#   mutate(Grid = NA) %>%
#   unique()
# 
# # merge source data and porewater data together
# June2022_event_source_pw_datetime <- rbind(event_pw_datetime, source_water_datetime)
# 
# 
# # 11. Final Clean-Up -----------------------------------------------------------
# 
# # write out csv of event dates and times
# write_rds(June2022_event_source_pw_datetime, file = "TMP_Event_June2022_META_PW_SOURCE_DateTime.rds")
# write.csv(June2022_event_source_pw_datetime, "TMP_Event_June2022_META_PW_SOURCE_DateTime.csv",
#           row.names = FALSE)
# 
# # delete google sheet files
# file.remove(c(files_pulse$name, files_source_water$name))


```

##Notes
```{r}
Sample_Year = "2022"

##Porewater samples that do not have metadata after checking
no_meta_flags <- c("TMP_ESTUARY_SOURCE_20220622_T0", "TMP_ESTUARY_SOURCE_20220622_T1", "TMP_SW_H3_20220615", "TMP_C_F6_20220615")
```

##Packages
```{r setup, include=FALSE}

library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(openxlsx)
library(googledrive)
library(googlesheets4)

#Packages that are required 
lapply(c(
  "dplyr", "ggplot2", "ggpubr", "stringr",
  "purrr", "tidyverse", "here", "broom", "tibble",
  "googledrive", "googlesheets4", "data.table", 
  "matrixStats", "gridExtra", "grid", "readxl", "stringr"), 
  library, character.only = TRUE)

#Packages for loading metadata
require(pacman)
pacman::p_load(janitor, # useful for simplifying column names
               googlesheets4, # read_sheet 
               googledrive, # drive_ functions
               plotrix,
               here) 

Sys.setenv(TZ = "America/New_York")
```

##Pull in 2022 Event Porewater Metadata
```{r pull in event porewater metadata for later, }

##Event metadata was entered separately from other 2022 metadata. 
#Used tempest_pw_june_event_pooled.R code to pull those together from google drive into a metadata file, which is then pulled in here. 
# "https://drive.google.com/drive/folders/1l6ChfOs6wU82ujsoAUw9dGNfoOVL4E8R"

##Read in event metadata
raw_metadata_event <- read.csv("TMP_Event_June2022_gridlevel_volumes.csv")

##Select only rows that have sample volumes and remove pooled samples
raw_metadata_event_samples <- raw_metadata_event %>% 
  filter(!is.na(DOC|Ions|Water_Isotopes|HR.MS|Saccharides|Amino_Acids|Pooled|CDOM_Subsample)) %>% 
  select(-c(Pooled, HR.MS))

##Pivot dataframe so have volume per sample type
raw_metadata_event_long <- raw_metadata_event_samples %>%
  rename(ISO = Water_Isotopes, 
         CDOM = CDOM_Subsample) %>% 
  pivot_longer(
    cols = c(DOC, Ions, ISO, Saccharides,
             Amino_Acids, CDOM),
    names_to = "Analyte",
    values_to = "Volume_mL"
  )


##Create IDs
raw_metadata_event_ids <- raw_metadata_event_long %>%
  mutate(Collection_Date_YYYYMMDD = as.numeric(gsub("-", "", Date))) %>%  ##Format date
  mutate(Analyte = gsub("_", "", Analyte)) %>%  ##Format analyte
  mutate(Plot = gsub("SEAWATER", "SALTWATER", Plot)) %>% ##Format Plot
  mutate(Plot_abb = case_when(
    Plot == "CONTROL" ~ "C",
    Plot == "FRESHWATER" ~ "FW",
    Plot == "SALTWATER" ~ "SW"
  )) %>% 
  mutate(Project = "COMPASS TEMPEST", 
         Plot = str_to_title(Plot), 
         sample_location = NA, 
         Evacuation_date_YYYYMMDD = as.numeric(Collection_Date_YYYYMMDD), 
         EST_EDT = "EDT") %>% 
  mutate(Sample_ID = paste("TMP",
                          Plot_abb,
                          "PW", 
                          Grid,
                          "15cm", 
                          Analyte, 
                          Timepoint, 
                          Collection_Date_YYYYMMDD, 
                          sep = "_")) %>% ##Create Sample IDs
  mutate(Vial_ID = paste(Plot_abb,
                          Grid,
                          "15", 
                          Analyte, 
                          Timepoint, 
                          Collection_Date_YYYYMMDD, 
                          sep = "_")) %>% ##Create Vial IDs
  mutate(
    Collection_Start_Time_24hrs = substr(Start_time, 1, 5), 
    Collection_Start_Time_24hrs = as.numeric(gsub(":", "", Collection_Start_Time_24hrs)), 
    Collection_End_Time_24hrs = substr(End_time, 1, 5), 
    Collection_End_Time_24hrs = as.numeric(gsub(":", "", Collection_End_Time_24hrs))
  ) %>% ##Format columns
  select(Project, Plot, Sample_ID, Volume_mL, Vial_ID, sample_location, Analyte, Evacuation_date_YYYYMMDD, Collection_Date_YYYYMMDD, Collection_Start_Time_24hrs, Collection_End_Time_24hrs, EST_EDT, Notes) %>% 
  filter(Volume_mL>0) ##Remove samples with no volume


```

## Pull in pre-treatment porewater metadata
```{r rest of metadata, include=FALSE}

# Run these to pull the TEMPEST porewater metadata into GitHub if not already there
inventory_directory1 <- "https://docs.google.com/spreadsheets/d/1rOAM_nK3SRaa9iNx1fvrmWzkIvYs-Prw/edit?gid=399569871#gid=399569871"

drive_download(inventory_directory1, overwrite = TRUE)

sheet_names1 <- excel_sheets("Jun2022_PW_FieldData.xlsx")
sheet_names1

raw_metadata_pt <- read_excel("Jun2022_PW_FieldData.xlsx", sheet = "Jun 22", skip = 5)

colnames(raw_metadata_pt) <- c("Plot", "Grid", "Depth_cm", "Volume_mL_20220613", "Volume_mL_20220615", "Bulk_mL_20220613",  "Bulk_mL_20220615", "Notes", "NA", "NA1")

##Convert to long format
raw_metadata_pt_long <- raw_metadata_pt %>%
  select("Plot", "Grid", "Depth_cm", "Volume_mL_20220613", "Volume_mL_20220615", "Notes") %>%
  mutate(
    Volume_mL_20220613 = as.character(Volume_mL_20220613), 
    Volume_mL_20220615 = as.character(Volume_mL_20220615) 
  ) %>%
  pivot_longer(
    cols = starts_with("Volume_mL_"),
    names_to = "Collection_Date_YYYYMMDD",
    names_prefix = "Volume_mL_",
    values_to = "Volume_mL"
  )


##Create IDs
raw_metadata_pt_ids <- raw_metadata_pt_long %>%
  mutate(Collection_Date_YYYYMMDD = as.numeric(Collection_Date_YYYYMMDD)) %>%  ##Format date
  mutate(Plot = gsub("SEAWATER", "SALTWATER", Plot)) %>% ##Format Plot
  mutate(Plot_abb = case_when(
    Plot == "CONTROL" ~ "C",
    Plot == "FRESHWATER" ~ "FW",
    Plot == "SALTWATER" ~ "SW"
  )) %>% 
  mutate(Start_time = case_when(
    Collection_Date_YYYYMMDD == "20220613" ~ "1100",
    Collection_Date_YYYYMMDD == "20220615" ~ "1200")) %>% 
  mutate(End_time = case_when(
    Collection_Date_YYYYMMDD == "20220613" ~ "1330",
    Collection_Date_YYYYMMDD == "20220615" ~ NA)) %>% 
  mutate(Project = "COMPASS TEMPEST", 
         Plot = str_to_title(Plot), 
         sample_location = NA, 
         Evacuation_date_YYYYMMDD = as.numeric(Collection_Date_YYYYMMDD), 
         EST_EDT = "EDT", 
         Analyte = NA) %>% 
  mutate(Sample_ID = paste("TMP",
                          Plot_abb,
                          "PW", 
                          Grid,
                          "15cm", 
                          Collection_Date_YYYYMMDD, 
                          sep = "_")) %>% ##Create Sample IDs
  mutate(Vial_ID = paste(Plot_abb,
                          Grid,
                          "15", 
                          Collection_Date_YYYYMMDD, 
                          sep = "_")) %>% ##Create Vial IDs
  mutate(
    Collection_Start_Time_24hrs = substr(Start_time, 1, 5), 
    Collection_Start_Time_24hrs = as.numeric(gsub(":", "", Collection_Start_Time_24hrs)), 
    Collection_End_Time_24hrs = substr(End_time, 1, 5), 
    Collection_End_Time_24hrs = as.numeric(gsub(":", "", Collection_End_Time_24hrs))
  ) %>% ##Format columns
  select(Project, Plot, Sample_ID, Volume_mL, Vial_ID, sample_location, Analyte, Evacuation_date_YYYYMMDD, Collection_Date_YYYYMMDD, 
         Collection_Start_Time_24hrs, Collection_End_Time_24hrs, EST_EDT, Notes) %>% 
  mutate(Volume_mL = as.numeric(Volume_mL)) %>% 
  filter(Volume_mL>0) ##Remove samples with no volume

```


## Combine event and pre-treatment metadata
```{r}

pw_metadata <- bind_rows(raw_metadata_event_ids, raw_metadata_pt_ids)

##Remove unneccesary notes and fix column data types
pw_metadata <- pw_metadata %>% 
  mutate(Notes = ifelse(str_detect(Notes, "vibe|Vibe"), NA, Notes)) %>% 
  mutate(Notes = ifelse(str_detect(Notes, "frog"), NA, Notes)) %>% 
  mutate(sample_location = as.character(sample_location))

unique(pw_metadata$Notes)

```


NOW COMBINE THE NEW INVENTORY WITH THE OLD
Porewater
# Prep POREWATER to be appended to the inventory file 
```{r, process data file to match inventory}

#Full Inventory Path 
# ####Need to change this link each time####
# inventory_path <- "https://docs.google.com/spreadsheets/d/1J4aPIRrKrPi_o8c7WzkY6cenWtoK8UgOrzIR5GiRBgg/edit?gid=680737830#gid=680737830"
# 
# #Folder to upload the new files to
# folder_path = as_id("1oEeiU48OULo0UOu0np3l1JhYYEQcnw7e")
# 
# # Run this to pull the TEMPEST porewater metadata into GitHub if not already there
# drive_download(inventory_path, overwrite = TRUE)
# 
# #Read in the master inventory
# master <- read_excel(
#   "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
#   sheet = "Porewater - Individual",
#   skip = 4
# )

# link directory to google drive location
inventory_path <- "https://drive.google.com/drive/folders/1oEeiU48OULo0UOu0np3l1JhYYEQcnw7e"

# check list of files found at the link
drive_ls(inventory_path)

# filter to necessary files
# June 2022 event
files_meta <- drive_ls(inventory_path) %>%
  filter(grepl("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", name))

# download files to local drive
lapply(c(files_meta$name), drive_download, 
       overwrite = TRUE)

# lapply(split(files_pulse, seq_len(nrow(files_pulse))), 
#        drive_download, overwrite = TRUE)


# "https://drive.google.com/drive/folders/1l6ChfOs6wU82ujsoAUw9dGNfoOVL4E8R"

#Read in the master inventory
master <- read_excel(
  "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
  sheet = "Porewater - Individual",
  skip = 4, 
  col_types = "text"
)


#clean out any empty sample_ID rows
master <- master %>%
  filter(!is.na(Sample_ID))

#Change column types
master <- master %>%
  mutate(
    Volume_mL = as.numeric(Volume_mL), 
    Evacuation_date_YYYYMMDD = as.numeric(Evacuation_date_YYYYMMDD), 
    Collection_Date_YYYYMMDD = as.numeric(Collection_Date_YYYYMMDD), 
    Collection_Start_Time_24hrs = as.numeric(Collection_Start_Time_24hrs), 
    Collection_End_Time_24hrs = as.numeric(Collection_End_Time_24hrs)
  )

# colnames(inventory_new)

# Align the columns
inventory_new <- pw_metadata %>%
  select(any_of(names(master)))

missing_cols <- setdiff(names(master), names(inventory_new))
missing_cols

for (col in missing_cols) {
  inventory_new[[col]] <- NA
}

# Append to the inventory
inventory_new <- inventory_new %>%
  select(names(master))

pw_inventory <- bind_rows(inventory_new, master)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS",
    "Site: TEMPEST",
    "Content: Lysimeter Porewater Inventory",
    "" ))

```


# Replace the POREWATER sheet in the inventory file
```{r, add the new inventory file}

# Run this to pull the TEMPEST porewater metadata into GitHub if not already there
# drive_download(inventory_path, overwrite = TRUE)


# Load existing workbook
wb <- loadWorkbook("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx")

# Remove existing sheet if it exists
if ("Porewater - Individual" %in% names(wb)) {
  removeWorksheet(wb, "Porewater - Individual")
}

# Add replacement sheet
addWorksheet(wb, "Porewater - Individual")

# Write header block
writeData(wb, "Porewater - Individual", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "Porewater - Individual",
  pw_inventory,
  startRow = nrow(header_rows)+1
)

##Bold the headers
header_style <- createStyle(
  textDecoration = "bold",
  halign = "center",
  valign = "center"
)
addStyle(
  wb,
  sheet = "Porewater - Individual",
  style = header_style,
  rows = 1:5,
  cols = 1:ncol(pw_inventory),
  gridExpand = TRUE
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", overwrite = TRUE)

```



## Pull in 2022 Pooled Metadata
```{r pull in event metadata for later, }

##Event metadata was entered separately from other 2022 metadata. 
#Used tempest_pw_june_event_pooled.R code to pull those together from google drive into a metadata file, which is then pulled in here. 
# "https://drive.google.com/drive/folders/1l6ChfOs6wU82ujsoAUw9dGNfoOVL4E8R"

##Pivot dataframe to get pooled and spe analytes with volumes
raw_metadata_pooled_analytes <- raw_metadata_event %>%
  select(-c(DOC, Ions, Water_Isotopes, Saccharides,
             Amino_Acids, CDOM_Subsample, Total_Volume)) %>% 
  rename(SPE = HR.MS) %>% 
  pivot_longer(
    cols = c(Pooled, SPE),
    names_to = "Analyte",
    values_to = "Volume_mL_Grid"
  ) %>% 
  mutate(Volume_mL_Grid = as.numeric(Volume_mL_Grid)) %>% 
  mutate(Volume_mL_Grid = ifelse(is.na(Volume_mL_Grid), 0, Volume_mL_Grid)) ##Replace NA values with zeroes

##Pivot dataframe so have volume per plot and timepoint
raw_metadata_pooled_samples <- raw_metadata_pooled_analytes %>%
    pivot_wider(
      id_cols = c(Plot, Date, Timepoint, Start_time, End_time, Analyte), 
      names_from = Grid, 
      values_from = Volume_mL_Grid, 
      names_prefix = "Grid_", 
    ) %>% 
    mutate(Total_Volume_mL = rowSums(across(c(Grid_B4, 
                                              Grid_C3, 
                                              Grid_C6, 
                                              Grid_D5, 
                                              Grid_E3, 
                                              Grid_F4, 
                                              Grid_F6, 
                                              Grid_H3, 
                                              Grid_H6, 
                                              Grid_I5)), na.rm = TRUE)) %>% 
    filter(Total_Volume_mL > 0)



##Create IDs
pooled_inventory_event <- raw_metadata_pooled_samples %>%
  mutate(Collection_Date_YYYYMMDD = as.numeric(gsub("-", "", Date))) %>%  ##Format date
  mutate(Analyte = toupper(Analyte)) %>% 
  mutate(Plot = gsub("SEAWATER", "SALTWATER", Plot)) %>% ##Format Plot
  mutate(Plot_abb = case_when(
    Plot == "CONTROL" ~ "C",
    Plot == "FRESHWATER" ~ "FW",
    Plot == "SALTWATER" ~ "SW"
  )) %>% 
  mutate(Project = "COMPASS TEMPEST", 
         Plot = str_to_title(Plot), 
         Sample_location = as.character(NA), 
         Evacuation_date_YYYYMMDD = as.numeric(Collection_Date_YYYYMMDD), 
         EST_EDT = "EDT") %>% 
  mutate(Sample_ID = paste("TMP",
                          Plot_abb,
                          "PW", 
                          "15cm", 
                          Analyte, 
                          Timepoint, 
                          Collection_Date_YYYYMMDD, 
                          sep = "_")) %>% ##Create Sample IDs
  mutate(Vial_ID = paste(Plot_abb,
                          "15", 
                          Analyte, 
                          Timepoint, 
                          Collection_Date_YYYYMMDD, 
                          sep = "_")) %>% ##Create Vial IDs
  mutate(
    Collection_Start_Time_24hrs = substr(Start_time, 1, 5), 
    Collection_Start_Time_24hrs = as.numeric(gsub(":", "", Collection_Start_Time_24hrs)), 
    Collection_End_Time_24hrs = substr(End_time, 1, 5), 
    Collection_End_Time_24hrs = as.numeric(gsub(":", "", Collection_End_Time_24hrs))
  ) %>% ##Format columns
    select(
      Project,
      Plot,
      Sample_ID,
      Sample_location,
      Total_Volume_mL,
      Vial_ID,
      Grid_B4, 
      Grid_C3, 
      Grid_C6, 
      Grid_D5, 
      Grid_E3, 
      Grid_F4, 
      Grid_F6, 
      Grid_H3, 
      Grid_H6, 
      Grid_I5, 
      Evacuation_date_YYYYMMDD, 
      Collection_Date_YYYYMMDD,
      Collection_Start_Time_24hrs, 
      Collection_End_Time_24hrs) 


```

## Pull in pre-treatment pooled metadata
```{r rest of metadata, include=FALSE}

##Convert to long format
raw_metadata_pt_pooled_long <- raw_metadata_pt %>%
  select("Plot", "Grid", "Depth_cm", "Bulk_mL_20220613", "Bulk_mL_20220615", "Notes") %>%
  mutate(
    Bulk_mL_20220613 = as.character(Bulk_mL_20220613), 
    Bulk_mL_20220615 = as.character(Bulk_mL_20220615) 
  ) %>%
  pivot_longer(
    cols = starts_with("Bulk_mL_"),
    names_to = "Collection_Date_YYYYMMDD",
    names_prefix = "Bulk_mL_",
    values_to = "Bulk_mL"
  ) %>% 
  mutate(Bulk_mL = as.numeric(Bulk_mL)) %>% 
  mutate(Bulk_mL = ifelse(is.na(Bulk_mL), 0, Bulk_mL)) ##Replace NA values with zeroes

##Pivot dataframe so have volume per plot and timepoint
raw_metadata_pooled_pt <- raw_metadata_pt_pooled_long %>%
    pivot_wider(
      id_cols = c(Plot, Collection_Date_YYYYMMDD), 
      names_from = Grid, 
      values_from = Bulk_mL, 
      names_prefix = "Grid_", 
    ) %>% 
    mutate(Total_Volume_mL = rowSums(across(c(Grid_B4, 
                                              Grid_C3, 
                                              Grid_C6, 
                                              Grid_D5, 
                                              Grid_E3, 
                                              Grid_F4, 
                                              Grid_F6, 
                                              Grid_H3, 
                                              Grid_H6, 
                                              Grid_I5)), na.rm = TRUE)) %>% 
    filter(Total_Volume_mL > 0)



##Create IDs
pooled_inventory_pt <- raw_metadata_pooled_pt %>%
  mutate(Plot = gsub("SEAWATER", "SALTWATER", Plot)) %>% ##Format Plot
  mutate(Plot_abb = case_when(
    Plot == "CONTROL" ~ "C",
    Plot == "FRESHWATER" ~ "FW",
    Plot == "SALTWATER" ~ "SW"
  )) %>% 
  mutate(Start_time = case_when(
    Collection_Date_YYYYMMDD == "20220613" ~ "1100",
    Collection_Date_YYYYMMDD == "20220615" ~ "1200")) %>% 
  mutate(End_time = case_when(
    Collection_Date_YYYYMMDD == "20220613" ~ "1330",
    Collection_Date_YYYYMMDD == "20220615" ~ NA)) %>% 
  mutate(Project = "COMPASS TEMPEST", 
         Plot = str_to_title(Plot), 
         Sample_location = as.character(NA), 
         Collection_Date_YYYYMMDD = as.numeric(Collection_Date_YYYYMMDD), 
         Evacuation_date_YYYYMMDD = as.numeric(Collection_Date_YYYYMMDD), 
         EST_EDT = "EDT") %>% 
  mutate(Sample_ID = paste("TMP",
                          Plot_abb,
                          "PW", 
                          "15cm", 
                          "POOLED", 
                          Collection_Date_YYYYMMDD, 
                          sep = "_")) %>% ##Create Sample IDs
  mutate(Vial_ID = paste(Plot_abb,
                          "15", 
                          Collection_Date_YYYYMMDD, 
                          sep = "_")) %>% ##Create Vial IDs
  mutate(
    Collection_Start_Time_24hrs = substr(Start_time, 1, 5), 
    Collection_Start_Time_24hrs = as.numeric(gsub(":", "", Collection_Start_Time_24hrs)), 
    Collection_End_Time_24hrs = substr(End_time, 1, 5), 
    Collection_End_Time_24hrs = as.numeric(gsub(":", "", Collection_End_Time_24hrs))
  ) %>% ##Format columns
    select(
      Project,
      Plot,
      Sample_ID,
      Sample_location,
      Total_Volume_mL,
      Vial_ID,
      Grid_B4, 
      Grid_C3, 
      Grid_C6, 
      Grid_D5, 
      Grid_E3, 
      Grid_F4, 
      Grid_F6, 
      Grid_H3, 
      Grid_H6, 
      Grid_I5, 
      Evacuation_date_YYYYMMDD, 
      Collection_Date_YYYYMMDD,
      Collection_Start_Time_24hrs, 
      Collection_End_Time_24hrs) 


```


## Combine event and pre-treatment pooled metadata
```{r}

pooled_inventory_all <- bind_rows(pooled_inventory_event, pooled_inventory_pt)

```


NOW COMBINE THE NEW INVENTORY WITH THE OLD
# Prep POOLED metadata to be appended to the inventory file 
```{r, process data file to match inventory}

#Read in the master inventory 
pooled_master <- read_excel(
  #path = inventory_path,
  "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
  sheet = "Porewater - Pooled",
  skip = 4, 
  col_types = "text"
)

#clean out any empty sample_ID rows
pooled_master <- pooled_master %>%
  filter(!is.na(Sample_ID))

#Change column types
pooled_master <- pooled_master %>%
  mutate(
    Total_Volume_mL = as.numeric(Total_Volume_mL), 
    Grid_B4 = as.numeric(Grid_B4), 
    Grid_C3 = as.numeric(Grid_C3), 
    Grid_C6 = as.numeric(Grid_C6), 
    Grid_D5 = as.numeric(Grid_D5), 
    Grid_E3 = as.numeric(Grid_E3), 
    Grid_F4 = as.numeric(Grid_F4), 
    Grid_F6 = as.numeric(Grid_F6), 
    Grid_H3 = as.numeric(Grid_H3), 
    Grid_H6 = as.numeric(Grid_H6), 
    Grid_I5 = as.numeric(Grid_I5),
    Evacuation_date_YYYYMMDD = as.numeric(Evacuation_date_YYYYMMDD), 
    Collection_Date_YYYYMMDD = as.numeric(Collection_Date_YYYYMMDD), 
    Collection_Start_Time_24hrs = as.numeric(Collection_Start_Time_24hrs), 
    Collection_End_Time_24hrs = as.numeric(Collection_End_Time_24hrs)
  )


# Align the columns
inventory_new <- pooled_inventory_all %>%
  select(any_of(names(pooled_master)))

missing_cols <- setdiff(names(pooled_master), names(inventory_new))
missing_cols

for (col in missing_cols) {
  inventory_new[[col]] <- NA
}

# Append to the inventory
inventory_new <- inventory_new %>%
  select(names(pooled_master))

pooled_inventory <- bind_rows(inventory_new, pooled_master)

#Set the header rows 
header_rows <- data.frame(
  col1 = c(
    "Project: COMPASS",
    "Site: TEMPEST",
    "Event: POOLED Porewater Inventory May 2022 - present",
    "" ))

```


# Replace the POOLED sheet in the inventory file
```{r, add the new inventory file}

# Load existing workbook
wb <- loadWorkbook("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx")

# Remove existing sheet if it exists
if ("Porewater - Pooled" %in% names(wb)) {
  removeWorksheet(wb, "Porewater - Pooled")
}

# Add replacement sheet
addWorksheet(wb, "Porewater - Pooled")

# Write header block
writeData(wb, "Porewater - Pooled", header_rows, startRow = 1, colNames = FALSE)

# Write inventory data below it
writeData(
  wb,
  "Porewater - Pooled",
  pooled_inventory,
  startRow = nrow(header_rows) + 1
)

# Write the header text into G4 (column 7, row 4)
writeData(
  wb,
  sheet = "Porewater - Pooled",
  x = "Breakdown_by_volume_mL",
  startCol = 7,
  startRow = 4,
  colNames = FALSE
)

# Merge G4:Q4 (columns 7-17)
mergeCells(
  wb,
  sheet = "Porewater - Pooled",
  cols = 7:16,
  rows = 4
)

##Bold the headers
header_style <- createStyle(
  textDecoration = "bold",
  halign = "center",
  valign = "center"
)
addStyle(
  wb,
  sheet = "Porewater - Pooled",
  style = header_style,
  rows = 1:5,
  cols = 1:21,
  gridExpand = TRUE
)

# Save changes back to the same file
saveWorkbook(wb, "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", overwrite = TRUE)

```


# Upload the new inventory files to Google Drive
```{r}

##inventory
drive_upload(
  media = "TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx",
  path = inventory_path, 
  name = "TEMPEST_PorewaterInventory_May2022_Present_updated",
  type = "spreadsheet"                    # Triggers conversion to Google Sheets
)

```
