---
title: "COMPASS_TEMPEST_Collate_All_Porewater_Data"
author: "Zoe Read"
date: "`r Sys.Date()`"
output: html_document
---

##Load required packages
```{r setup, include=FALSE}

library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(readxl)
library(tidyr)
library(tibble)
library(purrr)

#Packages for loading metadata
require(pacman)
pacman::p_load(janitor, # useful for simplifying column names
               googlesheets4, # read_sheet 
               googledrive, # drive_ functions
               plotrix,
               here) 

# detach("package:raster", unload = TRUE)

```

#Read in and collate all the collated analyte files
```{r}

##Should negative concentrations all be converted to zero? 

# Event samples: 
# 2022: T#
# 2023: AM/PM for PW, T# for Sourcewater
# 2023: T# and AM/PM for PW, AM/MID/PM for Sourcewater, High and Low Tide for Estuary Sourcewater (Removed AM/PM from PW ID, converted HT/LT to AM/PM for Estuary Sourcewater)

##Collated data by analyte
##Desired Sample ID format: TMP_FW_H6_15CM_20220620_T0

####DOC####
DOC <- read.csv("DOC/COMPASS_TEMPEST_DOC_AllData.csv")
DOC1 <- DOC %>% 
  ##Fix sample IDs
  mutate(Sample_ID = gsub("TMP_FW_E3_15_20220518", "TMP_FW_E3_15_20220520", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_SW_H6_15_20230417", "TMP_SW_H3_15_20230417", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_SW_H6_15_20230612", "TMP_SW_B4_15_20230612", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_C_C3_15_20230628", "TMP_C_B4_15_20230628", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_SW_C3_15_20260508 ", "TMP_SW_C3_30_20260508", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_SW_C6_15_20260508", "TMP_SW_C6_30_20260508", Sample_ID)) %>% 

  
  dplyr::select(c("Sample_ID", "npoc_mgL", "npoc_uM", "npoc_flag", "tdn_mgL", "tdn_uM", "tdn_flag")) %>% 
  mutate(Sample_ID = toupper (Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_15_", "_15CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_30_", "_30CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_50_", "_50CM_", Sample_ID)) %>% 
  mutate(Sample_ID = ifelse(str_detect(Sample_ID, "SOURCE"), gsub("20220622_T", "20220622_HR", Sample_ID), Sample_ID)) %>% 
  mutate(Sample_ID = gsub("RAIN_20241212", "RAINWATER_20241212", Sample_ID)) %>% 
  distinct() %>% 
  mutate(Sample_ID = gsub("TMP_SW_C3_15CM_20260508", "TMP_SW_C3_30CM_20260508", Sample_ID))
  

DOC2 <- DOC1 %>%
  mutate(Plot = stringr::str_extract(Sample_ID, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5") ,
         Sample_Type = stringr::str_extract(Sample_ID, "SOURCE|RAINWATER"), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         Date = stringr::str_extract(Sample_ID, "[0-9]{8}"),
         Depth = stringr::str_extract(Sample_ID, "15CM|30CM|50CM"),
         Depth = ifelse(str_detect(Sample_ID, "2022|2023|2024|2025|20260309|20260410") & Sample_Type == "PW", "15CM", Depth), ##Only 15 cm sampling until 2026 for PW
         Depth = ifelse(Sample_Type == "SOURCE|RAINWATER", NA, Depth), ##Only 15 cm sampling until 2026 for PW
         T_num = stringr::str_extract(Sample_ID, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID, "AM|MID|PM"), 
         TOD = case_when(
           Date == "20240612" & T_num == "T2" ~ "AM", 
           Date == "20240612" & T_num == "T3" ~ "PM", 
           Date == "20240613" & T_num == "T4" ~ "AM", 
           Date == "20240613" & T_num == "T5" ~ "PM", 
           TRUE ~ TOD), 

         Sample_ID_formatted = paste("TMP", Sample_Type, Plot, Grid, Depth, T_num, Date, TOD, sep = "_"), 
         Sample_ID_formatted = gsub("_NA", "", Sample_ID_formatted)) %>% 
  select(-Sample_ID)

####Nutrients####
Nutrients <- read.csv("Nutrients/COMPASS_TEMPEST_Nutrient_AllData.csv")
Nutrients1 <- Nutrients %>%
  ##Fix sample IDs
  mutate(Sample_ID = gsub("TMP_Source_Estuary_20240611_HT", "TMP_Source_Estuary_20240611_AM", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_Source_Estuary_20240611_LT", "TMP_Source_Estuary_20240611_PM", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_Source_Estuary_20240612_HT", "TMP_Source_Estuary_20240612_AM", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_Source_Estuary_20240612_LT", "TMP_Source_Estuary_20240612_PM", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_Source_Estuary_20240613_HT", "TMP_Source_Estuary_20240613_AM", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("TMP_Source_Estuary_20240613_LT", "TMP_Source_Estuary_20240613_PM", Sample_ID)) %>% 
  # mutate(Sample_ID = gsub("TMP_C_H6_20260602", "TMP_C_H6_20250602", Sample_ID)) %>% 

rename(
    NOx_mgL = NOx_Conc_mgL, 
    NOx_uM = NOx_Conc_uM, 
    NH3_mgL = NH3_Conc_mgL, 
    NH3_uM = NH3_Conc_uM, 
    PO4_mgL = PO4_Conc_mgL, 
    PO4_uM = PO4_Conc_uM
  ) %>%
  dplyr::select(c("Sample_ID", "NOx_mgL", "NOx_uM", "NOx_Conc_Flag", "NH3_mgL", "NH3_uM",  "NH3_Conc_Flag", "PO4_mgL", "PO4_uM", "PO4_Conc_Flag")) %>% 
  mutate(Sample_ID = toupper (Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_15_", "_15CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_30_", "_30CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_50_", "_50CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("RAINWATER_DEC10-12_2024", "RAINWATER_20241212", Sample_ID))  %>% 
  distinct() 


Nutrients2 <- Nutrients1 %>%
  mutate(Plot = stringr::str_extract(Sample_ID, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5") ,
         Sample_Type = stringr::str_extract(Sample_ID, "SOURCE|RAINWATER"), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         Date = stringr::str_extract(Sample_ID, "[0-9]{8}"),
         Depth = stringr::str_extract(Sample_ID, "15CM|30CM|50CM"),
         Depth = ifelse(str_detect(Sample_ID, "2022|2023|2024|2025|20260309|20260410") & Sample_Type == "PW", "15CM", Depth), ##Only 15 cm sampling until 2026 for PW
         Depth = ifelse(Sample_Type == "SOURCE|RAINWATER", NA, Depth), ##Only 15 cm sampling until 2026 for PW
         T_num = stringr::str_extract(Sample_ID, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID, "AM|MID|PM"), 
         TOD = case_when(
           Date == "20240612" & T_num == "T2" ~ "AM", 
           Date == "20240612" & T_num == "T3" ~ "PM", 
           Date == "20240613" & T_num == "T4" ~ "AM", 
           Date == "20240613" & T_num == "T5" ~ "PM", 
           TRUE ~ TOD), 
         
         Sample_ID_formatted = paste("TMP", Sample_Type, Plot, Grid, Depth, T_num, Date, TOD, sep = "_"), 
         Sample_ID_formatted = gsub("_NA", "", Sample_ID_formatted)) %>% 
  select(-Sample_ID)


####SO4/Cl####
SO4_Cl <- read.csv("Sulfate_Chloride/COMPASS_TEMPEST_SO4_Cl_AllData.csv")

SO4_Cl1 <- SO4_Cl %>%
  ##Fix sample IDs
  mutate(Sample_ID = gsub("TMP_SW_H6_20230612_15CM", "TMP_SW_B4_20230612_15CM", Sample_ID)) %>% 
  
  
  
  
  dplyr::select(c("Sample_ID", "SO4_ppm", "SO4_mM", "SO4_Conc_flag", "Cl_ppm", "Cl_mM", "Cl_Conc_flag", "salinity")) %>% 
  mutate(Sample_ID = toupper (Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_15_", "_15CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_30_", "_30CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_50_", "_50CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("RAINWATER_20241212", "RAINWATER_20241212", Sample_ID)) %>% 
  distinct() 


SO4_Cl2 <- SO4_Cl1 %>%
  mutate(Plot = stringr::str_extract(Sample_ID, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5") ,
         Sample_Type = stringr::str_extract(Sample_ID, "SOURCE|RAINWATER"), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         Date = stringr::str_extract(Sample_ID, "[0-9]{8}"),
         Depth = stringr::str_extract(Sample_ID, "15CM|30CM|50CM"),
         Depth = ifelse(str_detect(Sample_ID, "2022|2023|2024|2025|20260309|20260410") & Sample_Type == "PW", "15CM", Depth), ##Only 15 cm sampling until 2026 for PW
         Depth = ifelse(Sample_Type == "SOURCE|RAINWATER", NA, Depth), ##Only 15 cm sampling until 2026 for PW
         T_num = stringr::str_extract(Sample_ID, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID, "AM|MID|PM"), 
         TOD = case_when(
           Date == "20240612" & T_num == "T2" ~ "AM", 
           Date == "20240612" & T_num == "T3" ~ "PM", 
           Date == "20240613" & T_num == "T4" ~ "AM", 
           Date == "20240613" & T_num == "T5" ~ "PM", 
           TRUE ~ TOD), 
         
         Sample_ID_formatted = paste("TMP", Sample_Type, Plot, Grid, Depth, T_num, Date, TOD, sep = "_"), 
         Sample_ID_formatted = gsub("_NA", "", Sample_ID_formatted)) %>% 
  select(-Sample_ID)


####Sulfide####
Sulfide <- read.csv("Sulfide/COMPASS_TEMPEST_Sulfide_AllData.csv")
Sulfide1 <- Sulfide %>%
  dplyr::select(c("Sample_ID", "H2S_mean_uM", "H2S_sd_uM", "H2S_cv", "H2S_Conc_flag", "H2S_cv_flag")) %>% 
  mutate(Sample_ID = toupper (Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_15_", "_15CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_30_", "_30CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_50_", "_50CM_", Sample_ID)) %>% 
  mutate(Sample_ID = ifelse(str_detect(Sample_ID, "SOURCE|RAINWATER"), gsub("_15CM", "", Sample_ID), Sample_ID)) %>% 
  distinct() 


Sulfide2 <- Sulfide1 %>%
  mutate(Plot = stringr::str_extract(Sample_ID, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5") ,
         Sample_Type = stringr::str_extract(Sample_ID, "SOURCE|RAINWATER"), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         Date = stringr::str_extract(Sample_ID, "[0-9]{8}"),
         Depth = stringr::str_extract(Sample_ID, "15CM|30CM|50CM"),
         Depth = ifelse(str_detect(Sample_ID, "2022|2023|2024|2025|20260309|20260410") & Sample_Type == "PW", "15CM", Depth), ##Only 15 cm sampling until 2026 for PW
         T_num = stringr::str_extract(Sample_ID, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID, "AM|MID|PM"), 
         TOD = case_when(
           Date == "20240612" & T_num == "T2" ~ "AM", 
           Date == "20240612" & T_num == "T3" ~ "PM", 
           Date == "20240613" & T_num == "T4" ~ "AM", 
           Date == "20240613" & T_num == "T5" ~ "PM", 
           TRUE ~ TOD), 
         
         Sample_ID_formatted = paste("TMP", Sample_Type, Plot, Grid, Depth, T_num, Date, TOD, sep = "_"), 
         Sample_ID_formatted = gsub("_NA", "", Sample_ID_formatted)) %>% 
  select(-Sample_ID)

####pH####
pH <- read.csv("pH/COMPASS_TEMPEST_pH_AllData.csv")
colnames(pH)
pH <- pH %>%
  mutate(Sample_ID = toupper(Sample_ID)) %>% 
  dplyr::select(c("Sample_ID", "pH", "collection_temp_C", "temperature_flag", "pH_flag")) %>% 
  distinct() %>% 
  filter(!(Sample_ID == "TMP_SW_SOURCEWATER_20230607_1255_15CM" & temperature_flag == "temperature not recorded")) %>% 
  filter(!(Sample_ID == "TMP_ESTUARY_SOURCEWATER_20230607_1500_15CM" & temperature_flag == "temperature not recorded")) %>% 
  filter(!(Sample_ID == "TMP_FW_SOURCEWATER_20230607_1519_15CM" & temperature_flag == "temperature not recorded"))



```

##Combine data
```{r}
df_list <- list(DOC2, Nutrients2, SO4_Cl2, Sulfide2)
data <- df_list %>% reduce(full_join)
```

##Make table of sample IDs
```{r}
sampleID_format <- data %>% 
  mutate(Year = substr(Date, 1, 4)) %>% 
  select(c(Sample_Type, Year, T_num, TOD)) %>% 
  distinct()

write.csv(sampleID_format, "Sample_ID_formatting.csv")
```

##Need to figure out duplicates
```{r}
##Need to figure out duplicates
df_duplicates <- data %>%
  group_by(Sample_ID_formatted) %>%
  filter(n() > 1) %>%
  ungroup()

##Fix SO4/Cl duplicates
df_duplicates_SO4 <- SO4_Cl2 %>%
  group_by(Sample_ID_formatted) %>%
  filter(n() > 1) %>%
  ungroup()
# lots of duplicate SO4_Cl rows removed - should do this in collated SO4_Cl data
# 2024 Estuary sourcewater was incorrectly all labelled AM, some should be PM - DONE
## Some samples were ran twice in COMPASS_TEMPEST_20230603 and 20230605 SO4/Cl run, sometimes the values are similar and sometimes very different. Not sure what happened here. - used the first sample
## TMP_PW_SW_F4_15CM_20241213 ran twice, different Cl values - removed both

##Fix DOC duplicates
df_duplicates_DOC <- DOC2 %>%
  group_by(Sample_ID_formatted) %>%
  filter(n() > 1) %>%
  ungroup()
## 20230607_PM samples ran twice, sometimes the values are similar and sometimes very different. - AM samples were coded as PM samples, fixed the code for this. 
## TMP_PW_SW_C3_15CM_T3_20240612 ran twice, different values - removed


##Fix Nutrient duplicates
df_duplicates_Nutrients <- Nutrients2 %>%
  group_by(Sample_ID_formatted) %>%
  filter(n() > 1) %>%
  ungroup()
# 202300606 entered incorrectly, fixed 
# TEMPEST spelled incorrectly, fixed 
## TMP_C_C6_20230515 has zeroes for NOx in one run but not the other, the zero run was second but doesn't say it was a dilution, not sure what happened here. 
## Estuary sourcewater 20240611, 20240612, 20240613 ran twice, NOx was generally lower in one run

##Fix Sulfide duplicates
df_duplicates_Sulfide <- Sulfide2 %>%
  group_by(Sample_ID_formatted) %>%
  filter(n() > 1) %>%
  ungroup()


##Fix pH duplicates
df_duplicates_pH <- pH %>%
  group_by(Sample_ID) %>%
  filter(n() > 1) %>%
  ungroup()
##lots of duplicate pH rows removed - should do this in collated pH data
# TMP_SW_SOURCEWATER_20230607_1255_15CM has two entries? One says taken with YSI, no temperature - removed 
# TMP_FW_SOURCEWATER_20230607_1519_15CM has two entries? One has no temperature - removed
# TMP_ESTUARY_SOURCEWATER_20230607_1500_15CM has two entries? One has no temperature - removed 

```


## Pull in active porewater tracking inventory sheet from Google Drive: 
```{r rest of metadata, include=FALSE}

# #Run these to pull the TEMPEST porewater metadata into GitHub if not already there
# inventory_directory <- "https://docs.google.com/spreadsheets/d/1J4aPIRrKrPi_o8c7WzkY6cenWtoK8UgOrzIR5GiRBgg/edit?gid=1169944999#gid=1169944999"
# 
# drive_download(inventory_directory, overwrite = TRUE)

# link directory to google drive location
directory <- "https://drive.google.com/drive/folders/1oEeiU48OULo0UOu0np3l1JhYYEQcnw7e"

# check list of files found at the link
drive_ls(directory)

# filter to necessary files
# updated metadata event
files_pulse <- drive_ls(directory) %>%
  filter(grepl("TEMPEST_PorewaterInventory_May2022_Present_updated", name))

# download files to local drive
lapply(c(files_pulse$name), drive_download, 
       overwrite = TRUE)


sheet_names <- excel_sheets("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx")
sheet_names

raw_metadata_lys <- read_excel("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", sheet = "Porewater - Individual", skip = 4)

raw_metadata_sw <- read_excel("TEMPEST_PorewaterInventory_May2022_Present_updated.xlsx", sheet = "Source Water", skip = 4)

```


## Create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#Create IDs from what was collected for comparison later
##Lysimeter
raw_metadata_lys1 <- raw_metadata_lys %>%
  mutate(Sample_ID_meta = Sample_ID) %>% 
  mutate(Sample_ID = toupper (Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_15_", "_15CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_30_", "_30CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_50_", "_50CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("1OF1", "", Sample_ID)) %>% ##There were 2-3 vials collected at the same time, so this sample ID doesn't actually give more information. 
  mutate(Sample_ID = gsub("1OF2", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("1OF3", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("2OF2", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("2OF3", "", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("3OF3", "", Sample_ID)) 

raw_metadata_lys2 <- raw_metadata_lys1 %>%
  mutate(Plot = stringr::str_extract(Sample_ID, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5") ,
         Date = stringr::str_extract(Sample_ID, "[0-9]{8}"),
         Depth = stringr::str_extract(Sample_ID, "15CM|30CM|50CM"),
         T_num = stringr::str_extract(Sample_ID, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID, "AM|MID|PM"), 
         TOD = case_when(
           Date == "20240612" & T_num == "T2" ~ "AM", 
           Date == "20240612" & T_num == "T3" ~ "PM", 
           Date == "20240613" & T_num == "T4" ~ "AM", 
           Date == "20240613" & T_num == "T5" ~ "PM", 
           TRUE ~ TOD), 
         Sample_Type = stringr::str_extract(Sample_ID, "SOURCE_RAINWATER|SOURCE"), 
         Sample_Type = gsub("SOURCE_RAINWATER", "RAINWATER", Sample_Type), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         
         Sample_ID_formatted = paste("TMP", Sample_Type, Plot, Grid, Depth, T_num, Date, TOD, sep = "_"), 
         Sample_ID_formatted = gsub("_NA", "", Sample_ID_formatted)) %>% 
  select(-Sample_ID) %>% 
  distinct(Sample_ID_formatted, .keep_all = TRUE) %>% 
  select(c("Evacuation_date_YYYYMMDD", "Collection_Date_YYYYMMDD", "Collection_Start_Time_24hrs", "Collection_End_Time_24hrs", "EST_EDT", "Notes", "Sample_ID_formatted", "Sample_ID_meta"))


##Source Water
raw_metadata_sw1 <- raw_metadata_sw %>%
  mutate(Sample_ID_meta = Sample_ID) %>% 
  mutate(Sample_ID = toupper (Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_15_", "_15CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_30_", "_30CM_", Sample_ID)) %>% 
  mutate(Sample_ID = gsub("_50_", "_50CM_", Sample_ID)) %>% 
  mutate(Collection_Date_YYYYMMDD = ifelse(str_detect(Sample_ID, "20220622"), 20220622, Collection_Date_YYYYMMDD)) %>% 
  mutate(Collection_Time_24hrs = ifelse(Sample_ID == "TMP_Source_FW_HR0_20220622_0654", 0654, Collection_Time_24hrs)) %>% ##Fix 2022 sourcewater dates and times
  mutate(Collection_Time_24hrs = ifelse(Sample_ID == "TMP_Source_FW_HR0_20220622_0645", 0645, Collection_Time_24hrs)) %>%
  mutate(Collection_Time_24hrs = ifelse(Sample_ID == "TMP_Source_FW_HR0_20220622_0800", 0800, Collection_Time_24hrs)) %>%
  mutate(Collection_Time_24hrs = ifelse(Sample_ID == "TMP_Source_FW_HR0_20220622_0930", 0930, Collection_Time_24hrs)) %>%
  mutate(Collection_Time_24hrs = ifelse(Sample_ID == "TMP_Source_FW_HR0_20220622_1000", 1000, Collection_Time_24hrs)) %>%
  mutate(Collection_Time_24hrs = ifelse(Sample_ID == "TMP_Source_FW_HR0_20220622_1100", 1100, Collection_Time_24hrs)) %>%
  mutate(Collection_Time_24hrs = ifelse(Sample_ID == "TMP_Source_FW_HR0_20220622_1200", 1200, Collection_Time_24hrs)) %>%
  mutate(Collection_Time_24hrs = ifelse(Sample_ID == "TMP_Source_FW_HR0_20220622_1448", 1448, Collection_Time_24hrs)) %>%
  mutate(Sample_ID = gsub("RAIN_20241212", "RAINWATER_20241212", Sample_ID))

raw_metadata_sw2 <- raw_metadata_sw1 %>%
  mutate(Plot = stringr::str_extract(Sample_ID, 'SW|FW|_C_|ESTUARY'),
         Plot = gsub("_", "", Plot), 
         Grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5") ,
         Date = stringr::str_extract(Sample_ID, "[0-9]{8}"),
         Depth = stringr::str_extract(Sample_ID, "15CM|30CM|50CM"),
         T_num = stringr::str_extract(Sample_ID, "T0|T1|T2|T3|T4|T5|T6|HR0|HR1|HR2|HR3|HR4|HR5|HR6|HR7"),
         TOD = stringr::str_extract(Sample_ID, "AM|MID|PM"), 
         Sample_Type = stringr::str_extract(Sample_ID, "SOURCE_RAINWATER|SOURCE"), 
         Sample_Type = gsub("SOURCE_RAINWATER", "RAINWATER", Sample_Type), 
         Sample_Type = replace_na(Sample_Type, "PW"), 
         
         Sample_ID_formatted = paste("TMP", Sample_Type, Plot, Grid, Depth, T_num, Date, TOD, sep = "_"), 
         Sample_ID_formatted = gsub("_NA", "", Sample_ID_formatted)) %>% 
  select(-Sample_ID) %>% 
  distinct(Sample_ID_formatted, .keep_all = TRUE) %>% 
  select(c("Deployment_Date_YYYYMMDD", "Deployment_Time_24hrs" , "Collection_Date_YYYYMMDD", "Collection_Time_24hrs",  "EST_EDT", "Notes", "Sample_ID_formatted", "Sample_ID_meta"))


##Combine lys and sw data
metadata <- bind_rows(raw_metadata_lys2, raw_metadata_sw2)

```

## Check to see if samples run match metadata & merge info
```{r check sample ids with metadata, echo=FALSE, warning = FALSE}

#check to see if all samples are present in the metadata 
data$Sample_ID_formatted <- toupper(data$Sample_ID_formatted)
metadata$Sample_ID_formatted <- toupper(metadata$Sample_ID_formatted)

all_present <- all(data$Sample_ID_formatted %in% metadata$Sample_ID_formatted)

if (all_present) {
  message("All sample IDs are present in metadata.")
} else {
  message("Some sample IDs are missing from metadata.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(data$Sample_ID_formatted, metadata$Sample_ID_formatted)
  print(missing_ids)
}

#merge metadata with sample run data 
merged_data <- data %>%
  left_join(metadata, by = c("Sample_ID_formatted"))


##Samples missing from metadata: 

# 2022 estuary sourcewater samples ("TMP_SOURCE_ESTUARY_HR0_20220622", "TMP_SOURCE_ESTUARY_HR1_20220622") - not entered in metadata
# TMP_PW_SW_H3_15CM_20220615 - not entered in metadata, can't find scanned datasheets for these
# TMP_PW_C_F6_15CM_20220615 - not entered in metadata, can't find scanned datasheets for these
# TMP_PW_FW_F6_15CM_20230606 - scanned datasheet says no water collected but have SO4/Cl data? Not missing any of the other SW samples

# 2024 sourcewater samples don't match metadata
data_sel <- data %>% 
  filter(Sample_ID_formatted == "TMP_PW_SW_C6_15CM_20231211")

```

## Round data to sensitivity of the instrument 
```{r}

colnames(merged_data)
##Based on raw data from the instrument: 
#SO4/Cl/Salinity - 4 decimals 
#H2S/Fe - 3 decimals
#Nutrients - 6 decimals 
#DOC - ? 

final_data1 <- merged_data %>% 
  mutate(SO4_ppm = round(as.numeric(SO4_ppm), digits = 4), 
         SO4_mM = round(as.numeric(SO4_mM), digits = 4), 
         Cl_ppm = round(as.numeric(Cl_ppm), digits = 4), 
         Cl_mM = round(as.numeric(Cl_mM), digits = 4), 
         salinity = round(as.numeric(salinity), digits = 4), 
         H2S_mean_uM = round(as.numeric(H2S_mean_uM), digits = 3), 
         H2S_sd_uM = round(as.numeric(H2S_sd_uM), digits = 3), 
         NH3_mgL = round(as.numeric(NH3_mgL), digits = 6), 
         NH3_uM = round(as.numeric(NH3_uM), digits = 6), 
         PO4_mgL = round(as.numeric(PO4_mgL), digits = 6), 
         PO4_uM = round(as.numeric(PO4_uM), digits = 6), 
         NOx_mgL = round(as.numeric(NOx_mgL), digits = 6), 
         NOx_uM = round(as.numeric(NOx_uM), digits = 6))
```

## Format and write the alldata file out to the folder 
```{r, echo=FALSE}

##Order data by Year, Month, Day
# final_data1 <- final_data[order(final_data$Year, final_data$Month, final_data$Day, final_data$Zone, final_data$Lysimeter, final_data$Depth_cm),]

colnames(final_data1)

final_data2 <- final_data1 %>% 
  mutate(
    Year = as.numeric(substr(Date, 1, 4)), 
    Month = as.numeric(substr(Date, 5, 6)), 
    Day = as.numeric(substr(Date, 7, 8)), 
    date = as.numeric(Date), 
    Project = "COMPASS", 
    Region = "CB", 
    Site = "TMP", 
    Depth_cm = gsub("CM", "", Depth), 
    Timepoint = paste(T_num, TOD, sep = "_"), 
    Timepoint = gsub("_NA", "", Timepoint), 
    Timepoint = gsub("NA_", "", Timepoint), 
    ) %>% 
  ##Code in EDT/EST
  mutate(EST_EDT = case_when(
    Date >= 20220313 & Date <  20221106 ~ "EDT", 
    Date >= 20221106 & Date <  20230312 ~ "EST", 
    Date >= 20230312 & Date <  20231105 ~ "EDT", 
    Date >= 20231105 & Date <  20240310 ~ "EST", 
    Date >= 20240310 & Date <  20241103 ~ "EDT", 
    Date >= 20241103 & Date <  20250309 ~ "EST", 
    Date >= 20250309 & Date <  20251102 ~ "EDT", 
    Date >= 20251102 & Date <  20260308 ~ "EST",    
    Date >= 20260308 & Date <  20261101 ~ "EDT", 
    TRUE ~ EST_EDT
  ))
  
  
  
  rename(Sample_ID = Sample_ID_formatted, 
         NOx_flag = NOx_Conc_Flag, 
         NH3_flag = NH3_Conc_Flag, 
         PO4_flag = PO4_Conc_Flag, 
         SO4_flag = SO4_Conc_flag, 
         Cl_flag = Cl_Conc_flag, 
         H2S_flag = H2S_Conc_flag
         ) %>% 
  select(c("Project", "Region", "Year", "Month", "Day", "Site", "Sample_ID", "Plot", "Grid", "Date", "Depth_cm", "T_num", "TOD", "Sample_Type", 
           "Evacuation_date_YYYYMMDD", "Collection_Date_YYYYMMDD", "Collection_Start_Time_24hrs", "Collection_End_Time_24hrs", 
           "Deployment_Date_YYYYMMDD", "Deployment_Time_24hrs" , "Collection_Time_24hrs", "EST_EDT", "Notes", 
           "npoc_mgL", "npoc_uM", "npoc_flag", "tdn_mgL", "tdn_uM", "tdn_flag", 
           "NOx_mgL", "NOx_uM", "NOx_flag", "NH3_mgL", "NH3_uM", "NH3_flag", "PO4_mgL", "PO4_uM", "PO4_flag", 
           "SO4_ppm", "SO4_mM", "SO4_flag", "Cl_ppm", "Cl_mM", "Cl_flag", "salinity", 
           "H2S_mean_uM", "H2S_sd_uM", "H2S_cv", "H2S_flag"))


write.csv(final_data1, "COMPASS_TEMPEST_AllData_AllYears.csv", row.names = FALSE)

```

