---
title: "TEMPEST: Porewater CDOM"
author: "2024-2025 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true

---

##This code pulls in CDOM Abs and EEM processed data files from Jieun's folders on the Google Drive and renames them based on Jieun's Sample Log also from the Google Drive so that the names correspond to sample IDs. 
##In additiom, this outputs metadata for each sample ID, including the integration time, dilution, RSU_area_1s columns from the sample log. 

# Need to double check that these sample ID's match all the ones in the collated data for all other analytes and check against metadata/flag data that Jieun said is missing 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F)

# load packages
require(pacman)
pacman::p_load(tidyverse, # keep things tidy
               janitor, # useful for simplifying column names
               googlesheets4, # read_sheet 
               googledrive, # drive_ functions
               plotrix,
               here, 
               readxl, 
               readr, 
               purrr) 

#Try to deal with the timezone demons 
common_tz = "Etc/GMT+5"

Sys.setenv(TZ = "America/New_York")

# Authenticate with your Google account
drive_auth()

```

# Bring in the sample log:
```{r read in sample log, include=FALSE}

#Run these to pull the CDOM sample log into GitHub if not already there
inventory_directory <- "https://docs.google.com/spreadsheets/d/1ACmmjEuZ04tj0_gnS4SJ170UfjVXtvjB/edit?gid=1419188231#gid=1419188231"

drive_download(inventory_directory, overwrite = TRUE)

sheet_names <- excel_sheets("JK_CDOM_sample_log_TMP3_Jun24-25.xlsx")
sheet_names

sample_log <- read_excel("JK_CDOM_sample_log_TMP3_Jun24-25.xlsx", sheet = "log")
colnames(sample_log)
sample_log <- sample_log %>% 
  mutate(abs_inventory_vial_id = paste0(inventory_vial_id, "_ABS_DilCorr.csv")) %>% 
  mutate(eem_inventory_vial_id = paste0(inventory_vial_id, "_EEM_DilCorr_IFE_RamNorm.csv"))

```


##Pull processed absorbance data from google drive and rename files to save into github 
```{r read in and rename processed absorbance data}
##Processed absorbance data: https://drive.google.com/drive/folders/18gq7wUsBCfbxl4vDXvInE00T_AWML_rB

##Files in EEM folder
abs_folder_path <- "Abs_Data"
dir.create(abs_folder_path)

#folder link to id
abs_folder = "https://drive.google.com/drive/folders/18gq7wUsBCfbxl4vDXvInE00T_AWML_rB"
folder_id = drive_get(as_id(abs_folder))

#find files in folder
files = drive_ls(folder_id)

#loop dirs and download files inside them
for (i in seq_along(files$name)) {
  #list files
  i_dir = drive_ls(files[i, ])
  
  #download files
  for (file_i in seq_along(i_dir$name)) {
    #fails if already exists
      drive_download(
        as_id(i_dir$id[file_i]),
        path = paste0("Abs_Data/", i_dir$name[file_i]), 
        overwrite = TRUE
      )
  }
}

##Rename files
for(i in 1:nrow(sample_log)){
    
    # Create file paths with folder and file name
    old_file <- file.path(abs_folder_path, sample_log$abs_processed_file_name_in_gdrive[i])
    new_file <- file.path(abs_folder_path, sample_log$abs_inventory_vial_id[i])
    
    # Check if the original file actually exists
  if (file.exists(old_file)) {
    
    # Rename the file
    file.rename(from = old_file, to = new_file)
    
    cat("Success: Renamed", sample_log$abs_processed_file_name_in_gdrive[i], "to", sample_log$abs_inventory_vial_id[i], "\n")
  } else {
    cat("Warning: File", sample_log$abs_processed_file_name_in_gdrive[i], "not found.\n")
  }
}
```


##Pull processed EEM data from google drive and rename files to save into github 
```{r read in and rename processed eem data}
##Processed EEM data: https://drive.google.com/drive/folders/1PISZZLcDSNMOLkQKUy7MLEI18pLTHCn0

##Files in EEM folder
eem_folder_path <- "EEM_Data"
dir.create(eem_folder_path)

#folder link to id
eem_folder = "https://drive.google.com/drive/folders/1PISZZLcDSNMOLkQKUy7MLEI18pLTHCn0"
eem_folder_id = drive_get(as_id(eem_folder))

#find files in folder
eem_files = drive_ls(eem_folder_id)

#loop dirs and download files inside them
for (i in seq_along(eem_files$name)) {
  #list files
  eem_i_dir = drive_ls(eem_files[i, ])
  
  #download files
  for (file_i in seq_along(eem_i_dir$name)) {
    #fails if already exists
      drive_download(
        as_id(eem_i_dir$id[file_i]),
        path = paste0("EEM_Data/", eem_i_dir$name[file_i]), 
        overwrite = TRUE
      )
  }
}

for(i in 1:nrow(sample_log)){
    
    # Create file paths with folder and file name
    eem_old_file <- file.path(eem_folder_path, sample_log$EEM_processed_file_in_gdrive[i])
    eem_new_file <- file.path(eem_folder_path, sample_log$eem_inventory_vial_id[i])
    
    # Check if the original file actually exists
  if (file.exists(eem_old_file)) {
    
    # Rename the file
    file.rename(from = eem_old_file, to = eem_new_file)
    
    cat("Success: Renamed", sample_log$EEM_processed_file_in_gdrive[i], "to", 
        sample_log$eem_inventory_vial_id[i], "\n")
  } else {
    cat("Warning: File", sample_log$EEM_processed_file_in_gdrive[i], "not found.\n")
  }
}
```


##Create the metadata
```{r output metadata}
colnames(sample_log)
metadata <- sample_log %>% 
  rename(abs_file_id = abs_inventory_vial_id, 
         eem_file_id = eem_inventory_vial_id) %>% 
  select(c(inventory_vial_id, integration_time_s, dilution, RSU_area_1s, abs_file_id, eem_file_id))

write.csv(metadata, "CDOM_metadata.csv")

```

