---
title: "TEMPEST: Porewater Sulfide"
author: "2023-2025 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true

output_dir: "To Be Reviewed/PDF"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(readxl)
library(tidyr)
library(tibble)
library(openxlsx)
library(knitr)
```

##Grab 2023 and 2024 and 2025 All Data Files
```{r Pull in 2022-2024 data files, message=FALSE, include=FALSE}

#read in 2023 Sulfide Data:
dat23 <- read.csv("2023/TEMPEST_Sulfide_2023.csv")
colnames(dat23)

#read in 2024 Sulfide Data:
dat24 <- read.csv("2024/TEMPEST_Sulfide_2024.csv")
colnames(dat24)

#read in 2025 Sulfide Data:
dat25 <- read.csv("2025/TEMPEST_Sulfide_2025.csv")
# head(dat25)
colnames(dat25)

raw_dat <- bind_rows(dat23, dat24, dat25)

colnames(raw_dat)

##Convert Collection Date into Date format
raw_dat$Collection_Date <- as.character(raw_dat$Collection_Date)
raw_dat$Collection_Date_formatted <- as.Date(raw_dat$Collection_Date, format = "%Y%m%d")

##Extract Year and month from Collection Date
raw_dat$Year <- format(raw_dat$Collection_Date_formatted, format = "%Y")
raw_dat$Month <- format(raw_dat$Collection_Date_formatted, format = "%m")
raw_dat$Day <- format(raw_dat$Collection_Date_formatted, format = "%d")

raw_dat <- raw_dat %>% 
  mutate(Grid = ifelse(str_detect(Sample_ID, "SOURCE"), "Sourcewater", Grid))

all_dat <- raw_dat %>%
  mutate(
    Project = "COMPASS", 
    Site = "TEMPEST", 
    Depth_cm = gsub("cm", "", Depth), 
    Plot = Zone
  ) %>%
  rename(
    Collection_Date_YYYYMMDD = Collection_Date, 
    Evacuation_Date_YYYYMMDD = Evacuation_date_YYYMMDD
  )%>%
  dplyr::select(
         Project, Site, Year, Month, Day, Plot, Grid, Depth_cm, Event_Time, Time_of_day, Sample_ID, 
         Volume_mL, Evacuation_Date_YYYYMMDD, Collection_Date_YYYYMMDD, 
         Deployment_Date_YYYYMMDD, Deployment_Time_24hrs, Collection_Time_24hrs, 
         Collection_Start_Time_24hrs, Collection_End_Time_24hrs, EST_EDT, 
         H2S_mean_uM, H2S_sd_uM, H2S_cv, H2S_Conc_flag, H2S_cv_flag, H2S_QAQC_flag, 
         Analysis_rundate,  Run_notes, Field_notes
        # list columns in the order you want them
  )


##Set 4 decimal points max
all_dat <- all_dat %>%
  mutate(across(where(is.numeric), round, digits = 4)) ##This does make all salinities 0 when Cl_ppm is 0 instead of 0.000026

```


##Make Relevant Metadata Sheet
```{r make metadata, include = FALSE}

# colnames(all_dat)
# 
# #pull the column names from the all_dat dataframe and create a new dataframe with them 
# metadat <- data.frame(Column_Names = colnames(all_dat))
# metadat
# 
# metadat$Description <- ""
# metadat$Units <- ""
# metadat$Instrument_ifapplicable <- ""
# 
# #add information about each line: 
# 
# metadat$Description <- c("Project name", 
#                          "Site where sample collected", 
#                          "Plot where sample collected (SW = Saltwater, FW = Freshwater, C = 
#                          Control, Estuary (Sourcewater))", 
#                          "Grid where sample collected (Letter + Number)", 
#                          "Depth where sample collected", 
#                          "Event sampling time (T1 through T6)", 
#                          "Time of day of sample collection (AM, PM, Midday)",
#                          "Sample Identifier", 
#                          "Volume of sample collected (mL)", 
#                          "Date lysimeter evacuated", 
#                          "Date sample collected", 
#                          "Time lysimeter evacuated", 
#                          "Time sample collected", 
#                          "Time zone of sample collection", 
#                          "Sulfate concentration in ppm", 
#                          "Sulfate concentration in mM", 
#                          "Sulfate concentration flag base on standard curve (bdl=below detection limit (0), adl=above detection limit (highest   
#                          standard), Within_Range=within detection range)", 
#                          "Sulfate QAQC Flags generated during checks of standards, blanks, dups, and spks",
#                          "Chloride concentration in ppm", 
#                          "Chloride concentration in mM", 
#                          "Chloride concentration flag base on standard curve (bdl=below detection limit (0), adl=above detection limit (highest   
#                          standard), Within_Range=within detection range)", 
#                          "Chloride QAQC Flags generated during checks of standards, blanks, dups, and spks", 
#                          "Salinity calculated using the Knudsen equation: Salinity=((1.807*(Chloride ppm)+0.026)/1000", 
#                          "The date that samples were run on the Sulfide IC", 
#                          "Any notes generated during QAQC", 
#                          "Notes generated during field sample collection")
# 
# metadat$Units <-       c("NA", 
#                          "NA", 
#                          "NA",
#                          "NA",
#                          "NA", 
#                          "NA",
#                          "NA",
#                          "NA", 
#                          "mL", 
#                          "NA",
#                          "NA",
#                          "NA",
#                          "NA",
#                          "NA",
#                          "ppm", 
#                          "mM", 
#                          "NA",
#                          "NA", 
#                          "ppm", 
#                          "mM", 
#                          "NA", 
#                          "NA", 
#                          "ppt", 
#                          "NA", 
#                          "NA",
#                          "NA")
# 
# metadat$Instrument_ifapplicable <-  
#   
#                        c("NA", 
#                          "NA",
#                          "NA", 
#                          "NA", 
#                          "NA",
#                          "NA", 
#                          "NA",
#                          "NA",
#                          "NA", 
#                          "NA",
#                          "NA",
#                          "NA", 
#                          "NA",
#                          "NA",
#                          "Sulfide Integrion HPIC (Thermo Fisher)", 
#                          "Sulfide Integrion HPIC (Thermo Fisher)", 
#                          "NA", 
#                          "NA", 
#                          "Sulfide Integrion HPIC (Thermo Fisher)", 
#                          "Sulfide Integrion HPIC (Thermo Fisher)", 
#                          "NA", 
#                          "NA", 
#                          "Sulfide Integrion HPIC (Thermo Fisher)", 
#                          "NA", 
#                          "NA",
#                          "NA")


```

##Write out files
```{r write out combined data and metadata, include=FALSE}

#write out a csv of all the data to the main folder: 
write.csv(all_dat, "COMPASS_TEMPEST_Sulfide_AllData.csv")


# #write out a csv of the metadata associated with the data: 
# write.csv(metadat, "COMPASS_TEMPEST_Sulfide_Metadata.csv")

```

## Visualize Porewater Data by Plot   
```{r Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Remove Sourcewater
data_plotting <- all_dat %>%
  filter(!str_detect(Grid, "Sourcewater"), ignore.case = TRUE)%>%
  filter(!str_detect(Grid, "Rainwater"), ignore.case = TRUE)

data_plotting <- data_plotting %>% 
    mutate(
    plot_1 = case_when(
      Plot == "C" ~ "Control",
      Plot == "FW" ~ "Freshwater",
      Plot == "SW" ~ "Saltwater",
      TRUE ~ Plot  # keep everything else unchanged
    )
  ) %>%
  mutate(plot_1 = factor(plot_1, levels = c("Control", "Freshwater", "Saltwater")))

#Plot data by year and change colors based on Plot:
viz_H2S_plot <- ggplot(data_plotting, aes(x = as.character(Year), y = H2S_mean_uM, fill = plot_1)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ plot_1, scales="free") +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfide (uM)", title = "TEMPEST Porewater: Sulfide") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))+ 
  theme(legend.position = "none") + 
  scale_x_discrete(drop = TRUE)+
  geom_text(aes(label = ifelse(H2S_mean_uM == 0, "0", "")), vjust = -0.5) 

viz_H2S_plot


```

## Summarized Porewater data for Plot and Collection Date
```{r Summarize and Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting$Collection_Date_formatted <- as.Date(as.character(data_plotting$Collection_Date_YYYYMMDD), format = "%Y%m%d")

##Summarize SO4 and Cl by Date and Zone
Sulfide_avg <- data_plotting %>%
  group_by(Collection_Date_formatted, plot_1) %>%
  dplyr::summarize(mean_H2S = mean(H2S_mean_uM, na.rm=TRUE), 
                   sd_H2S = sd(H2S_mean_uM, na.rm=TRUE))

# Define the specific date breaks you want to display
specific_breaks <- as.Date(c("2023-01-01", "2024-01-01", "2025-01-01"))

viz_H2s_plot_avgs <- ggplot(Sulfide_avg, aes(x = Collection_Date_formatted, y = mean_H2S, color = plot_1)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_H2S - sd_H2S, ymax = mean_H2S + sd_H2S), width = 0.2) + 
  geom_line() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = "Date", y = "Sulfide (uM)", title = "TEMPEST Porewater: Sulfide") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") + # Format x-axis labels
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

viz_H2s_plot_avgs


```

## Boxplot summary for Porewater
```{r Porewater Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Boxplot Summarize SO4 and Cl by Date, Site, and Zone
viz_H2S_boxplot <- ggplot(data_plotting, aes(x = Collection_Date_YYYYMMDD,  y = H2S_mean_uM, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  # scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) +
  theme_classic() +
  labs(x = " ", y = "Sulfide (uM)", title = "TEMPEST Porewater: Sulfide") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_H2S_boxplot

# # Save as a PNG
# png("COMPASS_SynopticCB_Cl_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_cl_boxplot)  # For ggplot objects, use print()
# dev.off() 
# 
# # Save as a PNG
# png("COMPASS_SynopticCB_SO4_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_so4_boxplot)  # For ggplot objects, use print()
# dev.off() 

```

## Boxplot summary for Sourcewater
```{r Sourcewater Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Only Sourcewater
data_plotting_source <- all_dat %>%
  filter(str_detect(Grid, "Sourcewater"))

data_plotting_source <- data_plotting_source %>% 
    mutate(
    plot_1 = case_when(
      Plot == "FW" ~ "Freshwater",
      Plot == "SW" ~ "Saltwater",
      Plot == "ESTUARY" ~ "Estuary",
      TRUE ~ Plot  # keep everything else unchanged
    )
  ) %>%
  mutate(plot_1 = factor(plot_1, levels = c("Control", "Freshwater", "Saltwater", "Estuary")))


##Boxplot Summarize SO4 and Cl by Date, Site, and Plot
viz_H2S_boxplot <- ggplot(data_plotting, aes(x = Collection_Date_YYYYMMDD,  y = H2S_mean_uM, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  # scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2", 
    "Estuary" = "darkred"
  ), labels = c("Control", "Freshwater", "Saltwater", "Estuary")) +
  theme_classic() +
  labs(x = " ", y = "Sulfide (uM)", title = "TEMPEST Sourcewater: Sulfide") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_H2S_boxplot

# # Save as a PNG
# png("COMPASS_SynopticCB_Cl_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_cl_boxplot)  # For ggplot objects, use print()
# dev.off() 
# 
# # Save as a PNG
# png("COMPASS_SynopticCB_SO4_alldata.png", width = 2000, height = 1200, res = 100) 
# print(viz_so4_boxplot)  # For ggplot objects, use print()
# dev.off() 

```

## Boxplot summary by Grid
```{r Porewater grid boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Boxplot Summarize SO4 and Cl by Date, Site, and Zone
viz_H2S_boxplot <- ggplot(data_plotting, aes(x = Grid,  y = H2S_mean_uM, fill = plot_1)) +
  geom_boxplot() + 
  facet_wrap( ~ plot_1, scales="free_y", ncol = 1) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) +
  theme_classic() +
  labs(x = "Grid ID", y = "Sulfide (uM)", title = "TEMPEST Porewater: Sulfide by Grid") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_H2S_boxplot

```

##Calculate summary metrics for Porewater
```{r Porewater additional summary metrics, include=FALSE, dpi = 300}
# It would also be great to add to our collation codes a summary chunk that will calculate value ranges, means, and standard deviations (and errors) for each site, zone, depth for each year; also for each site zone, depth overall; and for each site and zone overall so that we have some easy access to metrics. Was just thinking this might be helpful today when re-collating the DIC data.

colnames(data_plotting)

##Metrics for each zone for each year
summary_by_year <- data_plotting %>%
  group_by(Year, plot_1) %>%
   reframe(
     min_H2S_ppm = min(H2S_mean_uM, na.rm=TRUE),
     max_H2S_ppm = max(H2S_mean_uM, na.rm=TRUE), 
     mean_H2S_ppm = mean(H2S_mean_uM, na.rm=TRUE), 
     sd_H2S_ppm = sd(H2S_mean_uM, na.rm=TRUE)) %>%
  mutate(across(where(is.numeric), round, digits = 4)) 

##Metrics for each zone for all years
summary_all_years <- data_plotting %>%
  group_by(plot_1) %>%
   reframe(
     min_H2S_ppm = min(H2S_mean_uM, na.rm=TRUE),
     max_H2S_ppm = max(H2S_mean_uM, na.rm=TRUE), 
     mean_H2S_ppm = mean(H2S_mean_uM, na.rm=TRUE), 
     sd_H2S_ppm = sd(H2S_mean_uM, na.rm=TRUE)) %>% 
  mutate(across(where(is.numeric), round, digits = 4)) 

```

##Calculate summary metrics for Sourcewater
```{r Sourcewater additional summary metrics, include=FALSE, dpi = 300}
# It would also be great to add to our collation codes a summary chunk that will calculate value ranges, means, and standard deviations (and errors) for each site, Grid, depth for each year; also for each site Grid, depth overall; and for each site and Grid overall so that we have some easy access to metrics. Was just thinking this might be helpful today when re-collating the DIC data.

colnames(data_plotting_source)

##Metrics for each Grid for each year
summary_by_year_source <- data_plotting_source %>%
  group_by(Year, plot_1) %>%
   reframe(
     min_H2S_ppm = min(H2S_mean_uM, na.rm=TRUE),
     max_H2S_ppm = max(H2S_mean_uM, na.rm=TRUE), 
     mean_H2S_ppm = mean(H2S_mean_uM, na.rm=TRUE), 
     sd_H2S_ppm = sd(H2S_mean_uM, na.rm=TRUE)) %>%
  mutate(across(where(is.numeric), round, digits = 4)) 

```


## Summary Tables for Porewater
```{r Porewater summary table metrics, echo = FALSE, message = FALSE, dpi = 100}

library(kableExtra)

kable(
  summary_by_year[, c("Year", "plot_1", "min_H2S_ppm",	"max_H2S_ppm",	"mean_H2S_ppm", "sd_H2S_ppm")],  
  format = "latex",
  booktabs = TRUE,
  caption = "Porewater Summary Statistics by Year"
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)

kable(
  summary_all_years[, c("plot_1", "min_H2S_ppm",	"max_H2S_ppm",	"mean_H2S_ppm", "sd_H2S_ppm")],  
  format = "latex",
  booktabs = TRUE,
  caption = "Porewater Summary Statistics by Plot"
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)

```

## Summary Tables for Sourcewater
```{r Sourcewater summary table metrics, echo = FALSE, message = FALSE, dpi = 100}

library(kableExtra)

kable(
  summary_by_year_source[, c("Year", "plot_1", "min_H2S_ppm",	"max_H2S_ppm",	"mean_H2S_ppm", "sd_H2S_ppm")],  
  format = "latex",
  booktabs = TRUE,
  caption = "Sourcewater Summary Statistics by Year"
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)

```


##Write out summarized data
```{r write out summarized data, include=FALSE}

#write out a csv of the Porewater Summary Statistics by Year to the main folder: 
write.csv(summary_by_year, "COMPASS_TEMPEST_PW_H2S_Summarized_By_Year.csv")

#write out a csv of the Porewater Summary Statistics by Grid to the main folder: 
write.csv(summary_all_years, "COMPASS_TEMPEST_PW_H2S_Summarized_Years_Combined.csv")

#write out a csv of the Sourcewater Summary Statistics by Year to the main folder: 
write.csv(summary_by_year_source, "COMPASS_TEMPEST_SOURCE_H2S_Summarized_By_Year.csv")

```



